---
title: "F2_map_construction"
author: "Ruijuan Li"
date: "4/18/2017"
output: html_document
---

# Goal
The goal of this analysis is to build genetic map using F2 genotyping information, F2 genotyping was done based on the SNPs we identified from its parents 

# install & load package 
```{r}
# install.packages("onemap")
# install.packages("tkrplot", type="source")
library(tcltk)
library(tkrplot)
library(onemap) 
library(ggplot2)
library(reshape)
library(dplyr)
```

# import & foramt data
```{r}
F2_geno_data <- read.table("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/data/F2_Final_SNP_Calls", header = T)
head(F2_geno_data) 

rownames(F2_geno_data) <- paste(F2_geno_data$CHROM, F2_geno_data$POS, sep = "_")
F2_geno_data <- F2_geno_data[, -c(1:6)]
dim(F2_geno_data) # 18226   166 
colnames(F2_geno_data) <- gsub("([[:print:]]+)(\\.)([[:digit:]])", "\\3", colnames(F2_geno_data))
head(F2_geno_data)  

### filter based on missing data less than 20% missing data for each marker... don't need to, because the data is already filtered based on missing rate 
# F2_geno_data$num <- apply(F2_geno_data,1, function(x) sum(!is.na(x)))
# F2_geno_data_filter <- (F2_geno_data$num/94)
# sum(is.na(F2_geno_data[1,])) 

## sample 200 markers to try this method ... 
# test <- sample_n(F2_geno_data, size = 200, replace = F) 
# dim(test)
```

# remove redundant SNPs

some SNPs should have exactly the same genotype across all individuals, they are redundant in terms of genetic map construction, so they should be removed. find those SNPs by doing correlation test. 

```{r}
# problem: 
# 1)  12 markers are not polymorphism across individuals 
# 2)  because there are missing data present in the matrix, when computing for correlation, only non missing data are accepted, so the pairwise correlation resulted in many markers that end up with same genotypes for the shared non-missing data, this cause error in cor() functon, I suppressed the "0 as standard deviation" warning message. finally found out there are 43422 pairs of SNPs that have this kind of problem 

# script "identify_identical_SNP.R" deal with problem 1 & 2 
# output file /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/F2/output/F2_SNP_correlation.Rdata is the correlation matrix for individuals in F2 population... 

# 3) which marker should keep when their correlation is 1? this is more complicated than I thought... hang on.... leave for later... 579 pairs of SNPs have correlation of 1 

# ####### pick highly correlated values, below are just practice test, with potential useful code for future use...  
# mat = matrix(runif(12), 3) 
# mat
# colnames(mat) <- c("test1", "test2", "test3", "test4")
# ?runif
# cor_mat <- cor(mat)
# cor_mat
# class(cor_mat)
# cor_mat[1,3] <- NA
# cor_mat
# nrow(which(cor_mat > 0.9 & lower.tri(cor_mat), arr.ind = T, useNames = T)) 
```

# format data for onemap 
```{r}
# F2_geno_data <- test

# formating data for onemap 
F2_geno_data <- as.matrix(F2_geno_data)
F2_geno_data[is.na(F2_geno_data)] <- "-"
F2_geno_data <- gsub("0/0", "A", F2_geno_data) 
F2_geno_data <- gsub("0/1", "H", F2_geno_data)
F2_geno_data <- gsub("1/1", "B", F2_geno_data)
# F2_geno_data <- t(F2_geno_data)

F2_geno_data <- as.data.frame(F2_geno_data)
dim(F2_geno_data) #  
write.table(F2_geno_data, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/data/F2_geno_data.txt")
# change file format in linux 
# cat F2_geno_data.txt | sed 's/"//g' | awk '{first = $1; $1 = ""; print $0}' | sed 's/ //g' > tmp
# tail -18226 tmp > tmp.1 

write.table(rownames(F2_geno_data), file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/data/marker_info.txt" )
# cat marker_info.txt | awk '{print "*"$2}' | tail -18226 | sed 's/"//g' > marker_info_reform.txt

# paste marker_info_reform.txt tmp.1 | awk '{print $1 " " $2}' > F2_geno_for_one_map.txt 
# cat header_one_map_input F2_geno_for_one_map.txt > F2_geno_for_one_map_final.txt 

# cross_pheno_data <- t(read.csv("F2_415_pheno.csv", header = T))
# cross_pheno_data
# cross_pheno_data <- as.data.frame(cross_pheno_data)
# cross_pheno_data
# change file format in linux 
# cat cross_pheno_data.txt | sed 's/\"\.\"/\"\-\"/g'  | sed 's/"//g' | awk '{print "*" $0}' > pheno.txt 

########### real analysis start from here ################################
data <- read.mapmaker(file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/data/F2_geno_for_one_map_final.txt")
# data
# save(data, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/data.Rdata") # the data saved here can be used later for map construction 

load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/data.Rdata")
# estimate two-point rf
twopts.f2 <- rf.2pts(data, LOD = 6, max.rf = 0.25) ##### run this on whitney....  
?rf.2pts
twopts.f2

# assign marker to linkage groups
mark.all.f2 <- make.seq(twopts.f2, "all")
mark.all.f2
?make.seq
LGs.f2 <- group(mark.all.f2, LOD = 6, max.rf = 0.25)
LGs.f2

#################### LG1 ##############################
LG1.f2 <- make.seq(LGs.f2, 1)
LG1.f2

# use touchdown
LG1.f2.ord <- order.seq(input.seq = LG1.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG1.f2.ord

# safe mode order
LG1.f2.safe <- make.seq(LG1.f2.ord, "safe")
LG1.f2.safe
# all order
LG1.f2.final <- make.seq(LG1.f2.ord, "force")
LG1.f2.final

# ripple
ripple.seq(LG1.f2.final, ws = 5, LOD = 6) 

################# LG2 ##################################
LG2.f2 <- make.seq(LGs.f2, 2)
LG2.f2

# use touchdown
LG2.f2.ord <- order.seq(input.seq = LG2.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG2.f2.ord

# safe mode order
LG2.f2.safe <- make.seq(LG2.f2.ord, "safe")
LG2.f2.safe
# all order
LG2.f2.final <- make.seq(LG2.f2.ord, "force")
LG2.f2.final

# ripple
ripple.seq(LG2.f2.final, ws = 5, LOD = 6) 

######################## LG3 ########################

LG3.f2 <- make.seq(LGs.f2, 3)
LG3.f2

# use touchdown
LG3.f2.ord <- order.seq(input.seq = LG3.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG3.f2.ord

# safe mode order
LG3.f2.safe <- make.seq(LG3.f2.ord, "safe")
LG3.f2.safe
# all order
LG3.f2.final <- make.seq(LG3.f2.ord, "force")
LG3.f2.final

# ripple
ripple.seq(LG3.f2.final, ws = 3, LOD = 6) 

######################## LG4 ######################
########### use Kosambi map function ################
set.map.fun(type = "kosambi")

LG4.f2 <- make.seq(LGs.f2, 4)
LG4.f2

# use touchdown
LG4.f2.ord <- order.seq(input.seq = LG4.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG4.f2.ord

# safe mode order
LG4.f2.safe <- make.seq(LG4.f2.ord, "safe")
LG4.f2.safe
# all order
LG4.f2.final <- make.seq(LG4.f2.ord, "force")
LG4.f2.final

######################################
#Markers               Position           
#
#60 FPBPN1720              0.00           
#59 FPBPN1183             21.79           
#58 FPBPN1466             23.96           
#22 FPBRN0403             99.48           
#23 FPBPN0589            123.34           
#24 FPBPN0595            124.96           
#25 FPBPN0006            132.46           
#26 FPBPN0294            138.91           
#27 FPBPN0205            151.91           
#28 FPBRN0426            171.45           
#29 FPBPN1534            177.44           
#30 FPBPN1503            178.58           
#31 FPBPN1712            187.25           
#32 FPBRN0434            192.02           
#33 FPBRN0876            207.88           
#40 FPBPN1533            270.61           
#41 FPBPN0296            272.67           
#42 FPBPN0399            284.70           
#43 FPBPN1863            304.96           
#44 FPBRN0153            311.49           
#45 FPBPN0005            324.22           
#46 FPBRN0154            325.08           
#47 FPBPN0291            330.16           
#48 FPBPN0283            331.39           
#53 FPBPN1083            335.94           
#49 FPBPN0790            339.05           
#50 FPBPN1637            341.32           
#51 FPBPN1563            345.21           
#52 FPBPN1651            346.72           
#54 FPBPN1982            357.41           
#55 FPBPN1702            360.15           
#56 FPBPN0045            365.75           
#34 FPBPN1414            380.72           
#35 FPBPN0559            390.62           
#36 FPBPN1369            395.71           
#37 FPBPN0197            400.49           
#38 FPBPN0164            409.48           
#39 FPBPN1107            419.04           
#57 FPBPN0082            440.07           
#
#39 markers            log-likelihood: -2303.75 
###############################################

# ripple
ripple.seq(LG4.f2.final, ws = 5, LOD = 6)

################### use haldane map function ##########
set.map.fun(type = "haldane")

LG4h.f2 <- make.seq(LGs.f2, 4)
LG4h.f2

# use touchdown
LG4h.f2.ord <- order.seq(input.seq = LG4h.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG4h.f2.ord

# safe mode order
LG4h.f2.safe <- make.seq(LG4h.f2.ord, "safe")
LG4h.f2.safe
# all order
LG4h.f2.final <- make.seq(LG4h.f2.ord, "force")
LG4h.f2.final
### almost no difference between haldane & kosambi fxn, except that 
# kosambi gives shorter map distance... the marker order is a little 
# different (not much) from FnP's order got from mapdisto kosambi funcion. 
######################################################

###################### LG5 ########################

LG5.f2 <- make.seq(LGs.f2, 5)
LG5.f2

# use touchdown
LG5.f2.ord <- order.seq(input.seq = LG5.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG5.f2.ord

# safe mode order
LG5.f2.safe <- make.seq(LG5.f2.ord, "safe")
LG5.f2.safe
# all order
LG5.f2.final <- make.seq(LG5.f2.ord, "force")
LG5.f2.final

################# LG6 ##############################
LG6.f2 <- make.seq(LGs.f2, 6)
LG6.f2

# use touchdown
LG6.f2.ord <- order.seq(input.seq = LG6.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG6.f2.ord

# safe mode order
LG6.f2.safe <- make.seq(LG6.f2.ord, "safe")
LG6.f2.safe
# all order
LG6.f2.final <- make.seq(LG6.f2.ord, "force")
LG6.f2.final

################# LG7 ##############################
LG7.f2 <- make.seq(LGs.f2, 7)
LG7.f2

# use touchdown
LG7.f2.ord <- order.seq(input.seq = LG7.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG7.f2.ord

# safe mode order
LG7.f2.safe <- make.seq(LG7.f2.ord, "safe")
LG7.f2.safe
# all order
LG7.f2.final <- make.seq(LG7.f2.ord, "force")
LG7.f2.final

################# LG8 ##############################
LG8.f2 <- make.seq(LGs.f2, 8)
LG8.f2

# use touchdown
LG8.f2.ord <- order.seq(input.seq = LG8.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG8.f2.ord

# safe mode order
LG8.f2.safe <- make.seq(LG8.f2.ord, "safe")
LG8.f2.safe
# all order
LG8.f2.final <- make.seq(LG8.f2.ord, "force")
LG8.f2.final

################# LG9 ##############################
LG9.f2 <- make.seq(LGs.f2, 9)
LG9.f2

# use touchdown
LG9.f2.ord <- order.seq(input.seq = LG9.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG9.f2.ord

# safe mode order
LG9.f2.safe <- make.seq(LG9.f2.ord, "safe")
LG9.f2.safe
# all order
LG9.f2.final <- make.seq(LG9.f2.ord, "force")
LG9.f2.final

################# LG10 ##############################
LG10.f2 <- make.seq(LGs.f2, 10)
LG10.f2

# use touchdown
LG10.f2.ord <- order.seq(input.seq = LG10.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG10.f2.ord

# safe mode order
LG10.f2.safe <- make.seq(LG10.f2.ord, "safe")
LG10.f2.safe
# all order
LG10.f2.final <- make.seq(LG10.f2.ord, "force")
LG10.f2.final

################# LG11/A09 ##############################
LG11.f2 <- make.seq(LGs.f2, 11)
LG11.f2

# use touchdown
LG11.f2.ord <- order.seq(input.seq = LG11.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG11.f2.ord

# safe mode order
LG11.f2.safe <- make.seq(LG11.f2.ord, "safe")
LG11.f2.safe
# all order
LG11.f2.final <- make.seq(LG11.f2.ord, "force")
LG11.f2.final

################# LG12/A10 ##############################
LG12.f2 <- make.seq(LGs.f2, 12)
LG12.f2

# use touchdown
LG12.f2.ord <- order.seq(input.seq = LG12.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG12.f2.ord

# safe mode order
LG12.f2.safe <- make.seq(LG12.f2.ord, "safe")
LG12.f2.safe
# all order
LG12.f2.final <- make.seq(LG12.f2.ord, "force")
LG12.f2.final

################# LG13/C01 ##############################
LG13.f2 <- make.seq(LGs.f2, 13)
LG13.f2

# use touchdown
LG13.f2.ord <- order.seq(input.seq = LG13.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG13.f2.ord

# safe mode order
LG13.f2.safe <- make.seq(LG13.f2.ord, "safe")
LG13.f2.safe
# all order
LG13.f2.final <- make.seq(LG13.f2.ord, "force")
LG13.f2.final

################# LG14/C01-1 ##############################
LG14.f2 <- make.seq(LGs.f2, 14)
LG14.f2

# use touchdown
LG14.f2.ord <- order.seq(input.seq = LG14.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG14.f2.ord

# safe mode order
LG14.f2.safe <- make.seq(LG14.f2.ord, "safe")
LG14.f2.safe
# all order
LG14.f2.final <- make.seq(LG14.f2.ord, "force")
LG14.f2.final

################# LG15/C02 ##############################
LG15.f2 <- make.seq(LGs.f2, 15)
LG15.f2

# use touchdown
LG15.f2.ord <- order.seq(input.seq = LG15.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG15.f2.ord

# safe mode order
LG15.f2.safe <- make.seq(LG15.f2.ord, "safe")
LG15.f2.safe
# all order
LG15.f2.final <- make.seq(LG15.f2.ord, "force")
LG15.f2.final

################# LG16/C02-2 ##############################
LG16.f2 <- make.seq(LGs.f2, 16)
LG16.f2

# use touchdown
LG16.f2.ord <- order.seq(input.seq = LG16.f2, n.init = 5,
                         subset.search = "twopt",
                         twopt.alg = "rcd", THRES = 6,
                         draw.try = T, wait = 1,
                         touchdown = T)

LG16.f2.ord

# safe mode order
LG16.f2.safe <- make.seq(LG16.f2.ord, "safe")
LG16.f2.safe
# all order
LG16.f2.final <- make.seq(LG16.f2.ord, "force")
LG16.f2.final

################# LG17/C02-3 ##############################
LG17.f2 <- make.seq(LGs.f2, 17)
LG17.f2

# use touchdown
LG17.f2.ord <- order.seq(input.seq = LG17.f2, n.init = 2,
                         subset.search = "twopt",
                         twopt.alg = "rcd", THRES = 6,
                         draw.try = T, wait = 1,
                         touchdown = T)

LG17.f2.ord

# safe mode order
LG17.f2.safe <- make.seq(LG17.f2.ord, "safe")
LG17.f2.safe
# all order
LG17.f2.final <- make.seq(LG17.f2.ord, "force")
LG17.f2.final

################# LG18/C003 ##############################
LG18.f2 <- make.seq(LGs.f2, 18)
LG18.f2

# use touchdown
LG18.f2.ord <- order.seq(input.seq = LG18.f2, n.init = 5,
                         subset.search = "twopt",
                         twopt.alg = "rcd", THRES = 6,
                         draw.try = T, wait = 1,
                         touchdown = T)

LG18.f2.ord

# safe mode order
LG18.f2.safe <- make.seq(LG18.f2.ord, "safe")
LG18.f2.safe
# all order
LG18.f2.final <- make.seq(LG18.f2.ord, "force")
LG18.f2.final

################# LG19/C04 ##############################
LG19.f2 <- make.seq(LGs.f2, 19)
LG19.f2

# use touchdown
LG19.f2.ord <- order.seq(input.seq = LG19.f2, n.init = 5,
                         subset.search = "twopt",
                         twopt.alg = "rcd", THRES = 6,
                         draw.try = T, wait = 1,
                         touchdown = T)

LG19.f2.ord

# safe mode order
LG19.f2.safe <- make.seq(LG19.f2.ord, "safe")
LG19.f2.safe
# all order
LG19.f2.final <- make.seq(LG19.f2.ord, "force")
LG19.f2.final

################# LG20/C05 ##############################
LG20.f2 <- make.seq(LGs.f2, 20)
LG20.f2

# use touchdown
LG20.f2.ord <- order.seq(input.seq = LG20.f2, n.init = 5,
                         subset.search = "twopt",
                         twopt.alg = "rcd", THRES = 6,
                         draw.try = T, wait = 1,
                         touchdown = T)

LG20.f2.ord

# safe mode order
LG20.f2.safe <- make.seq(LG20.f2.ord, "safe")
LG20.f2.safe
# all order
LG20.f2.final <- make.seq(LG20.f2.ord, "force")
LG20.f2.final

################# LG21/C06 ##############################
LG21.f2 <- make.seq(LGs.f2, 21)
LG21.f2

# use touchdown
LG21.f2.ord <- order.seq(input.seq = LG21.f2, n.init = 5,
                         subset.search = "twopt",
                         twopt.alg = "rcd", THRES = 6,
                         draw.try = T, wait = 1,
                         touchdown = T)

LG21.f2.ord

# safe mode order
LG21.f2.safe <- make.seq(LG21.f2.ord, "safe")
LG21.f2.safe
# all order
LG21.f2.final <- make.seq(LG21.f2.ord, "force")
LG21.f2.final

################# LG22/C07 ##############################
LG22.f2 <- make.seq(LGs.f2, 22)
LG22.f2

# use touchdown
LG22.f2.ord <- order.seq(input.seq = LG22.f2, n.init = 5,
                         subset.search = "twopt",
                         twopt.alg = "rcd", THRES = 6,
                         draw.try = T, wait = 1,
                         touchdown = T)

LG22.f2.ord

# safe mode order
LG22.f2.safe <- make.seq(LG22.f2.ord, "safe")
LG22.f2.safe
# all order
LG22.f2.final <- make.seq(LG22.f2.ord, "force")
LG22.f2.final

################# LG23/C08 ##############################
LG23.f2 <- make.seq(LGs.f2, 23)
LG23.f2

# use touchdown
LG23.f2.ord <- order.seq(input.seq = LG23.f2, n.init = 5,
                         subset.search = "twopt",
                         twopt.alg = "rcd", THRES = 6,
                         draw.try = T, wait = 1,
                         touchdown = T)

LG23.f2.ord

# safe mode order
LG23.f2.safe <- make.seq(LG23.f2.ord, "safe")
LG23.f2.safe
# all order
LG23.f2.final <- make.seq(LG23.f2.ord, "force")
LG23.f2.final

################# LG24/C09 ##############################
LG24.f2 <- make.seq(LGs.f2, 24)
LG24.f2

# use touchdown
LG24.f2.ord <- order.seq(input.seq = LG24.f2, n.init = 5,
                         subset.search = "twopt",
                         twopt.alg = "rcd", THRES = 6,
                         draw.try = T, wait = 1,
                         touchdown = T)

LG24.f2.ord

# safe mode order
LG24.f2.safe <- make.seq(LG24.f2.ord, "safe")
LG24.f2.safe
# all order
LG24.f2.final <- make.seq(LG24.f2.ord, "force")
LG24.f2.final

################# LG25/C09-1 ##############################
LG25.f2 <- make.seq(LGs.f2, 25)
LG25.f2

# use touchdown
LG25.f2.ord <- order.seq(input.seq = LG25.f2, n.init = 5,
                         subset.search = "twopt",
                         twopt.alg = "rcd", THRES = 6,
                         draw.try = T, wait = 1,
                         touchdown = T)

LG25.f2.ord

# safe mode order
LG25.f2.safe <- make.seq(LG25.f2.ord, "safe")
LG25.f2.safe
# all order
LG25.f2.final <- make.seq(LG25.f2.ord, "force")
LG25.f2.final




###################################################
########### drop marker with double crossover ###########
data <- read.mapmaker(file = "F2_415_dropp_geno_Ruijuan.txt")
data

# estimate two-point rf
twopts.f2 <- rf.2pts(data, LOD = 3, max.rf = 0.5)
?rf.2pts

# assign marker to linkage groups
mark.all.f2 <- make.seq(twopts.f2, "all")
?make.seq
LGs.f2 <- group(mark.all.f2, LOD = 3, max.rf = 0.5)
LGs.f2

#################### LG1 ##############################
LG1.f2 <- make.seq(LGs.f2, 1)
LG1.f2

# use touchdown
LG1.f2.ord <- order.seq(input.seq = LG1.f2, n.init = 5,
                        subset.search = "twopt",
                        twopt.alg = "rcd", THRES = 6,
                        draw.try = T, wait = 1,
                        touchdown = T)

LG1.f2.ord

# safe mode order
LG1.f2.safe <- make.seq(LG1.f2.ord, "safe")
LG1.f2.safe
# all order
LG1.f2.final <- make.seq(LG1.f2.ord, "force")
LG1.f2.final

# ripple
ripple.seq(LG1.f2.final, ws = 5, LOD = 3) 
``` 
