---
title: "KIAT_F1_data_analysis"
author: "Ruijuan Li"
date: "August 12, 2016"
output: html_document
---

```{r}
# 1) download file from ftp using ftp_file_transfer.sh
# 1.1) unzip files using gunzip *  
# 2) get first 1000 sequences for each file extract_1000_reads.sh, actually worked with all data instead of the 1st 1000 seqs
# 3) QC file using fastQC: fastqc *.fq -o /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/2016_summer/fastqc_result
# 4) download Brassica_napus.annotation_v5.cds.fa to whitney 
# 5) collapse fastqc summary result into one file & get overrepresented sequences using ./make_summary_report.sh 

# 6) reformat summary.all.txt file 
setwd("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/")
fastQC_summary <- read.delim("./fastQC_result/summary_all.txt", header = F)
dim(fastQC_summary)
head(fastQC_summary)
tail(fastQC_summary)

fastQC_summary_wide <- reshape(fastQC_summary, timevar="V2",idvar="V3",direction="wide")
colnames(fastQC_summary_wide) <- gsub("(V1)(.)([[:print:]]+)","\\3",colnames(fastQC_summary_wide))
dim(fastQC_summary_wide)

write.csv(fastQC_summary_wide, file = "fastQC_summary_wide.csv")

# 7) checked the source of overrepresented sequences using web nucleotide blast tool, found that most of them are from 
# cholorplast, mitochondra, rRNA, tansposon, etc. Some are illumina or TruSeq adapters 

# blast sequences against Trimmomatic adapters suggest that TruSeq3-PE-2.fa should be used to trimm the data. (Q: how to deal with the others? will they affect the final result? search... )

# Qs on plstids seqs contamination: 1) why are there these sequences? check the lib making step 2) will they affect downstream analysis if kept? CDS mapping VS genome mapping 3) If yes, how to remove them? (I need to figure these Qs out.)

# 8) using kallisto to build the reference cds index 
# kallisto index -k 19 -i Brassica_napus.annotation_v5.cds.19.kai /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5.cds.fa

# 9) map to reference cds index and get read count file using bunchrun_kallisto.sh (using est_counts from Kallisto)
```

```{r}
##### STAR is only installed on coloma!!! 
# 1) get B.napus reference genome: Brassica_napus_v4.1.chromosomes.fa 
# make STAR reference index: STAR --runMode genomeGenerate --genomeDir star_nome/ --genomeFastaFiles Brassica_napus_v4.1.chromosomes.fa --runThreadN 2 
# mapping using run_star.sh 
 
# STAR --genomeDir /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/star_genome --readFilesIn /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/2016_summer/automate_transfer/F1/414F1-gea-1_1.fq /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/2016_summer/automate_transfer/F1/414F1-gea-1_2.fq --runThreadN 6

### running stranded fr-secondstrand cufflinks 
# screen -r 35026.ttys020.coloma (running cufflinks now)  
```

# data summary statisitics read count number & expressed genes 
```{r}
F1.read.count.normalized <- read.table("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/read.count.normalized.tsv", header = T, check.names = F)
rownames(F1.read.count.normalized) <- F1.read.count.normalized[,1]
F1.read.count.normalized <- F1.read.count.normalized[,-1]

# format data 
head(F1.read.count.normalized)
dim(F1.read.count.normalized) # 101040     24 
colnames(F1.read.count.normalized)
colnames(F1.read.count.normalized) <- sub("_2.fq","",colnames(F1.read.count.normalized),fixed = TRUE) # remove _2.fq 

# sample description  
new_sample_ID.F1

# replace sample ID 
colnames(F1.read.count.normalized) <- new_sample_ID.F1
head(F1.read.count.normalized)
save(F1.read.count.normalized,file="/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/F1.read.count.normalized.Rdata")

# check number of genes with reads number > 10 in each lib # make summary table 
expressed_gene_number.F1 <- 
sapply(colnames(F1.read.count.normalized), function(RNAlib){
  sum(F1.read.count.normalized[,RNAlib]>3)
}
)
expressed_gene_number.F1 <- as.data.frame(expressed_gene_number.F1)
expressed_gene_number.F1

# top 10 highly expressed genes in each lib 
most_highly_expressed_genes.F1 <- 
sapply(colnames((F1.read.count.normalized)), function(RNAlib){
  rownames(head(F1.read.count.normalized[order(F1.read.count.normalized[,RNAlib],decreasing = TRUE),], 10))
}
)
# write.csv(t(most_highly_expressed_genes.F1), file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/most_highly_expressed_genes.F1.csv")

temp <- sample_des.F1[,c(5,6,7,10)]
rownames(temp) <- paste(sample_des.F1[,5], sample_des.F1[,6], sample_des.F1[,7], sep = "_")
sample_statistics.F1 <- merge(temp, expressed_gene_number.F1, by=0) 
colnames(sample_statistics.F1) <- c("sample_ID", "Cultivar", "Tissue", "rep", "total_reads", "expressed_gene_number")
head(sample_statistics.F1)
colnames(sample_statistics.F1)
# write.csv(sample_statistics.F1, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/sample_statisitcs.F1.csv")
```

# Expression analysis 1) formatting data   
```{r}
F1.read.count <- read.table("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/read.count.tsv", header = T, check.names = F)

rownames(F1.read.count) <- F1.read.count[,1]
F1.read.count <- F1.read.count[,-1]

# format data 
head(F1.read.count)
dim(F1.read.count) # 101040     24 
colnames(F1.read.count)
colnames(F1.read.count) <- sub("_2.fq","",colnames(F1.read.count),fixed = TRUE) # remove _2.fq 

# sample description  
sample_des.F1 <- read.csv("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/F1_summary.csv")
dim(sample_des.F1)
sorted_sample_des.F1 <- sample_des.F1[order(sample_des.F1$Sample.ID),]

sorted_sample_des.F1[,5:8]
new_sample_ID.F1 <- paste(sorted_sample_des.F1$cultivar, sorted_sample_des.F1$Stage, sorted_sample_des.F1$rep, sep = "_")
new_sample_ID.F1

# calculate GB size of libs 
mean(as.numeric(sub("Gb","",sample_des.F1$TotalBases.Gb.))) 
min(as.numeric(sub("Gb","",sample_des.F1$TotalBases.Gb.))) 
max(as.numeric(sub("Gb","",sample_des.F1$TotalBases.Gb.))) 

# replace sample ID 
colnames(F1.read.count)
colnames(F1.read.count) <- new_sample_ID.F1
head(F1.read.count)
save(F1.read.count,file="/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/F1.read.count.Rdata")
```

# expression analysis 2) set up sample description 
```{r}
# load necessary libs & functions 
library(edgeR)
library(ggplot2)

# filter based on read count, assign group, normalize, design matrix, calculate dispersion   
# set up group 
load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/F1.read.count.Rdata")
F1.read.count <- F1.read.count[,colSums(F1.read.count) > 1000000]  
dim(F1.read.count) # 101040     24 
colnames(F1.read.count)

F1.read.count.sample<-data.frame(file=colnames(F1.read.count),
                                         batch=factor(gsub("(414F1|415F1)(_)(flowering|early-silique|late-silique|Young)(_)(1|2|3)","\\5",colnames(F1.read.count))),  
                                         genotype=factor(gsub("(414F1|415F1)(_)(flowering|early-silique|late-silique|Young)(_)(1|2|3)","\\1",colnames(F1.read.count))),	
                                         tissue=factor(gsub("(414F1|415F1)(_)(flowering|early-silique|late-silique|Young)(_)(1|2|3)","\\3",colnames(F1.read.count))),	
                                         group=factor(gsub("(414F1|415F1)(_)(flowering|early-silique|late-silique|Young)(_)(1|2|3)","\\1\\3",colnames(F1.read.count)))
)

F1.read.count.sample
ftable(F1.read.count.sample,row.vars="tissue",col.vars=c("batch","genotype"))

# filter based on read count 
F1.read.count.small <- F1.read.count[rowSums(F1.read.count > 10) >= 3,]
dim(F1.read.count.small) # 60176  24 
save(F1.read.count.small, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/F1.read.count.small.Rdata")
```

# expression analysis 3) voom transformation # think more about the design in voom transformation... 
```{r}
# voom transformation 
# source("https://bioconductor.org/biocLite.R")
# biocLite("DESeq2")
load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/F1.read.count.small.Rdata")
library("DESeq2")

dds.F1 <- DESeqDataSetFromMatrix(countData = round(F1.read.count.small), colData = F1.read.count.sample, design = ~ batch + genotype*tissue)

vsd.F1 <- varianceStabilizingTransformation(dds.F1)
vstMat.F1 <- assay(vsd.F1)
colnames(vstMat.F1) <- colnames(F1.read.count.small)
save(vstMat.F1, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/vstMat.F1.Rdata")
```

# MDS, clustering 
```{r}
load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/vstMat.F1.Rdata")
load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/F1.read.count.small.Rdata")

# normalize 
dge.data.F1 <- DGEList(counts=F1.read.count.small, group=F1.read.count.sample$group)
dge.data.F1 <- calcNormFactors(dge.data.F1, method = "TMM") 
dge.data.F1$sample
hist(dge.data.F1$samples$norm.factors)

# MDS to check sample seperation, also using dengrogram.  
plotMDS(dge.data.F1, method = "bcv",labels = rownames(dge.data.F1$samples))
mds.F1 <- plotMDS(dge.data.F1, method = "bcv",labels = dge.data.F1$samples$group)

x.F1 <- as.data.frame(mds.F1$x)
y.F1 <- as.data.frame(mds.F1$y)
distance_matrix.F1 <- merge(x.F1, y.F1, by="row.names")
distance_matrix.F1$group <- gsub("(414F1|415F1)(_)(flowering|early-silique|late-silique|Young)(_)(1|2|3)","\\1\\3",distance_matrix.F1$Row.names)
distance_matrix.F1$gt <- gsub("(414F1|415F1)(_)(flowering|early-silique|late-silique|Young)(_)(1|2|3)","\\1",distance_matrix.F1$Row.names)
distance_matrix.F1$tissue <- gsub("(414F1|415F1)(_)(flowering|early-silique|late-silique|Young)(_)(1|2|3)","\\3",distance_matrix.F1$Row.names)

colnames(distance_matrix.F1) <- c("lib","x","y","group","gt","tissue")
head(distance_matrix.F1)

# making color MDS figure 
p <- ggplot(data = distance_matrix.F1)
p <- p + geom_point(aes(x, y, color=factor(tissue)), size=3) 
p <- p + labs(y = "BCV distance 2", x="BCV distance 1")
# p <- p + facet_grid(~gt)
p

# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/MDS_2.png", width = 11, height = 8)

# clustering to check sample seperation 
# install.packages("ggdendro")
library(ggdendro)
library(pvclust)

hc.F1 <- hclust(dist(t(vstMat.F1[1:1000,])))
# test <- ggdendrogram(hc.F1, theme_dendro = F, rotate = T)

ggdata <- dendro_data(as.dendrogram(hc.F1))
ggdata$labels$gt <- gsub("(414F1|415F1)(_)(flowering|early-silique|late-silique|Young)(_)(1|2|3)","\\1",ggdata$labels$label) 
ggdata$labels$tissue <- gsub("(414F1|415F1)(_)(flowering|early-silique|late-silique|Young)(_)(1|2|3)","\\3",ggdata$labels$label)
ggdata$labels$group <- paste(ggdata$labels$gt, ggdata$labels$tissue, sep = "_")
ggdata$labels

# start ggplot here 
p2 <- ggplot(data = segment(ggdata))
p2 <- p2 + geom_segment(aes(x=x, y=y, xend=xend, yend=yend)) + theme_dendro()
p2 <- p2 + geom_text(data = label(ggdata), aes(x = x, y = y, label = label, hjust=0, color=group)) + coord_flip() + scale_y_reverse(expand=c(0.2, 0))
p2
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/clustering.png", width = 13, height = 8)
```

# expression analysis 
```{r}
# pairwise comparison using GLM model 
# design matrix
design.F1 <- model.matrix(~0+group, data = F1.read.count.sample)

# w/o batch effect
dge.data.F1 <- estimateGLMCommonDisp(dge.data.F1, design.F1,verbose = TRUE) # Disp = 0.08264 , BCV = 0.2875 
dge.data.F1 <- estimateGLMTrendedDisp(dge.data.F1,design.F1)
dge.data.F1 <- estimateGLMTagwiseDisp(dge.data.F1,design.F1)
plotBCV(dge.data.F1)

## w/o batch effect
fit.F1 <- glmFit(dge.data.F1, design.F1)
colnames(fit.F1$design)

# between genotypes, 414 as reference level  
lrt.young.F1 <- glmLRT(fit.F1, contrast = c(0,0,0,-1,0,0,0,1))
topTags(lrt.young.F1) 
summary(de.young.F1 <- decideTestsDGE(lrt.young.F1, p=0.05))
DEgene.young.F1 <- topTags(lrt.young.F1,n = Inf)$table[topTags(lrt.young.F1,n = Inf)$table$FDR<0.05,]
# -1  1640
# 0  56152
# 1   2384

lrt.flowering.F1 <- glmLRT(fit.F1, contrast = c(0,-1,0,0,0,1,0,0))
topTags(lrt.flowering.F1)
summary(de.flowering.F1 <- decideTestsDGE(lrt.flowering.F1, p=0.05))
DEgene.flowering.F1 <- topTags(lrt.flowering.F1,n = Inf)$table[topTags(lrt.flowering.F1,n = Inf)$table$FDR<0.05,]
# -1   961
# 0  58526
# 1    689

lrt.early.silique.F1 <- glmLRT(fit.F1, contrast = c(-1,0,0,0,1,0,0,0))
topTags(lrt.early.silique.F1)
summary(de.early.silique.F1 <- decideTestsDGE(lrt.early.silique.F1, p=0.05))
DEgene.early.silique.F1 <- topTags(lrt.early.silique.F1,n = Inf)$table[topTags(lrt.early.silique.F1,n = Inf)$table$FDR<0.05,]
# -1    69
# 0  60061
# 1     46

lrt.late.silique.F1 <- glmLRT(fit.F1, contrast = c(0,0,-1,0,0,0,1,0))
topTags(lrt.late.silique.F1)
summary(de.late.silique.F1 <- decideTestsDGE(lrt.late.silique.F1, p=0.05))
DEgene.late.silique.F1 <- topTags(lrt.late.silique.F1,n = Inf)$table[topTags(lrt.late.silique.F1,n = Inf)$table$FDR<0.05,]
# -1   170
# 0  59840
# 1    166

### reformat for ggplot barplot 
DEGs_number_between_gt <- data.frame(young = as.data.frame(summary(de.young.F1 <- decideTestsDGE(lrt.young.F1, p=0.05)))$Freq,
                                     flowering = as.data.frame(summary(de.flowering.F1 <- decideTestsDGE(lrt.flowering.F1, p=0.05)))$Freq, 
                                     early.silique = as.data.frame(summary(de.early.silique.F1 <- decideTestsDGE(lrt.early.silique.F1, p=0.05)))$Freq,
                                     late.silique = as.data.frame(summary(de.late.silique.F1 <- decideTestsDGE(lrt.late.silique.F1, p=0.05)))$Freq)

rownames(DEGs_number_between_gt) <- c("down", "no", "up")
DEGs_number_between_gt <- DEGs_number_between_gt[c("down", "up"),]
DEGs_number_between_gt

DEGs_number_between_gt.melt <- melt(DEGs_number_between_gt)
DEGs_number_between_gt.melt$DE <- rep(c("down", "up"), 4)
colnames(DEGs_number_between_gt.melt) <- c("tissue", "number", "DE")
DEGs_number_between_gt.melt

# reorder: up 1st down 2nd 
DEGs_number_between_gt.melt$DE <- factor(DEGs_number_between_gt.melt$DE, levels = c("up", "down"))

DEGs_number_between_gt.melt <- DEGs_number_between_gt.melt[order(DEGs_number_between_gt.melt$DE),]
DEGs_number_between_gt.melt

### making ggplot for DEGs 
p.DEGs_number_between_gt <- ggplot(data = DEGs_number_between_gt.melt)
p.DEGs_number_between_gt <- p.DEGs_number_between_gt + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number), stat = "identity")
p.DEGs_number_between_gt <- p.DEGs_number_between_gt + facet_grid(~tissue)
p.DEGs_number_between_gt <- p.DEGs_number_between_gt + labs(y = "number of differentially expressed genes", x = "")

p.DEGs_number_between_gt
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/DEGs_number_between_gt.png", width = 11, height = 8)

colnames(fit.F1$design)
# between developmental stages 
### 414 
lrt.young.vs.flowering.414 <- glmLRT(fit.F1, contrast = c(0,1,0,-1,0,0,0,0))
topTags(lrt.young.vs.flowering.414)
summary(de.young.vs.flowering.414 <- decideTestsDGE(lrt.young.vs.flowering.414, p=0.05))
DEgene.young.vs.flowering.414 <- topTags(lrt.young.vs.flowering.414,n = Inf)$table[topTags(lrt.young.vs.flowering.414,n = Inf)$table$FDR<0.05,]
# -1 12706
# 0  30576
# 1  16894

lrt.flowering.vs.early.silique.414 <- glmLRT(fit.F1, contrast = c(1,-1,0,0,0,0,0,0))
topTags(lrt.flowering.vs.early.silique.414)
summary(de.flowering.vs.early.silique.414 <- decideTestsDGE(lrt.flowering.vs.early.silique.414, p=0.05))
DEgene.flowering.vs.early.silique.414 <- topTags(lrt.flowering.vs.early.silique.414,n = Inf)$table[topTags(lrt.flowering.vs.early.silique.414,n = Inf)$table$FDR<0.05,]
# -1  8920
# 0  44337
# 1   6919

lrt.early.vs.late.silique.414 <- glmLRT(fit.F1, contrast = c(-1,0,1,0,0,0,0,0))
topTags(lrt.early.vs.late.silique.414)
summary(de.early.vs.late.silique.414 <- decideTestsDGE(lrt.early.vs.late.silique.414, p=0.05))
DEgene.early.vs.late.silique.414 <- topTags(lrt.early.vs.late.silique.414,n = Inf)$table[topTags(lrt.early.vs.late.silique.414,n = Inf)$table$FDR<0.05,]
# -1  3049
# 0  53803
# 1   3324

### ggplot 
### reformat for ggplot barplot 
DEGs_number_between_tissue.414 <- data.frame("flowering-vs-young" = as.data.frame(summary(de.young.vs.flowering.414 <- decideTestsDGE(lrt.young.vs.flowering.414, p=0.05)))$Freq, 
                                     "early.slique-vs-flowering"= as.data.frame(summary(de.flowering.vs.early.silique.414 <- decideTestsDGE(lrt.flowering.vs.early.silique.414, p=0.05)))$Freq, 
                                     "late.silique-vs-early.silique" = as.data.frame(summary(de.early.vs.late.silique.414 <- decideTestsDGE(lrt.early.vs.late.silique.414, p=0.05)))$Freq)

rownames(DEGs_number_between_tissue.414) <- c("down", "no", "up")
DEGs_number_between_tissue.414 <- DEGs_number_between_tissue.414[c("down", "up"),]
DEGs_number_between_tissue.414

DEGs_number_between_tissue.414.melt <- melt(DEGs_number_between_tissue.414)
DEGs_number_between_tissue.414.melt$DE <- rep(c("down", "up"), 3)
colnames(DEGs_number_between_tissue.414.melt) <- c("tissue", "number", "DE")
DEGs_number_between_tissue.414.melt

# reorder: up 1st down 2nd 
DEGs_number_between_tissue.414.melt$DE <- factor(DEGs_number_between_tissue.414.melt$DE, levels = c("up", "down"))

DEGs_number_between_tissue.414.melt <- DEGs_number_between_tissue.414.melt[order(DEGs_number_between_tissue.414.melt$DE),]
DEGs_number_between_tissue.414.melt

### making ggplot for DEGs 
p.DEGs_number_between_tissue.414 <- ggplot(data = DEGs_number_between_tissue.414.melt)
p.DEGs_number_between_tissue.414 <- p.DEGs_number_between_tissue.414 + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number), stat = "identity")
p.DEGs_number_between_tissue.414 <- p.DEGs_number_between_tissue.414 + facet_grid(~tissue)
p.DEGs_number_between_tissue.414 <- p.DEGs_number_between_tissue.414 + labs(y = "number of differentially expressed genes", x = "")

p.DEGs_number_between_tissue.414
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/DEGs_number_between_tissue.414.png", width = 11, height = 8)  

### 415 
lrt.young.vs.flowering.415 <- glmLRT(fit.F1, contrast = c(0,0,0,0,0,1,0,-1))
topTags(lrt.young.vs.flowering.415)
summary(de.young.vs.flowering.415 <- decideTestsDGE(lrt.young.vs.flowering.415, p=0.05))
DEgene.young.vs.flowering.415 <- topTags(lrt.young.vs.flowering.415,n = Inf)$table[topTags(lrt.young.vs.flowering.415,n = Inf)$table$FDR<0.05,]
# -1 12450
# 0  32614
# 1  15112

lrt.flowering.vs.early.silique.415 <- glmLRT(fit.F1, contrast = c(0,0,0,0,1,-1,0,0))
topTags(lrt.flowering.vs.early.silique.415)
summary(de.flowering.vs.early.silique.415 <- decideTestsDGE(lrt.flowering.vs.early.silique.415, p=0.05))
DEgene.flowering.vs.early.silique.415 <- topTags(lrt.flowering.vs.early.silique.415,n = Inf)$table[topTags(lrt.flowering.vs.early.silique.415,n = Inf)$table$FDR<0.05,]
# -1  7666
# 0  47946
# 1   4564

lrt.early.vs.late.silique.415 <- glmLRT(fit.F1, contrast = c(0,0,0,0,-1,0,1,0))
topTags(lrt.early.vs.late.silique.415)
summary(de.early.vs.late.silique.415 <- decideTestsDGE(lrt.early.vs.late.silique.415, p=0.05))
DEgene.early.vs.late.silique.415 <- topTags(lrt.early.vs.late.silique.415,n = Inf)$table[topTags(lrt.early.vs.late.silique.415,n = Inf)$table$FDR<0.05,]
# -1  3074
# 0  53197
# 1   3905

### ggplot 
### reformat for ggplot barplot 
DEGs_number_between_tissue.415 <- data.frame("flowering-vs-young" = as.data.frame(summary(de.young.vs.flowering.415 <- decideTestsDGE(lrt.young.vs.flowering.415, p=0.05)))$Freq, 
                                     "early.slique-vs-flowering"= as.data.frame(summary(de.flowering.vs.early.silique.415 <- decideTestsDGE(lrt.flowering.vs.early.silique.415, p=0.05)))$Freq, 
                                     "late.silique-vs-early.silique" = as.data.frame(summary(de.early.vs.late.silique.415 <- decideTestsDGE(lrt.early.vs.late.silique.415, p=0.05)))$Freq)

rownames(DEGs_number_between_tissue.415) <- c("down", "no", "up")
DEGs_number_between_tissue.415 <- DEGs_number_between_tissue.415[c("down", "up"),]
DEGs_number_between_tissue.415

DEGs_number_between_tissue.415.melt <- melt(DEGs_number_between_tissue.415)
DEGs_number_between_tissue.415.melt$DE <- rep(c("down", "up"), 3)
colnames(DEGs_number_between_tissue.415.melt) <- c("tissue", "number", "DE")
DEGs_number_between_tissue.415.melt

# reorder: up 1st down 2nd 
DEGs_number_between_tissue.415.melt$DE <- factor(DEGs_number_between_tissue.415.melt$DE, levels = c("up", "down"))

DEGs_number_between_tissue.415.melt <- DEGs_number_between_tissue.415.melt[order(DEGs_number_between_tissue.415.melt$DE),]
DEGs_number_between_tissue.415.melt

### making ggplot for DEGs 
p.DEGs_number_between_tissue.415 <- ggplot(data = DEGs_number_between_tissue.415.melt)
p.DEGs_number_between_tissue.415 <- p.DEGs_number_between_tissue.415 + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number), stat = "identity")
p.DEGs_number_between_tissue.415 <- p.DEGs_number_between_tissue.415 + facet_grid(~tissue)
p.DEGs_number_between_tissue.415 <- p.DEGs_number_between_tissue.415 + labs(y = "number of differentially expressed genes", x = "")

p.DEGs_number_between_tissue.415
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/DEGs_number_between_tissue.415.png", width = 11, height = 8)  
### 

```

# venn diagram between different stages of 414 & 415  
```{r}
# calcuate overlaps & unique genes 
dim(DEgene.young.vs.flowering.414) # 29600     5 
dim(DEgene.young.vs.flowering.415) # 27562     5
summary(rownames(DEgene.young.vs.flowering.414) %in% rownames(DEgene.young.vs.flowering.415)) 
#    Mode   FALSE    TRUE    NA's 
# logical    8561   21039       0

require(VennDiagram)

# young vs flowering
grid.newpage()
png(filename = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/young.vs.bolting.png")
venn.plot <- draw.pairwise.venn(29600, 27562, 21039, # 1st number for group A, 2nd for group B, 3rd for overlaps 
                                c("414", "415"),
                                fill = c("lightblue","pink"),
                                lty          = "blank",
                                cex = 2,
                                cat.cex = 1.5,
                                cat.pos = 9)
grid.draw(venn.plot)
dev.off()

# flowering vs early silique 
dim(DEgene.flowering.vs.early.silique.414) # 15839     5 
dim(DEgene.flowering.vs.early.silique.415) # 12230     5
summary(rownames(DEgene.flowering.vs.early.silique.414) %in% rownames(DEgene.flowering.vs.early.silique.415)) 
#    Mode   FALSE    TRUE    NA's 
# logical    6584    9255       0 

# young vs bolting
grid.newpage()
png(filename = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/flowering.vs.early.silique.png")
venn.plot <- draw.pairwise.venn(15839, 12230, 9255, # 1st number for group A, 2nd for group B, 3rd for overlaps 
                                c("414", "415"),
                                fill = c("lightblue","pink"),
                                lty          = "blank",
                                cex = 2,
                                cat.cex = 1.5,
                                cat.pos = 9)
grid.draw(venn.plot)
dev.off()

# early silique vs late silique 
dim(DEgene.early.vs.late.silique.414) # 6373    5 
dim(DEgene.early.vs.late.silique.415) # 6979    5
summary(rownames(DEgene.early.vs.late.silique.414) %in% rownames(DEgene.early.vs.late.silique.415)) 
#    Mode   FALSE    TRUE    NA's 
# logical    2652    3721       0

# young vs bolting
grid.newpage()
png(filename = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/early.vs.late.silique.png")
venn.plot <- draw.pairwise.venn(6373, 6979, 3721, # 1st number for group A, 2nd for group B, 3rd for overlaps 
                                c("414", "415"),
                                fill = c("lightblue","pink"),
                                lty          = "blank",
                                cex = 2,
                                cat.cex = 1.5,
                                cat.pos = 9)
grid.draw(venn.plot)
dev.off()
```

# GO enrichment analysis 
```{r}
# GO enrichment analysis 
# load all necessary functions 
source("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/function_BnRNAseq.R")

# young 
DEgene.young.GO.ORA.F1 <- GOseq.Bn.ORA(rownames(DEgene.young.F1))
DEgene.young.GO.ORA.F1
write.csv(DEgene.young.GO.ORA.F1, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/DEgene.young.GO.ORA.F1.csv")

# flowering 
DEgene.flowering.GO.ORA.F1 <- GOseq.Bn.ORA(rownames(DEgene.flowering.F1))
DEgene.flowering.GO.ORA.F1
write.csv(DEgene.flowering.GO.ORA.F1, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/DEgene.flowering.GO.ORA.F1.csv")

# early silique 
DEgene.early.silique.GO.ORA.F1 <- GOseq.Bn.ORA(rownames(DEgene.early.silique.F1))
DEgene.early.silique.GO.ORA.F1
write.csv(DEgene.early.silique.GO.ORA.F1, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/DEgene.early.silique.GO.ORA.F1.csv")

# late silique 
DEgene.late.silique.GO.ORA.F1 <- GOseq.Bn.ORA(rownames(DEgene.late.silique.F1))
DEgene.late.silique.GO.ORA.F1
write.csv(DEgene.late.silique.GO.ORA.F1, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/DEgene.late.silique.GO.ORA.F1.csv")

### between stages 
### 414 
# young vs flowering 
DEgene.young.vs.flowering.414.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.young.vs.flowering.414))
DEgene.young.vs.flowering.414.GO.ORA
nrow(DEgene.young.vs.flowering.414.GO.ORA)
write.csv(DEgene.young.vs.flowering.414.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/DEgene.young.vs.flowering.414.GO.ORA.csv")

# flowering vs early silique 
DEgene.flowering.vs.early.silique.414.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.flowering.vs.early.silique.414))
DEgene.flowering.vs.early.silique.414.GO.ORA
nrow(DEgene.flowering.vs.early.silique.414.GO.ORA)
write.csv(DEgene.flowering.vs.early.silique.414.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/DEgene.flowering.vs.early.silique.GO.ORA.csv")

# early vs late silique 
DEgene.silique.414.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.early.vs.late.silique.414))
DEgene.silique.414.GO.ORA
nrow(DEgene.silique.414.GO.ORA)
write.csv(DEgene.silique.414.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/DEgene.silique.414.GO.ORA.csv")

### 415 
# young vs flowering 
DEgene.young.vs.flowering.415.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.young.vs.flowering.415))
DEgene.young.vs.flowering.415.GO.ORA
write.csv(DEgene.young.vs.flowering.415.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/DEgene.young.vs.flowering.415.GO.ORA.csv")

# flowering vs early silique 
DEgene.flowering.vs.early.silique.415.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.flowering.vs.early.silique.415))
DEgene.flowering.vs.early.silique.415.GO.ORA
write.csv(DEgene.flowering.vs.early.silique.415.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/DEgene.flowering.vs.early.silique.415.GO.ORA.csv")

# early vs late silique 
DEgene.silique.415.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.early.vs.late.silique.415))
DEgene.silique.415.GO.ORA
write.csv(DEgene.silique.415.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/F1/DEgene.silique.415.GO.ORA.csv")

```

# draw figure for fatty acid genes expression over time for 414 & 415 
```{r}
source("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/function_BnRNAseq.R")

fatty.acid.genes.annotation <-  read.csv("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/fatty_acid_gene_name.csv", header = F)
rownames(fatty.acid.genes.annotation) <- fatty.acid.genes.annotation$V1
fatty.acid.genes.interaction.F1.ID <- rownames(fatty.acid.genes.annotation) 

expression.pattern.Bn.F1.with.annot(fatty.acid.genes.interaction.F1.ID, fatty.acid.genes.annotation)

ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/fatty.acid.interaction.F1.png", width = 8, height = 20)

three.fatty.acid.genes <- c("BnaA05g06830D","BnaA10g09300D",  "BnaC09g50630D") 
expression.pattern.Bn.F1.with.annot(three.fatty.acid.genes, fatty.acid.genes.annotation)
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/fatty.acid.three.F1.png", width = 11, height = 8) 
```

### other important genes 
# Acetyl-CoA-ACP-transacetylase: GO:0004313 none 
# B-Ketoacyl-ACP-synthase: IPR014030 --> A 
# Malonyl-CoA-ACP-transacetylase: IPR016036 --> B  
# B-ketoacyl-ACP-reductase: GO:0004316 --> C 
# Enoyl-ACP-hydrase: IPR018376 --> D
# Enoyl-ACP-reductase: GO:0016631 & GO:0004319 none 
# Palmitoyl-thioesterase: IPR002472 --> E 
```{r}
# B-Ketoacyl-ACP-synthase 
A.ID <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/B-Ketoacyl-ACP-synthase", header = F)
p.A <- expression.pattern.Bn.F1(A.ID)
p.A <- p + labs(y = "mean expression value", x="tissue", title="B-Ketoacyl-ACP-synthase")
p.A 
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/5f.png", width = 8, height = 20)

# Malonyl-CoA-ACP-transacetylase  
B.ID <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/Malonyl-CoA-ACP-transacetylase", header = F)
dim(B.ID)
p.B <- expression.pattern.Bn.F1(B.ID)
p.B <- p.B + labs(y = "mean expression value", x="tissue", title="Malonyl-CoA-ACP-transacetylase")
p.B
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/5g.png", width = 8, height = 10)

# B-ketoacyl-ACP-reductase: GO:0004316 --> C 
C.ID <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/B-ketoacyl-ACP-reductase", header = F)
dim(C.ID)
p.C <- expression.pattern.Bn.F1(C.ID)
p.C <- p.C + labs(y = "mean expression value", x="tissue", title="B-ketoacyl-ACP-reductase") 
p.C
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/5h.png", width = 8, height = 8)

# Enoyl-ACP-hydrase: IPR018376 --> D
D.ID <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/Enoyl-ACP-hydrase", header = F)
dim(D.ID)
p.D <- expression.pattern.Bn.F1(D.ID)
p.D <- p.D + labs(y = "mean expression value", x="tissue", title="Enoyl-ACP-hydrase") 
p.D
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/5i.png", width = 8, height = 20)

# Palmitoyl-thioesterase: IPR002472 --> E 
E.ID <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/Palmitoyl-thioesterase", header = F)
dim(E.ID)
p.E <- expression.pattern.Bn.F1(E.ID)
p.E <- p.E + labs(y = "mean expression value", x="tissue", title="Palmitoyl-thioesterase") 
p.E
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/F1/5j.png", width = 8, height = 20) 
``` 






















