---
title: "KIAT_RNA_seq"
author: "Ruijuan Li"
date: "August 1, 2016"
output: html_document
---

# download data, QC, and mapping (using kallisto)
```{r}
# 1) download file from ftp using ftp_file_transfer.sh # ftp works only under whitney. 
# 1.1) unzip files using gunzip *  
# 2) get first 1000 sequences for each file extract_1000_reads.sh, actually worked with all data instead of the 1st 1000 seqs
# 3) QC file using fastQC: fastqc *.fq -o /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/2016_summer/fastqc_result
# 4) download Brassica_napus.annotation_v5.cds.fa to whitney 
# 5) collapse fastqc summary result into one file & get overrepresented sequences using ./make_summary_report.sh 

# 6) reformat summary.all.txt file 
setwd("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/")
fastQC_summary <- read.delim("./fastQC_result/summary_all.txt", header = F)
dim(fastQC_summary)
head(fastQC_summary)
tail(fastQC_summary)

fastQC_summary_wide <- reshape(fastQC_summary, timevar="V2",idvar="V3",direction="wide")
colnames(fastQC_summary_wide) <- gsub("(V1)(.)([[:print:]]+)","\\3",colnames(fastQC_summary_wide))
dim(fastQC_summary_wide)

write.csv(fastQC_summary_wide, file = "fastQC_summary_wide.csv")

# 7) checked the source of overrepresented sequences using web nucleotide blast tool, found that most of them are from 
# cholorplast, mitochondra, rRNA, tansposon, etc. Some are illumina or TruSeq adapters 

# blast sequences against Trimmomatic adapters suggest that TruSeq3-PE-2.fa should be used to trimm the data. (Q: how to deal with the others? will they affect the final result? search... )

# Qs on plstids seqs contamination: 1) why are there these sequences? check the lib making step 2) will they affect downstream analysis if kept? CDS mapping VS genome mapping 3) If yes, how to remove them? (I need to figure these Qs out.)

# 8) using kallisto to build the reference cds index 
# kallisto index -k 19 -i Brassica_napus.annotation_v5.cds.19.kai /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5.cds.fa

# 9) map to reference cds index and get read count file using bunchrun_kallisto.sh (using est_counts from Kallisto)
```

# using STAR 
```{r}
# 1) get B.napus reference genome: 

# add "gene_id" colomn to gff3 file 
# run with modified gff3 file 
# STAR --runMode genomeGenerate --genomeDir star_genome/ --genomeFastaFiles Brassica_napus_v4.1.chromosomes.fa --sjdbGTFfile Brassica_napus.annotation_v5_modified_modified.gff3 --runThreadN 6 --sjdbGTFtagExonParentTranscript Parent --sjdbGTFfeatureExon CDS (screen -r 10070.ttys003.coloma) 

# 2) mapping 
# STAR --genomeDir /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/star_genome --readFilesIn /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/2016_summer/raw_data/$i"_1.fq" /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/2016_summer/raw_data/$i"_2.fq" --outSAMtype BAM SortedByCoordinate --sjdbGTFfile /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5_modified_modified.gff3 --quantMode TranscriptomeSAM GeneCounts --twopassMode Basic –alignIntronMax 15000 --outFilterIntronMotifs RemoveNoncanonical --runThreadN 6 --sjdbGTFtagExonParentTranscript Parent --sjdbGTFfeatureExon CDS --outReadsUnmapped fasta (done) 

### because of the naming problems of All1_Yeong3_160621 & All1_Yeong2_160621, two seperate STAR mapping were made for them using the below codes: 
# STAR --genomeDir /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/star_genome --readFilesIn ../All1_Yeong2_160621_1.fq ../All1_Yeong2_160621_2.fq --outSAMtype BAM SortedByCoordinate --sjdbGTFfile /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5_modified_modified.gff3 --quantMode TranscriptomeSAM GeneCounts --twopassMode Basic –alignIntronMax 15000 --outFilterIntronMotifs RemoveNoncanonical --runThreadN 6 --sjdbGTFtagExonParentTranscript Parent --sjdbGTFfeatureExon CDS 
### very high percentage of unmapped!!!!!! 

# All1_Yeong3_160621 sample  
# STAR --genomeDir /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/star_genome --readFilesIn /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/2016_summer/raw_data/All1_Yeong3_160621_1.fq /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/2016_summer/raw_data/All1_Yeong3_160621_2.fq --outSAMtype BAM SortedByCoordinate --sjdbGTFfile /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5_modified_modified.gff3 --quantMode TranscriptomeSAM GeneCounts --twopassMode Basic –alignIntronMax 15000 --outFilterIntronMotifs RemoveNoncanonical --runThreadN 6 --sjdbGTFtagExonParentTranscript Parent --sjdbGTFfeatureExon CDS --outReadsUnmapped fastx 

# get unmapped reads with command "STAR --genomeDir /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/star_genome --readFilesIn /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/2016_summer/raw_data/Ae_Hu_2_1.fq /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/2016_summer/raw_data/Ae_Hu_2_2.fq --outSAMtype BAM SortedByCoordinate --sjdbGTFfile /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5_modified_modified.gff3 --quantMode TranscriptomeSAM GeneCounts --twopassMode Basic –alignIntronMax 15000 --outFilterIntronMotifs RemoveNoncanonical --runThreadN 6 --sjdbGTFtagExonParentTranscript Parent --sjdbGTFfeatureExon CDS --outReadsUnmapped Fastx" 
```

### find out why low mapping % 
```{r}
# 1) because of rRNA contamination??? 
# hypothesis: too high percentage of unmapped because of rRNA contamination & rRNA is absent in genome seq 
# 1) keep unmapped reads from STAR mapping by using command: STAR --genomeDir /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/star_genome --readFilesIn *_1.fq *_2.fq --outSAMtype BAM SortedByCoordinate --sjdbGTFfile /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5_modified_modified.gff3 --quantMode TranscriptomeSAM GeneCounts --twopassMode Basic –alignIntronMax 15000 --outFilterIntronMotifs RemoveNoncanonical --runThreadN 6 --sjdbGTFtagExonParentTranscript Parent --sjdbGTFfeatureExon CDS --outReadsUnmapped Fastx
# 2) download rRNA referece seq from SILVA 

# SILVA_128_SSURef_Nr99_tax_silva.fasta 
# SILVA_128_LSURef_tax_silva.fasta
# NCBI_rRNA_seq.fa (NCBI viradeaplantea rRNA)

# 3) compare B.napus genome & rRNA by blast ...  
# blast "blastn -db /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus_v4.1.chromosomes.fa -query NCBI_rRNA_seq.fa -out NCBIrRNA_napa_genome.blast.out -evalue 1e-6 -outfmt 6" found that out of the 520 rRNA seqs, 421 of them have significant hit in napus genone. so this is not the problem. 

# 2) because of genes from other organism??? 
# 2.1) mapping to combined B.rapa & B.oleracea genome to see the mapping result 
# 2.1.1) B.rapa used v1.5 genome version; there are two genome asesmblies for B.oleracea: one variety is cabbage, one is TO1000. While TO1000 is the representive one, I decide to make three combinations: a) B.rapa & cabbage b) B.rapa & TO1000 c) all three together. 
# About the genome data, there are two places I can get genome data fro B.oleracea: NCBI & http://brassicadb.org/brad/downloadOverview.php.After reading genome papers & NCBI inforamtion, I decide to use genome information from NCBI. 

# So for cabbage, its genome data is GCA_000604025.1_BOL_v1.0_genomic.fna (scaffold level assembly)
# for TO1000, there are two versions, GCF_000695525.1_BOL_genomic.fna & GCA_000695525.1_BOL_genomic.fna, I used GCF_000695525.1_BOL_genomic.fna. 

# I) build the genome index (rapa & TO1000)
# STAR --runMode genomeGenerate --genomeDir TO1000.rapa.star.genome/ --genomeFastaFiles TO10000.rapa.fa --sjdbGTFfile TO10000.rapa.gff --runThreadN 8 --sjdbGTFtagExonParentTranscript Parent --sjdbGTFfeatureExon CDS --limitGenomeGenerateRAM 53099915306   
# II) map to genome 
# no improvement on mapping rate 

### John found BWA has very high mapping rate, so it is because we didn't use STAR right? working to get the right STAR parameters.... but since TopHat also gives John low mapping rate, now John is trying to figure out why BWA gives such a high mapping rate.   

### include mt, chlroplast genome and  build the genome again 
# download napus Mt, napus chloroplast and rRNA from all plants, added these seqs into reference genome, rebuilt the genome using the same gff3 file 
# STAR --runMode genomeGenerate --genomeDir star_genome_chl_mt_rRNAall/ --genomeFastaFiles Brassica_napus_v4.1.chromosomes.chl.mt.rRNAall.fa --sjdbGTFfile Brassica_napus.annotation_v5_modified_modified.gff3 --runThreadN 6 --sjdbGTFtagExonParentTranscript Parent --sjdbGTFfeatureExon CDS

# map to the new ref genome to see whether I can get higher mapping rate 
# STAR --genomeDir /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/star_genome_chl_mt_rRNAall/ --readFilesIn ../1_1.fq ../1_2.fq --outSAMtype BAM SortedByCoordinate --sjdbGTFfile /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5_modified_modified.gff3 --quantMode TranscriptomeSAM GeneCounts --twopassMode Basic –alignIntronMax 15000 --outFilterIntronMotifs RemoveNoncanonical --runThreadN 6 --sjdbGTFtagExonParentTranscript Parent --sjdbGTFfeatureExon CDS --outReadsUnmapped Fastx 

# result, no mapping rate improvement, why? because the chloroplast & Mt are not the ones that I should use? use different ones? I am confused... 

#### map unmapped reads to Mt, chloroplast, and rRNA to see the mapping rate (using STAR, BWA)
# bwa index -p mitochondria_chloroplast_rRNA_all_bwa ../mitochondria_chloroplast_rRNA_all.fa 
# bwa mem -t 8 /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/non_coding/bwa_index/mitochondria_chloroplast_rRNA_all_bwa Unmapped.out.mate1 > unmapped_mate1_mt_chl_rRNA_bwa.se.sam 
# about half is mapped from BWA 

# Now I trim another sample to see whether I get high mapping rate for this sample trimmed with overrepresented seqs. 
# yes, with paired end mapping, 20% unampped, single end get 3% unmapped. 

# map to Arabidopsis genome 

# sequences trimmed with custom adapters map to napus genome 

```

# assemble unmapped reads to see what they hit 
```{r}
# 1) get unmapped reads by running STAR again for Da-Ol-1 young tissue 
# 2) Trinity to assemble unmapped reads
# 2.1) decide which library type my unmapped read1 is after visulizaing it in IGV, determined the library type for unmappedmate1 is F, unmappedmate2 is R then. paired end is FR  

# Trinity --seqType fq --single 1.star.dir.no.annotation.mapping/Unmapped.out.mate2.replaceID,All1_Yeong2_160621.star.dir.w.unmapped/Unmapped.out.mate2.replaceID,All1_Yeong3_160621.star.dir.w.unmapped/Unmapped.out.mate2.replaceID --SS_lib_type R --max_memory 50G --CPU 4 --output trinity_out_unmapped_mate2 

# next, blast trinity assembled transcripts to NCBI see homology 
# blastn -db /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/NCBI/nt/nt -query Trinity.fasta -out Trinity.nt.blast.out -evalue 1e-6 -outfmt 0 -max_target_seqs 1 -num_threads 6 

# 3) how to get reads with multi mapping and too many mapping ??? (under John's help)
# /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/johndavis/Ruijuan/Multi_Mapped 

# 4) Trinity assemble paired end raw reads 
# Trinity --seqType fq --left 1_1.fq --right 1_2.fq --SS_lib_type FR --max_memory 50G --CPU 8 --output trinity_out_all

# also, check I get the right bolting data, by checking earlier email, realize that bolting might not be good data. 

# also Richard suggested 
# 1) download other SRA data to see whether using my pipeline those data get high mapping rate, if not, then my pipeline might be problematic. If yes, then data has problem 

# downloading data now... 

# 2) map reads to transcriptome to see mapping rate. If very low mapping rate to CDS, that means most of the unmapped are from non coding regions...  

```

# cufflinks 
```{r}
### talked with Julin: use IGV to visualize the sequence alignment to decide library type 
# then after visualize genome, bam file, and gff file, we decide our library type is fr-secondstrand 

########### running cufflinks:
# cufflinks -u -p 10 -o cufflink_output --max-bundle-frags 10000000 --library-type fr-secondstrand -g /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5_modified_modified.gff3 -b /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus_v4.1.chromosomes.fa Aligned.sortedByCoord.out.bam  

####### cufflinks takes forever to run, why? 
# by checking the progress of cufflinks, I found that it took 15 days to get one transcript.gtf file for one lib. and mostly the progress is stucked at the last 5%. Why? Also at the begnining of the command, it says "segementation fault, core dumped", but still makes progress. previously, I had cufflinks running on one core and only takes 3 days to finish 11 libs. And I was using exactly the same command. Why??? 

# test on one lib: 2_paired.star.trim.dir 
# change the core number to 6 because John is using 8, if I still ask for 10, it might cause problem. Also add 
# && 
# echo "success"
# to the end of the command to make sure I get it run correctly. now it is running for sample 2, waiting for the result, fingers crossed!!! 
# since cufflinks can only use one core during some step, the best practice should be make make it several cufflinks run in parallel, each using one core? 
# this still runs very slow.... 

### after talking to Richard group, we decide to stop Trinity and run cufflinks only on 14 cores to see whether this will speed up assembly. 
# library, now still runs very slow, now I decide not to use --max-bundle-frags option to see whether this can speed up the assembly. 



```

# SNP calling  
```{r}
# index reference genome fasta file 
# samtools faidx Brapa_sequence_v1.5.fa  

# 1) merge libarary for Da-Ae and Da-Ol-1 
# cat Ae*1.fq 5_1.fq 6_1.fq 05-11-Final-8ul_1.fq 8_1.fq > Ae_1.fq
# cat Ae*2.fq 5_2.fq 6_2.fq 05-11-Final-8ul_2.fq 8_2.fq > Ae_2.fq

# cat All*1.fq 1_1.fq 2_1.fq 3_1.fq 4_1.fq > Ol_1.fq
# cat All*2.fq 1_2.fq 2_2.fq 3_2.fq 4_2.fq > Ol_2.fq

# trim Da-Ae & Da-Ol sample using: java -jar /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/bin/Trimmomatic-0.36/trimmomatic-0.36.jar PE ${i}_1.fq ${i}_2.fq ${i}_paired_1.fq ${i}_unpaired_1.fq ${i}_paired_2.fq ${i}_unpaired_2.fq ILLUMINACLIP:Bradseq_adapter.fa:2:30:10 LEADING:3 TRAILING:3 SLIDINGWINDOW:4:15 MINLEN:36 -threads 2 

# 2) do STAR mapping for Da-Ae & Da-Ol-1 paired-end seq 

### STAR mapping of merged data 
# STAR --genomeDir /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/star_genome --readFilesIn *_1.fq *_2.fq --outSAMtype BAM SortedByCoordinate --sjdbGTFfile /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5_modified_modified.gff3 --quantMode TranscriptomeSAM GeneCounts --twopassMode Basic –alignIntronMax 15000 --outFilterIntronMotifs RemoveNoncanonical --runThreadN 2 --sjdbGTFtagExonParentTranscript Parent --sjdbGTFfeatureExon CDS --outReadsUnmapped Fastx 

# (visualizing the MDS plot and hierachical plot, I found bolting are not mixed between genotypes, its that they are not clustering with themselve, which might because the two genotypes have different flowering time? Da_Ae's bolting is more similar to young whereas Da-Ol's bolting is more similar to flowering.)  

### remove PCR duplicate 
# samtools rmdup -s Aligned.sortedByCoord.out.bam Ae_rmdup.bam
# samtools rmdup -s Aligned.sortedByCoord.out.bam Ol_rmdup.bam 

### index bam file 
# samtools index Ae_rmdup.bam
# samtools index Ol_rmdup.bam

### add read group & SNP calling ... 
# bamaddrg -b Ae_paired.star.trim.dir/Ae_rmdup.bam -s Ae -r Ae -b Ol_paired.star.trim.dir/Ol_rmdup.bam -s Ol -r Ol | freebayes_v0.9.21-7-g7dd41db --fasta-reference /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus_v4.1.chromosomes.fa --stdin --ploidy 4 --min-alternate-count 4 --min-alternate-fraction 0.4 --dont-left-align-indels > Ae_Ol.vcf 

############### 
################### 
# opps, forget that bolting data should not be included for SNP calling !!! 
# mapping w/o bolting data 
# 1) combine all trimmed reads for Da-Ae and Da-Ol-1 paired end reads 
# cat Ae_Yeong_2_paired_1.fq Ae_Yeong_3_paired_1.fq Ae_Gae_2_paired_1.fq Ae_Gae_3_paired_1.fq Ae_Cho_2_paired_1.fq Ae_Cho_3_paired_1.fq Ae_Hu_2_paired_1.fq Ae_Hu_3_paired_1.fq 5_paired_1.fq 6_paired_1.fq 05-11-Final-8ul_paired_1.fq 8_paired_1.fq > Ae_no_bolting_1.fq

# cat Ae_Yeong_2_paired_2.fq Ae_Yeong_3_paired_2.fq Ae_Gae_2_paired_2.fq Ae_Gae_3_paired_2.fq Ae_Cho_2_paired_2.fq Ae_Cho_3_paired_2.fq Ae_Hu_2_paired_2.fq Ae_Hu_3_paired_2.fq 5_paired_2.fq 6_paired_2.fq 05-11-Final-8ul_paired_2.fq 8_paired_2.fq > Ae_no_bolting_2.fq 

# cat All1_Yeong2_160621_paired_1.fq All1_Yeong3_160621_paired_1.fq All1_Gae_2_paired_1.fq All1_Gae_3_paired_1.fq All1_Cho_2_paired_1.fq All1_Cho_3_paired_1.fq All1_Hu_2_paired_1.fq All1_Hu_3_paired_1.fq 1_paired_1.fq 2_paired_1.fq 3_paired_1.fq 4_paired_1.fq > Ol_no_bolting_1.fq 

# cat All1_Yeong2_160621_paired_2.fq All1_Yeong3_160621_paired_2.fq All1_Gae_2_paired_2.fq All1_Gae_3_paired_2.fq All1_Cho_2_paired_2.fq All1_Cho_3_paired_2.fq All1_Hu_2_paired_2.fq All1_Hu_3_paired_2.fq 1_paired_2.fq 2_paired_2.fq 3_paired_2.fq 4_paired_2.fq > Ol_no_bolting_2.fq

## STAR mapping 
## remove PCR duplicate 
# samtools rmdup -s Aligned.sortedByCoord.out.bam Ae_no_bolting_rmdup.bam 
# samtools rmdup -s Aligned.sortedByCoord.out.bam Ol_no_bolting_rmdup.bam 

### index bam file 
# samtools index Ae_no_bolting_rmdup.bam 
# samtools index Ol_no_bolting_rmdup.bam 

### add read group & SNP calling ... 
# bamaddrg -b Ae_paired.star.trim.dir/Ae_no_bolting_rmdup.bam -s Ae -r Ae -b Ol_paired.star.trim.dir/Ol_no_bolting_rmdup.bam -s Ol -r Ol | freebayes_v0.9.21-7-g7dd41db --fasta-reference /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus_v4.1.chromosomes.fa --stdin --ploidy 4 --min-alternate-count 4 --min-alternate-fraction 0.4 --dont-left-align-indels > Ae_Ol_no_bolting.vcf (screen -r 35910.ttys009.coloma)  
```

# Expression analysis 1) formatting data   
```{r}
parent.read.count <- read.table("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/read_count/read.count.tsv", header = T, check.names = F)
rownames(parent.read.count) <- parent.read.count[,1]
parent.read.count <- parent.read.count[,-1]

# format data 
head(parent.read.count)
dim(parent.read.count) # 101040     54 
colnames(parent.read.count)
even_indexes<-seq(2,length(colnames(parent.read.count)),2)
parent.read.count.one <- parent.read.count[,even_indexes]
dim(parent.read.count.one) # 101040     27 
colnames(parent.read.count.one)
colnames(parent.read.count.one) <- sub("_2.fq","",colnames(parent.read.count.one),fixed = TRUE) # remove _2.fq 

# sample description  
sample_des <- read.csv("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/parent_summary_corrected.csv")
dim(sample_des) # 27 22
sorted_sample_des <- sample_des[order(sample_des$SampleID),]

sorted_sample_des[,4:7]
new_sample_ID <- paste(sorted_sample_des$Cultivar, sorted_sample_des$Stage, sorted_sample_des$rep, sep = "_")
new_sample_ID

# replace sample ID 
colnames(parent.read.count.one) <- new_sample_ID
head(parent.read.count.one)
# save(parent.read.count.one,file="/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.one.Rdata")
```

# data summary statisitics read count number & expressed genes 
```{r}
parent.read.count.normalized <- read.table("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/read.count.normalized.tsv", header = T, check.names = F)
rownames(parent.read.count.normalized) <- parent.read.count.normalized[,1]
parent.read.count.normalized <- parent.read.count.normalized[,-1]

# format data 
head(parent.read.count.normalized)
dim(parent.read.count.normalized) # 101040     54 
colnames(parent.read.count.normalized)
even_indexes<-seq(2,length(colnames(parent.read.count.normalized)),2)
parent.read.count.one.normalized <- parent.read.count.normalized[,even_indexes]
dim(parent.read.count.one.normalized) # 101040     27 
colnames(parent.read.count.one.normalized)
colnames(parent.read.count.one.normalized) <- sub("_2.fq","",colnames(parent.read.count.one.normalized),fixed = TRUE) # remove _2.fq 

# sample description  
new_sample_ID

# replace sample ID 
colnames(parent.read.count.one.normalized) <- new_sample_ID
head(parent.read.count.one.normalized)
save(parent.read.count.one.normalized,file="/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.one.normalized.Rdata")

# check number of genes with reads number > 10 in each lib # make summary table 
expressed_gene_number <- 
sapply(colnames(parent.read.count.one.normalized), function(RNAlib){
  sum(parent.read.count.one.normalized[,RNAlib]>3)
}
)
expressed_gene_number <- as.data.frame(expressed_gene_number)
expressed_gene_number

# top 10 highly expressed genes in each lib 
most_highly_expressed_genes <- 
sapply(colnames((parent.read.count.one.normalized)), function(RNAlib){
  rownames(head(parent.read.count.one.normalized[order(parent.read.count.one.normalized[,RNAlib],decreasing = TRUE),], 10))
}
)
# write.csv(t(most_highly_expressed_genes), file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/most_highly_expressed_genes.csv")

temp <- sample_des[,c(4,5,6,8)]
rownames(temp) <- paste(sample_des[,4], sample_des[,5], sample_des[,6], sep = "_")
sample_statistics <- merge(temp, expressed_gene_number, by=0) 
colnames(sample_statistics) <- c("sample_ID", "Cultivar", "Tissue", "rep", "total_reads", "expressed_gene_number")
head(sample_statistics)
colnames(sample_statistics) 
# write.csv(sample_statistics, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/sample_statisitcs.csv")

```

# expression analysis 2) set up sample description 
```{r}
# load necessary libs & functions 
library(edgeR)
library(ggplot2)

# filter based on read count, assign group, normalize, design matrix, calculate dispersion   
# set up group 
load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.one.Rdata")
parent.read.count.one <- parent.read.count.one[,colSums(parent.read.count.one) > 1000000]  
dim(parent.read.count.one) # 101040     27 
parent.read.count.one.sample<-data.frame(file=colnames(parent.read.count.one),
                             batch=factor(gsub("(Da-Ae|Da-Ol-1)(_)(Young|flowering|early-silique|late-silique|bolting)(_)(1|2|3)","\\5",colnames(parent.read.count.one))),  
                             genotype=factor(gsub("(Da-Ae|Da-Ol-1)(_)(Young|flowering|early-silique|late-silique|bolting)(_)(1|2|3)","\\1",colnames(parent.read.count.one))),	
                             stage=factor(gsub("(Da-Ae|Da-Ol-1)(_)(Young|flowering|early-silique|late-silique|bolting)(_)(1|2|3)","\\3",colnames(parent.read.count.one))),	
                             group=factor(gsub("(Da-Ae|Da-Ol-1)(_)([[:print:]]+)(_)(1|2|3)","\\1\\3",colnames(parent.read.count.one)))
)

parent.read.count.one.sample
ftable(parent.read.count.one.sample,row.vars="stage",col.vars=c("batch","genotype"))

# filter based on read count 
parent.read.count.one.small <- parent.read.count.one[rowSums(parent.read.count.one > 10) >= 3,]
dim(parent.read.count.one.small) # 60975    27 
# save(parent.read.count.one.small, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.one.small.Rdata")
```

# expression analysis 3) voom transformation 
```{r}
# voom transformation 
# source("https://bioconductor.org/biocLite.R")
# biocLite("DESeq2")
load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.one.small.Rdata")
library("DESeq2")

dds.parent <- DESeqDataSetFromMatrix(countData = round(parent.read.count.one.small), colData = parent.read.count.one.sample, design = ~ batch + genotype*stage)

vsd.parent <- varianceStabilizingTransformation(dds.parent)
vstMat.parent <- assay(vsd.parent)
colnames(vstMat.parent) <- colnames(parent.read.count.one)
# save(vstMat.parent, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/vstMat.parent.Rdata")
```

# expression analysis 4) MDS, clustering, and expression analysis 
```{r}
load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/vstMat.parent.Rdata")
load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.one.small.Rdata")

# normalize 
dge.data.parent <- DGEList(counts=parent.read.count.one.small, group=parent.read.count.one.sample$group)
dge.data.parent <- calcNormFactors(dge.data.parent, method = "TMM") 
dge.data.parent$sample
hist(dge.data.parent$samples$norm.factors)

# MDS to check sample seperation, also using dengrogram.  
mds <- plotMDS(dge.data.parent, method = "bcv",labels = dge.data.parent$samples$group)

x <- as.data.frame(mds$x)
y <- as.data.frame(mds$y)
distance_matrix <- merge(x, y, by="row.names")
distance_matrix$group <- gsub("(Da-Ae|Da-Ol-1)(_)(Young|flowering|early-silique|late-silique|bolting)(_)(1|2|3)","\\1\\3",distance_matrix$Row.names)
distance_matrix$gt <- gsub("(Da-Ae|Da-Ol-1)(_)(Young|flowering|early-silique|late-silique|bolting)(_)(1|2|3)","\\1",distance_matrix$Row.names)
distance_matrix$tissue <- gsub("(Da-Ae|Da-Ol-1)(_)(Young|flowering|early-silique|late-silique|bolting)(_)(1|2|3)","\\3",distance_matrix$Row.names)

colnames(distance_matrix) <- c("lib","x","y","group","gt","tissue")
head(distance_matrix)

# making color MDS figure 
p.mds <- ggplot(data = distance_matrix) + theme_gray()
p.mds <- p.mds + geom_point(aes(x, y, color=factor(group)), size=3) 
p.mds <- p.mds + labs(y = "BCV distance 2", x="BCV distance 1")
# p.mds <- p.mds + facet_grid(~gt)
p.mds      

ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/cDNA_mapped/MDS.by.gt.png", width = 11, height = 8)

# clustering to check sample seperation 
library(ggdendro)

hc.parent <- hclust(dist(t(vstMat.parent)))

ggdata.parent <- dendro_data(as.dendrogram(hc.parent))
ggdata.parent$labels$gt <- gsub("(Da-Ae|Da-Ol-1)(_)(Young|flowering|early-silique|late-silique|bolting)(_)(1|2|3)","\\1",ggdata.parent$labels$label) 
ggdata.parent$labels$tissue <- gsub("(Da-Ae|Da-Ol-1)(_)(Young|flowering|early-silique|late-silique|bolting)(_)(1|2|3)","\\3",ggdata.parent$labels$label)
ggdata.parent$labels$group <- paste(ggdata.parent$labels$gt, ggdata.parent$labels$tissue, sep = "_")
ggdata.parent$labels

# start ggplot here 
p.dengro <- ggplot(data = segment(ggdata.parent))
p.dendro <- p.dengro + geom_segment(aes(x=x, y=y, xend=xend, yend=yend)) + theme_dendro()
p.dendro <- p.dendro + geom_text(data = label(ggdata.parent), aes(x = x, y = y, label = label, hjust=0, color=group)) + coord_flip() + scale_y_reverse(expand=c(0.2, 0))
p.dendro
# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/clustering_new.png", width = 15, height = 8) 

# pairwise comparison using GLM model 
# design matrix
# design.parent.batch <- model.matrix(~batch+group, data = parent.read.count.one.sample) 
design.parent <- model.matrix(~0+group, data = parent.read.count.one.sample)

# w/o batch effect
dge.data.parent <- estimateGLMCommonDisp(dge.data.parent, design.parent,verbose = TRUE) # Disp = 0.25572 , BCV = 0.5057
dge.data.parent <- estimateGLMTrendedDisp(dge.data.parent,design.parent)
dge.data.parent <- estimateGLMTagwiseDisp(dge.data.parent,design.parent)
plotBCV(dge.data.parent)

## w/o batch effect
fit.parent <- glmFit(dge.data.parent, design.parent)

lrt.young <- glmLRT(fit.parent, contrast = c(0,0,0,0,1,0,0,0,0,-1))
topTags(lrt.young)
summary(de.young <- decideTestsDGE(lrt.young, p=0.05))
DEgene.young <- topTags(lrt.young,n = Inf)$table[topTags(lrt.young,n = Inf)$table$FDR<0.05,]
# -1  4553
# 0  51999
# 1   4423

lrt.flowering <- glmLRT(fit.parent, contrast = c(0,0,1,0,0,0,0,-1,0,0))
topTags(lrt.flowering)
summary(de.flowering <- decideTestsDGE(lrt.flowering, p=0.05))
DEgene.flowering <- topTags(lrt.flowering,n = Inf)$table[topTags(lrt.flowering,n = Inf)$table$FDR<0.05,]
# -1  2877
# 0  54894
# 1   3204

lrt.early.silique <- glmLRT(fit.parent, contrast = c(0,1,0,0,0,0,-1,0,0,0))
topTags(lrt.early.silique)
summary(de.early.silique <- decideTestsDGE(lrt.early.silique, p=0.05))
DEgene.early.silique <- topTags(lrt.early.silique,n = Inf)$table[topTags(lrt.early.silique,n = Inf)$table$FDR<0.05,]
# -1  3200
# 0  54280
# 1   3495

lrt.late.silique <- glmLRT(fit.parent, contrast = c(0,0,0,1,0,0,0,0,-1,0))
topTags(lrt.late.silique)
summary(de.late.silique <- decideTestsDGE(lrt.late.silique, p=0.05))
DEgene.late.silique <- topTags(lrt.late.silique,n = Inf)$table[topTags(lrt.late.silique,n = Inf)$table$FDR<0.05,]
# -1  3745
# 0  53562
# 1   3668

lrt.bolting <- glmLRT(fit.parent, contrast = c(1,0,0,0,0,-1,0,0,0,0))
topTags(lrt.bolting)
summary(de.bolting <- decideTestsDGE(lrt.bolting, p=0.05))
DEgene.bolting <- topTags(lrt.bolting,n = Inf)$table[topTags(lrt.bolting,n = Inf)$table$FDR<0.05,]
# -1  2794
# 0  56327
# 1   1854

### barplot for DEGs number 
### reformat for ggplot barplot 
DEGs_number_between_gt.parent <- data.frame(young = as.data.frame(summary(de.young <- decideTestsDGE(lrt.young, p=0.05)))$Freq,
                                            bolting = as.data.frame(summary(de.bolting <- decideTestsDGE(lrt.bolting, p=0.05)))$Freq,
                                     flowering = as.data.frame(summary(de.flowering <- decideTestsDGE(lrt.flowering, p=0.05)))$Freq, 
                                     early.silique = as.data.frame(summary(de.early.silique <- decideTestsDGE(lrt.early.silique, p=0.05)))$Freq,
                                     late.silique = as.data.frame(summary(de.late.silique <- decideTestsDGE(lrt.late.silique, p=0.05)))$Freq)

rownames(DEGs_number_between_gt.parent) <- c("down", "no", "up")
DEGs_number_between_gt.parent <- DEGs_number_between_gt.parent[c("down", "up"),]
DEGs_number_between_gt.parent

library(reshape2)
DEGs_number_between_gt.melt.parent <- melt(DEGs_number_between_gt.parent)
DEGs_number_between_gt.melt.parent$DE <- rep(c("down", "up"), 5)
colnames(DEGs_number_between_gt.melt.parent) <- c("tissue", "number", "DE")
DEGs_number_between_gt.melt.parent

# reorder: up 1st down 2nd 
DEGs_number_between_gt.melt.parent$DE <- factor(DEGs_number_between_gt.melt.parent$DE, levels = c("up", "down"))

DEGs_number_between_gt.melt.parent <- DEGs_number_between_gt.melt.parent[order(DEGs_number_between_gt.melt.parent$DE),]
DEGs_number_between_gt.melt.parent

### making ggplot for DEGs 
p.DEGs_number_between_gt.parent <- ggplot(data = DEGs_number_between_gt.melt.parent) + theme_grey()
p.DEGs_number_between_gt.parent <- p.DEGs_number_between_gt.parent + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number), stat = "identity")
p.DEGs_number_between_gt.parent <- p.DEGs_number_between_gt.parent + facet_grid(~tissue)
p.DEGs_number_between_gt.parent <- p.DEGs_number_between_gt.parent + geom_text(data=DEGs_number_between_gt.melt.parent,aes(x=DE,y=number*1.05,label=factor(number),color=DE)) 
p.DEGs_number_between_gt.parent <- p.DEGs_number_between_gt.parent + labs(y = "number of differentially expressed genes", x = "")

p.DEGs_number_between_gt.parent
ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/cDNA_mapped/DEGs_number_between_gt.parent.png", width = 11, height = 8)

## between developmental stages 
# Da-Ae 
lrt.young.vs.bolting.ae <- glmLRT(fit.parent, contrast = c(1,0,0,0,-1,0,0,0,0,0))
topTags(lrt.young.vs.bolting.ae)
summary(de.young.vs.bolting.ae <- decideTestsDGE(lrt.young.vs.bolting.ae, p=0.05))
DEgene.young.vs.bolting.ae <- topTags(lrt.young.vs.bolting.ae,n = Inf)$table[topTags(lrt.young.vs.bolting.ae,n = Inf)$table$FDR<0.05,]
# -1   442
# 0  59872
# 1    661

lrt.bolting.vs.flowering.ae <- glmLRT(fit.parent, contrast = c(-1,0,1,0,0,0,0,0,0,0))
topTags(lrt.bolting.vs.flowering.ae)
summary(de.bolting.vs.flowering.ae <- decideTestsDGE(lrt.bolting.vs.flowering.ae, p=0.05))
DEgene.bolting.vs.flowering.ae <- topTags(lrt.bolting.vs.flowering.ae,n = Inf)$table[topTags(lrt.bolting.vs.flowering.ae,n = Inf)$table$FDR<0.05,]
# -1   736
# 0  57389
# 1   2850

lrt.flowering.vs.early.silique.ae <- glmLRT(fit.parent, contrast = c(0,1,-1,0,0,0,0,0,0,0))
topTags(lrt.flowering.vs.early.silique.ae)
summary(de.flowering.vs.early.silique.ae <- decideTestsDGE(lrt.flowering.vs.early.silique.ae, p=0.05))
DEgene.flowering.vs.early.silique.ae <- topTags(lrt.flowering.vs.early.silique.ae,n = Inf)$table[topTags(lrt.flowering.vs.early.silique.ae,n = Inf)$table$FDR<0.05,]
# -1  3811
# 0  56054
# 1   1110

lrt.early.vs.late.silique.ae <- glmLRT(fit.parent, contrast = c(0,-1,0,1,0,0,0,0,0,0))
topTags(lrt.early.vs.late.silique.ae)
summary(de.early.vs.late.silique.ae <- decideTestsDGE(lrt.early.vs.late.silique.ae, p=0.05))
DEgene.early.vs.late.silique.ae <- topTags(lrt.early.vs.late.silique.ae,n = Inf)$table[topTags(lrt.early.vs.late.silique.ae,n = Inf)$table$FDR<0.05,]
# -1   285
# 0  60430
# 1    260

### ggplot 
### reformat for ggplot barplot 
DEGs_number_between_tissue.ae <- data.frame("bolting-vs-young" = as.data.frame(summary(de.young.vs.bolting.ae <- decideTestsDGE(lrt.young.vs.bolting.ae, p=0.05)))$Freq, 
                                            "flowering-vs-bolting" = as.data.frame(summary(de.bolting.vs.flowering.ae <- decideTestsDGE(lrt.bolting.vs.flowering.ae, p=0.05)))$Freq,
                                     "early.slique-vs-flowering"= as.data.frame(summary(de.flowering.vs.early.silique.ae <- decideTestsDGE(lrt.flowering.vs.early.silique.ae, p=0.05)))$Freq, 
                                     "late.silique-vs-early.silique" = as.data.frame(summary(de.early.vs.late.silique.ae <- decideTestsDGE(lrt.early.vs.late.silique.ae, p=0.05)))$Freq)

rownames(DEGs_number_between_tissue.ae) <- c("down", "no", "up")
DEGs_number_between_tissue.ae <- DEGs_number_between_tissue.ae[c("down", "up"),]
DEGs_number_between_tissue.ae

DEGs_number_between_tissue.ae.melt.parent <- melt(DEGs_number_between_tissue.ae)
DEGs_number_between_tissue.ae.melt.parent$DE <- rep(c("down", "up"), 4)
colnames(DEGs_number_between_tissue.ae.melt.parent) <- c("tissue", "number", "DE")
DEGs_number_between_tissue.ae.melt.parent

# reorder: up 1st down 2nd 
DEGs_number_between_tissue.ae.melt.parent$DE <- factor(DEGs_number_between_tissue.ae.melt.parent$DE, levels = c("up", "down"))

DEGs_number_between_tissue.ae.melt.parent <- DEGs_number_between_tissue.ae.melt.parent[order(DEGs_number_between_tissue.ae.melt.parent$DE),]
DEGs_number_between_tissue.ae.melt.parent

### making ggplot for DEGs 
p.DEGs_number_between_tissue.ae.parent <- ggplot(data = DEGs_number_between_tissue.ae.melt.parent) + theme_grey()
p.DEGs_number_between_tissue.ae.parent <- p.DEGs_number_between_tissue.ae.parent + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number), stat = "identity")
p.DEGs_number_between_tissue.ae.parent <- p.DEGs_number_between_tissue.ae.parent + facet_grid(~tissue)
p.DEGs_number_between_tissue.ae.parent <- p.DEGs_number_between_tissue.ae.parent + geom_text(data=DEGs_number_between_tissue.ae.melt.parent,aes(x=DE,y=number*1.05,label=factor(number),color=DE)) 
p.DEGs_number_between_tissue.ae.parent <- p.DEGs_number_between_tissue.ae.parent + labs(y = "number of differentially expressed genes", x = "")

p.DEGs_number_between_tissue.ae.parent
ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/cDNA_mapped/DEGs_number_between_tissue.ae.parent.png", width = 11, height = 8)

### Da-Ol-1 
lrt.young.vs.bolting.ol <- glmLRT(fit.parent, contrast = c(0,0,0,0,0,1,0,0,0,-1))
topTags(lrt.young.vs.bolting.ol)
summary(de.young.vs.bolting.ol <- decideTestsDGE(lrt.young.vs.bolting.ol, p=0.05))
DEgene.young.vs.bolting.ol <- topTags(lrt.young.vs.bolting.ol,n = Inf)$table[topTags(lrt.young.vs.bolting.ol,n = Inf)$table$FDR<0.05,]
# -1  2683
# 0  54969
# 1   3323

lrt.bolting.vs.flowering.ol <- glmLRT(fit.parent, contrast = c(0,0,0,0,0,-1,0,1,0,0))
topTags(lrt.bolting.vs.flowering.ol)
summary(de.bolting.vs.flowering.ol <- decideTestsDGE(lrt.bolting.vs.flowering.ol, p=0.05))
DEgene.bolting.vs.flowering.ol <- topTags(lrt.bolting.vs.flowering.ol,n = Inf)$table[topTags(lrt.bolting.vs.flowering.ol,n = Inf)$table$FDR<0.05,]
# -1   914
# 0  58313
# 1   1748

lrt.flowering.vs.early.silique.ol <- glmLRT(fit.parent, contrast = c(0,0,0,0,0,0,1,-1,0,0))
topTags(lrt.flowering.vs.early.silique.ol)
summary(de.flowering.vs.early.silique.ol <- decideTestsDGE(lrt.flowering.vs.early.silique.ol, p=0.05))
DEgene.flowering.vs.early.silique.ol <- topTags(lrt.flowering.vs.early.silique.ol,n = Inf)$table[topTags(lrt.flowering.vs.early.silique.ol,n = Inf)$table$FDR<0.05,]
# -1  4845
# 0  54087
# 1   2043

lrt.early.vs.late.silique.ol <- glmLRT(fit.parent, contrast = c(0,0,0,0,0,0,-1,0,1,0))
topTags(lrt.early.vs.late.silique.ol)
summary(de.early.vs.late.silique.ol <- decideTestsDGE(lrt.early.vs.late.silique.ol, p=0.05))
DEgene.early.vs.late.silique.ol <- topTags(lrt.early.vs.late.silique.ol,n = Inf)$table[topTags(lrt.early.vs.late.silique.ol,n = Inf)$table$FDR<0.05,]
# -1  1042
# 0  58199
# 1   1734

### ggplot 
### reformat for ggplot barplot 
DEGs_number_between_tissue.ol <- data.frame("bolting-vs-young" = as.data.frame(summary(de.young.vs.bolting.ol <- decideTestsDGE(lrt.young.vs.bolting.ol, p=0.05)))$Freq, 
                                            "flowering-vs-bolting" = as.data.frame(summary(de.bolting.vs.flowering.ol <- decideTestsDGE(lrt.bolting.vs.flowering.ol, p=0.05)))$Freq,
                                     "early.slique-vs-flowering"= as.data.frame(summary(de.flowering.vs.early.silique.ol <- decideTestsDGE(lrt.flowering.vs.early.silique.ol, p=0.05)))$Freq, 
                                     "late.silique-vs-early.silique" = as.data.frame(summary(de.early.vs.late.silique.ol <- decideTestsDGE(lrt.early.vs.late.silique.ol, p=0.05)))$Freq)

rownames(DEGs_number_between_tissue.ol) <- c("down", "no", "up")
DEGs_number_between_tissue.ol <- DEGs_number_between_tissue.ol[c("down", "up"),]
DEGs_number_between_tissue.ol

DEGs_number_between_tissue.ol.melt.parent <- melt(DEGs_number_between_tissue.ol)
DEGs_number_between_tissue.ol.melt.parent$DE <- rep(c("down", "up"), 4)
colnames(DEGs_number_between_tissue.ol.melt.parent) <- c("tissue", "number", "DE")
DEGs_number_between_tissue.ol.melt.parent

# reorder: up 1st down 2nd 
DEGs_number_between_tissue.ol.melt.parent$DE <- factor(DEGs_number_between_tissue.ol.melt.parent$DE, levels = c("up", "down"))

DEGs_number_between_tissue.ol.melt.parent <- DEGs_number_between_tissue.ol.melt.parent[order(DEGs_number_between_tissue.ol.melt.parent$DE),]
DEGs_number_between_tissue.ol.melt.parent

### making ggplot for DEGs 
p.DEGs_number_between_tissue.ol.parent <- ggplot(data = DEGs_number_between_tissue.ol.melt.parent) + theme_grey()
p.DEGs_number_between_tissue.ol.parent <- p.DEGs_number_between_tissue.ol.parent + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number), stat = "identity")
p.DEGs_number_between_tissue.ol.parent <- p.DEGs_number_between_tissue.ol.parent + facet_grid(~tissue)
p.DEGs_number_between_tissue.ol.parent <- p.DEGs_number_between_tissue.ol.parent + geom_text(data=DEGs_number_between_tissue.ol.melt.parent,aes(x=DE,y=number*1.05,label=factor(number),color=DE)) 
p.DEGs_number_between_tissue.ol.parent <- p.DEGs_number_between_tissue.ol.parent + labs(y = "number of differentially expressed genes", x = "")

p.DEGs_number_between_tissue.ol.parent
ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/cDNA_mapped/DEGs_number_between_tissue.ol.parent.png", width = 11, height = 8) 

```

# explore the silique problem, much fewer genes were differentially expressed between early & late silique in Da-Ae than in Da-Ol-1, why? (actually De-Ae has fewer DEGs between all stages, need to find out why later) 
```{r}
# calcuate overlaps & unique genes 
dim(DEgene.young.vs.bolting.ol) # 6006    5 
dim(DEgene.young.vs.bolting.ae) # 1103    5
summary(rownames(DEgene.young.vs.bolting.ol) %in% rownames(DEgene.young.vs.bolting.ae)) 
# Mode   FALSE    TRUE    NA's 
# logical    5556     450       0 
dim(DEgene.bolting.vs.flowering.ol) # 2662    5 
dim(DEgene.bolting.vs.flowering.ae) # 3586    5 
summary(rownames(DEgene.bolting.vs.flowering.ol) %in% rownames(DEgene.bolting.vs.flowering.ae)) 
#    Mode   FALSE    TRUE    NA's 
# logical    1528    1134       0
dim(DEgene.flowering.vs.early.silique.ol) # 6888    5 
dim(DEgene.flowering.vs.early.silique.ae) # 4921    5 
summary(rownames(DEgene.flowering.vs.early.silique.ol) %in% rownames(DEgene.flowering.vs.early.silique.ae)) 
#   Mode   FALSE    TRUE    NA's 
# logical    3200    3688       0 
dim(DEgene.early.vs.late.silique.ol) # 2776 5 
dim(DEgene.early.vs.late.silique.ae) # 545 5 
summary(rownames(DEgene.early.vs.late.silique.ol) %in% rownames(DEgene.early.vs.late.silique.ae)) 
#    Mode   FALSE    TRUE    NA's 
# logical    2436     340       0 

# draw venn diagram 
# install.packages("VennDiagram")
require(VennDiagram)

# young vs bolting
grid.newpage()
png(filename = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/young.vs.bolting.png")
venn.plot <- draw.pairwise.venn(6006, 1103, 450, # 1st number for group A, 2nd for group B, 3rd for overlaps 
                                c("Da-Ol-1", "Da-Ae"),
                                fill = c("lightblue","pink"),
                                lty          = "blank",
                                cex = 2,
                                cat.cex = 1.5,
                                cat.pos = 9)
grid.draw(venn.plot)
dev.off()

# bolting vs flowering 
grid.newpage()
png(filename = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/bolting.vs.flowering.png")
venn.plot <- draw.pairwise.venn(2662, 3586, 1134, # 1st number for group A, 2nd for group B, 3rd for overlaps 
                                c("Da-Ol-1", "Da-Ae"),
                                fill = c("lightblue","pink"),
                                lty          = "blank",
                                cex = 2,
                                cat.cex = 1.5,
                                cat.pos = 9)
grid.draw(venn.plot)
dev.off()

# flowering vs early silique 
grid.newpage()
png(filename = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/flowering.vs.early.silique.png")
venn.plot <- draw.pairwise.venn(6888, 4921, 3688, # 1st number for group A, 2nd for group B, 3rd for overlaps 
                                c("Da-Ol-1", "Da-Ae"),
                                fill = c("lightblue","pink"),
                                lty          = "blank",
                                cex = 2,
                                cat.cex = 1.5,
                                cat.pos = 9)
grid.draw(venn.plot)
dev.off()

# early vs late silique 
grid.newpage()
png(filename = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/early.vs.late.silique.png")
venn.plot <- draw.pairwise.venn(2776, 545, 340, # 1st number for group A, 2nd for group B, 3rd for overlaps 
                                c("Da-Ol-1", "Da-Ae"),
                                fill = c("lightblue","pink"),
                                lty          = "blank",
                                cex = 2,
                                cat.cex = 1.5,
                                cat.pos = 9)
grid.draw(venn.plot)
dev.off()

# explore the uniquely differentially expressed genes in Da-Ol-1, what are their expression patterns in Da-Ae? 
# 1) find genes that are only differentially expressed in Da-Ol-1 
index <- which(!(rownames(DEgene.early.vs.late.silique.ol) %in% rownames(DEgene.early.vs.late.silique.ae)))

# 2) extract the expression values of these genes from early, late silique of Da-Ae & Da-Ol-1 
uniq.gene.ol <- parent.read.count.one.small[index,]
uniq.gene.ol <- uniq.gene.ol[,c("Da-Ae_early-silique_3", "Da-Ol-1_early-silique_3", "Da-Ol-1_late-silique_3", "Da-Ae_late-silique_3", "Da-Ae_early-silique_1", "Da-Ae_early-silique_2", "Da-Ae_late-silique_1", "Da-Ae_late-silique_2", "Da-Ol-1_early-silique_1", "Da-Ol-1_early-silique_2", "Da-Ol-1_late-silique_1", "Da-Ol-1_late-silique_2")]

dim(uniq.gene.ol) # 2436   12 
```

# fatty acid biosynthesis pathways genes (GO:0006633)
```{r}
fatty_acid_gene_index <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/fatty_acid_biosynthesis_genes_ID.txt", header = F)
head(fatty_acid_gene_index)
nrow(fatty_acid_gene_index)
rownames(fatty_acid_gene_index) <- fatty_acid_gene_index$V1

# extract only fatty acid genes from dataset 
fatty.acid.genes <- (merge(parent.read.count.one.small, fatty_acid_gene_index, by="row.names"))[,1:28]
head(fatty.acid.genes)
rownames(fatty.acid.genes) <- fatty.acid.genes$Row.names
fatty.acid.genes <- fatty.acid.genes[,-1]
dim(fatty.acid.genes) # 207 27
fatty.acid.genes$total <- rowSums(fatty.acid.genes)
dim(fatty.acid.genes)
```

# new design to see how many fatty acid genes' expression are differentially expressed between genotypes and during developmental stages  
```{r}
source("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/function_BnRNAseq.R")
# assign group 
ftable(parent.read.count.one.sample,row.vars=c("stage"),col.vars=c("batch","genotype"))

# normalize 
dge.new <- DGEList(counts=parent.read.count.one.small, group=parent.read.count.one.sample$group)
dim(dge.new) # [1] 30876    28 
dge.new <- calcNormFactors(dge.new, method = "TMM") 
dge.new$sample
# design matrix 
parent.read.count.one.sample$genotype <- as.factor(parent.read.count.one.sample$genotype)
parent.read.count.one.sample$tissue <- as.factor(parent.read.count.one.sample$stage)
parent.read.count.one.sample$genotype <- relevel(parent.read.count.one.sample$genotype,ref="Da-Ol-1")
parent.read.count.one.sample$tissue <- relevel(parent.read.count.one.sample$tissue,ref="Young")

design.new <- model.matrix(~tissue*genotype,data = parent.read.count.one.sample) 
colnames(design.new)

# calculate dispersion
dge.new <- estimateGLMCommonDisp(dge.new, design.new,verbose = TRUE) # Disp = 0.25646 , BCV = 0.5064 
dge.new <- estimateGLMTrendedDisp(dge.new,design.new)
dge.new <- estimateGLMTagwiseDisp(dge.new,design.new)
# plotBCV(dge.new)

# use interaction as coefficient and see how many fatty acid genes are important for interaction term... 
fit.new <- glmFit(dge.new, design.new)
lrt.new.interaction <- glmLRT(fit.new,coef = c("tissuebolting:genotypeDa-Ae", "tissueearly-silique:genotypeDa-Ae", "tissueflowering:genotypeDa-Ae", "tissuelate-silique:genotypeDa-Ae"))
topTags(lrt.new.interaction)
DEgene.new.interaction <- topTags(lrt.new.interaction,n = Inf)$table[topTags(lrt.new.interaction,n = Inf)$table$FDR<0.05,]
nrow(DEgene.new.interaction) # number of genes with interaction effect in any tissue types 

# genes for tissue types 
lrt.new.tissue <- glmLRT(fit.new,coef = c("tissuebolting", "tissueearly-silique", "tissueflowering", "tissuelate-silique"))
topTags(lrt.new.tissue)
DEgene.new.tissue <- topTags(lrt.new.tissue,n = Inf)$table[topTags(lrt.new.tissue,n = Inf)$table$FDR<0.05,]
nrow(DEgene.new.tissue) # number of genes for tissue effect  22580  
 
# genes for gt 
lrt.new.gt <- glmLRT(fit.new,coef = c("genotypeDa-Ae", "tissuebolting:genotypeDa-Ae", "tissueearly-silique:genotypeDa-Ae", "tissueflowering:genotypeDa-Ae", "tissuelate-silique:genotypeDa-Ae"))
topTags(lrt.new.gt)
DEgene.new.gt <- topTags(lrt.new.gt,n = Inf)$table[topTags(lrt.new.gt,n = Inf)$table$FDR<0.05,]
nrow(DEgene.new.gt) # number of genes for gt effect # 15098   

dim(fatty.acid.genes) # 207 fatty acid genes expressed in my data  

index_2 <- which(rownames(fatty.acid.genes) %in% rownames(DEgene.new.interaction)) 
rownames(fatty.acid.genes[index_2,])
ID.15.fatty.acid.genes <- rownames(fatty.acid.genes[which(rownames(fatty.acid.genes) %in% rownames(DEgene.new.interaction)),])
fatty.acid.genes.annotation <- read.csv("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/fatty_acid_gene_name.csv", header = F)

expression.pattern.Bn.parent.with.annot(ID.15.fatty.acid.genes, fatty.acid.genes.annotation)

# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/fatty.acid.interaction.png", width = 8, height = 20)

# get only 3 genes 
ID <- c("BnaA05g06830D","BnaA10g09300D",  "BnaC09g50630D") 
expression.pattern.Bn.parent.with.annot(ID, fatty.acid.genes.annotation)

# ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/fatty.acid.three.png", width = 11, height = 8)
```

# GO enrichment analysis 
```{r}
# load all necessary functions 
# source("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/function_BnRNAseq.R")

# young 
DEgene.young.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.young))
DEgene.young.GO.ORA
write.csv(DEgene.young.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/DEgene.young.GO.ORA.csv")

# bolting 
DEgene.bolting.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.bolting))
DEgene.bolting.GO.ORA
write.csv(DEgene.bolting.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/DEgene.bolting.GO.ORA.csv")

# flowering 
DEgene.flowering.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.flowering))
DEgene.flowering.GO.ORA
write.csv(DEgene.flowering.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/DEgene.flowering.GO.ORA.csv")

# early silique 
DEgene.early.silique.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.early.silique))
DEgene.early.silique.GO.ORA
write.csv(DEgene.early.silique.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/DEgene.early.silique.GO.ORA.csv")

DEgene.late.silique.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.late.silique))
DEgene.late.silique.GO.ORA
write.csv(DEgene.late.silique.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/DEgene.late.silique.GO.ORA.csv")

DEgene.silique.ae.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.early.vs.late.silique.ae))
DEgene.silique.ae.GO.ORA
write.csv(DEgene.late.silique.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/DEgene.silique.ae.GO.ORA.csv")

DEgene.silique.ol.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.early.vs.late.silique.ol))
DEgene.silique.ol.GO.ORA
write.csv(DEgene.silique.ol.GO.ORA, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/DEgene.silique.ol.GO.ORA.csv")

DEgene.young.vs.bolting.ol.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.young.vs.bolting.ol))
DEgene.young.vs.bolting.ol.GO.ORA

DEgene.young.vs.bolting.ae.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.young.vs.bolting.ae))
DEgene.young.vs.bolting.ae.GO.ORA

DEgene.bolting.vs.flowering.ol.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.bolting.vs.flowering.ol))
DEgene.bolting.vs.flowering.ol.GO.ORA

DEgene.bolting.vs.flowering.ae.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.bolting.vs.flowering.ae))
DEgene.bolting.vs.flowering.ae.GO.ORA

DEgene.flowering.vs.early.silique.ol.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.flowering.vs.early.silique.ol))
DEgene.bolting.vs.flowering.ol.GO.ORA

DEgene.flowering.vs.early.silique.ae.GO.ORA <- GOseq.Bn.ORA(rownames(DEgene.flowering.vs.early.silique.ae))
DEgene.bolting.vs.flowering.ae.GO.ORA
```

### other important genes 
# Acetyl-CoA-ACP-transacetylase: GO:0004313 none 
# B-Ketoacyl-ACP-synthase: IPR014030 --> A 
# Malonyl-CoA-ACP-transacetylase: IPR016036 --> B  
# B-ketoacyl-ACP-reductase: GO:0004316 --> C 
# Enoyl-ACP-hydrase: IPR018376 --> D
# Enoyl-ACP-reductase: GO:0016631 & GO:0004319 none 
# Palmitoyl-thioesterase: IPR002472 --> E 
```{r}
source("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/function_BnRNAseq.R")

# B-Ketoacyl-ACP-synthase  
A.ID <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/B-Ketoacyl-ACP-synthase", header = F)
dim(A.ID)
expression.pattern.Bn.parent(A.ID) 
# ggsave("~/Desktop/Brassica_project/KIAT_RNA_seq/figure/A.png", width = 8, height = 20) 

# Malonyl-CoA-ACP-transacetylase  
B.ID <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/Malonyl-CoA-ACP-transacetylase", header = F) 
dim(B.ID)
expression.pattern.Bn.parent(B.ID)
# ggsave("~/Desktop/Brassica_project/KIAT_RNA_seq/figure/B.png", width = 8, height = 10) 

# B-ketoacyl-ACP-reductase: GO:0004316 --> C 
C.ID <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/B-ketoacyl-ACP-reductase", header = F) 
dim(C.ID)
expression.pattern.Bn.parent(C.ID)
# ggsave("~/Desktop/Brassica_project/KIAT_RNA_seq/figure/C.png", width = 8, height = 8) 

# Enoyl-ACP-hydrase: IPR018376 --> D
D.ID <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/Enoyl-ACP-hydrase", header = F)
dim(D.ID)
expression.pattern.Bn.parent(D.ID)
# ggsave("~/Desktop/Brassica_project/KIAT_RNA_seq/figure/D.png", width = 8, height = 20) 

# Palmitoyl-thioesterase: IPR002472 --> E 
E.ID <- read.delim("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/Palmitoyl-thioesterase", header = F) 
dim(E.ID)
expression.pattern.Bn.parent(E.ID)  
# ggsave("~/Desktop/Brassica_project/KIAT_RNA_seq/figure/E.png", width = 8, height = 20) 
```

# get Arabidopsis hit for every B.napus cds (reciprocal blast, need to write out the bash script for reciprocal blast...)
```{r}
# blast: reciprocal blast bewteen Arabidopsis & B.napus 
# merge file to get Arabidopsis hit for each B.napus gene:  
ara.ID <- read.delim("~/Desktop/Brassica_project/reference/ara_ID.txt")

# awk 'NR==FNR{a[$1]=$0}NR>FNR{if ($2 in a) print $0 "\t" a[$2]}' ara_ID.txt B.napus_vs_ara.final | awk 'BEGIN{FS="\t"}{print $1 "\t" $3 "\t" $5}' > napus_vs_ara.final.final  

napus_vs_ara.final <- read.delim("~/Desktop/Brassica_project/reference/napus_vs_ara.final")
rownames(napus_vs_ara.final) <- napus_vs_ara.final$napus_ID
head(napus_vs_ara.final)

napus_vs_ara.non_reciprocal <- read.delim("~/Desktop/Brassica_project/reference/napus_vs_ara.non_reciprocal")
colnames(napus_vs_ara.non_reciprocal) <- c("napus_ID", "ara_ID", "description")
rownames(napus_vs_ara.non_reciprocal) <- napus_vs_ara.non_reciprocal$napus_ID
head(napus_vs_ara.non_reciprocal)
```

# Q: whether 1cm/early or 5cm/late silique is the better choice for RNAseq for looking at differences in fatty acid synthesis. 
```{r}
# 1) used GO# & IPR# related to different fatty acid pathways as keywords to search for gene IDs the reason that I didn't use Arabidopsis blast results is: the inconsistency between the annotation of gene list from FnP & the annotation in Arabidopsis dataset. 

# data 
# gene list related to fatty acid biosynthesis 
rownames(fatty.acid.genes) # fatty acid biosynthesis pathways genes (GO:0006633) 
A.ID # B-Ketoacyl-ACP-synthase: IPR014030
B.ID # Malonyl-CoA-ACP-transacetylase: IPR016036 
C.ID # B-ketoacyl-ACP-reductase: GO:0004316
D.ID # Enoyl-ACP-hydrase: IPR018376
E.ID # Palmitoyl-thioesterase: IPR002472 

length(unique(c(rownames(fatty.acid.genes), as.character(A.ID$V1), as.character(B.ID$V1), as.character(C.ID$V1), as.character(D.ID$V1), as.character(E.ID$V1)))) # 247 genes in total 

# all the expression result 
DEgene.early.vs.late.silique.ae
DEgene.early.vs.late.silique.ol
DEgene.early.silique
DEgene.late.silique

# extract fold change value & p-value from DEG dataset 
# between early & late silique 
DEgene.early.vs.late.silique.ae[rownames(DEgene.early.vs.late.silique.ae) %in% rownames(fatty.acid.genes),]
DEgene.early.vs.late.silique.ae[rownames(DEgene.early.vs.late.silique.ae) %in% A.ID$V1,]
DEgene.early.vs.late.silique.ae[rownames(DEgene.early.vs.late.silique.ae) %in% B.ID$V1,]
DEgene.early.vs.late.silique.ae[rownames(DEgene.early.vs.late.silique.ae) %in% C.ID$V1,]
DEgene.early.vs.late.silique.ae[rownames(DEgene.early.vs.late.silique.ae) %in% D.ID$V1,]
DEgene.early.vs.late.silique.ae[rownames(DEgene.early.vs.late.silique.ae) %in% E.ID$V1,]

gene.1 <- DEgene.early.vs.late.silique.ae[rownames(DEgene.early.vs.late.silique.ae) %in% rownames(fatty.acid.genes),]
gene.1
# only one fatty acid related gene is differentially expressed between early & late silique for Da-Ae 

DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% rownames(fatty.acid.genes),] 
DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% A.ID$V1,]
DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% B.ID$V1,]
DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% C.ID$V1,]
DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% D.ID$V1,]
DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% E.ID$V1,]

oil.synthesis.related.early.vs.late.ol <- unique(rbind(
  DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% rownames(fatty.acid.genes),], 
  DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% A.ID$V1,],
  DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% B.ID$V1,],
  DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% C.ID$V1,],
  DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% D.ID$V1,],
  DEgene.early.vs.late.silique.ol[rownames(DEgene.early.vs.late.silique.ol) %in% E.ID$V1,]
))

gene.9 <- oil.synthesis.related.early.vs.late.ol # 10 genes are differentially expressed between early & late silique for Da-Ol-1 

# between two genotypes at early & late silique stages 
DEgene.early.silique[rownames(DEgene.early.silique) %in% rownames(fatty.acid.genes),]
DEgene.early.silique[(rownames(DEgene.early.silique) %in% A.ID$V1),]
DEgene.early.silique[(rownames(DEgene.early.silique) %in% B.ID$V1),]
DEgene.early.silique[(rownames(DEgene.early.silique) %in% C.ID$V1),]
DEgene.early.silique[(rownames(DEgene.early.silique) %in% D.ID$V1),]
DEgene.early.silique[(rownames(DEgene.early.silique) %in% E.ID$V1),]

oil.synthesis.related.early.silique <- unique(rbind(
  DEgene.early.silique[rownames(DEgene.early.silique) %in% rownames(fatty.acid.genes),],
  DEgene.early.silique[(rownames(DEgene.early.silique) %in% A.ID$V1),],
  DEgene.early.silique[(rownames(DEgene.early.silique) %in% B.ID$V1),],
  DEgene.early.silique[(rownames(DEgene.early.silique) %in% C.ID$V1),],
  DEgene.early.silique[(rownames(DEgene.early.silique) %in% D.ID$V1),],
  DEgene.early.silique[(rownames(DEgene.early.silique) %in% E.ID$V1),]
))

gene.28 <- oil.synthesis.related.early.silique # 28 genes are differentially expressed between Da-Ae & Da-Ol-1 at early silique stage 
gene.28
# write.table(gene.28, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/data/gene.28.txt")

DEgene.late.silique[rownames(DEgene.late.silique) %in% rownames(fatty.acid.genes),]
DEgene.late.silique[(rownames(DEgene.late.silique) %in% A.ID$V1),]
DEgene.late.silique[(rownames(DEgene.late.silique) %in% B.ID$V1),]
DEgene.late.silique[(rownames(DEgene.late.silique) %in% C.ID$V1),]
DEgene.late.silique[(rownames(DEgene.late.silique) %in% D.ID$V1),]
DEgene.late.silique[(rownames(DEgene.late.silique) %in% E.ID$V1),]

oil.synthesis.related.late.silique <- unique(rbind(
  DEgene.late.silique[rownames(DEgene.late.silique) %in% rownames(fatty.acid.genes),],
  DEgene.late.silique[(rownames(DEgene.late.silique) %in% A.ID$V1),],
  DEgene.late.silique[(rownames(DEgene.late.silique) %in% B.ID$V1),],
  DEgene.late.silique[(rownames(DEgene.late.silique) %in% C.ID$V1),],
  DEgene.late.silique[(rownames(DEgene.late.silique) %in% D.ID$V1),],
  DEgene.late.silique[(rownames(DEgene.late.silique) %in% E.ID$V1),]
))

gene.36 <- oil.synthesis.related.late.silique # 36 genes are differentially expressed between Da-Ae & Da-Ol-1 at late silique stage 
gene.36
write.table(gene.36, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/data/gene.36.txt")

gene.annotation <- read.csv("~/Desktop/Brassica_project/KIAT_RNA_seq/data/gene_annotation.csv", header = F)
gene.annotation <- gene.annotation[,c(1:2)]
rownames(gene.annotation) <- gene.annotation$V1

### combined gene list 
gene <- read.table("~/Desktop/Brassica_project/KIAT_RNA_seq/data/gene.txt")
gene

rownames(gene.annotation) %in% gene$V1

## make tables 
gene.1 <- merge(gene.1, gene.annotation, by="row.names")[,c(1,2,3,4,5,6,8)] 
gene.1 
colnames(gene.1) <- c("geneID", "logFC", "logCPM", "LR", "PValue", "FDR", "Arabidopsis.hit")
write.csv(gene.1, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/data/gene.1.csv")

gene.9 <- merge(gene.9, gene.annotation, by="row.names")[,c(1,2,3,4,5,6,8)]
gene.9
colnames(gene.9) <- c("geneID", "logFC", "logCPM", "LR", "PValue", "FDR", "Arabidopsis.hit")
write.csv(gene.9, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/data/gene.9.csv")

gene.28 <- merge(gene.28, gene.annotation, by="row.names")[,c(1,2,3,4,5,6,8)]
gene.28
colnames(gene.28) <- c("geneID", "logFC", "logCPM", "LR", "PValue", "FDR", "Arabidopsis.hit")
write.csv(gene.28, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/data/gene.28.csv")

gene.36 <- merge(gene.36, gene.annotation, by="row.names")[,c(1,2,3,4,5,6,8)]
gene.36
colnames(gene.36) <- c("geneID", "logFC", "logCPM", "LR", "PValue", "FDR", "Arabidopsis.hit")
write.csv(gene.36, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/data/gene.36.csv")

rownames(fatty.acid.genes) 
vstMat.parent

### are fatty acid genes expressed at higher level in early or late silique? 
summary(de.early.vs.late.silique.ae.1 <- decideTestsDGE(lrt.early.vs.late.silique.ae, p=1))
DEgene.early.vs.late.silique.ae.1 <- topTags(lrt.early.vs.late.silique.ae,n = Inf)$table[topTags(lrt.early.vs.late.silique.ae,n = Inf)$table$FDR<0.075,]

summary(de.early.vs.late.silique.ol.1 <- decideTestsDGE(lrt.early.vs.late.silique.ol, p=1))
DEgene.early.vs.late.silique.ol.1 <- topTags(lrt.early.vs.late.silique.ol,n = Inf)$table[topTags(lrt.early.vs.late.silique.ol,n = Inf)$table$FDR<0.075,]

all.fatty.acid.genes <- unique(c(rownames(fatty.acid.genes), as.character(A.ID$V1), as.character(B.ID$V1), as.character(C.ID$V1), as.character(D.ID$V1), as.character(E.ID$V1)))

# check fatty acid genes expression 
all.fatty.acid.genes.ol.1.early.vs.late <- DEgene.early.vs.late.silique.ol.1[rownames(DEgene.early.vs.late.silique.ol.1) %in% all.fatty.acid.genes,]
dim(all.fatty.acid.genes.ol.1.early.vs.late)

sum(all.fatty.acid.genes.ol.1.early.vs.late$logFC >0) # 135 # 4
sum(all.fatty.acid.genes.ol.1.early.vs.late$logFC < 0) # 108 # 5

all.fatty.acid.genes.ae.1.early.vs.late <- DEgene.early.vs.late.silique.ae.1[rownames(DEgene.early.vs.late.silique.ae.1) %in% all.fatty.acid.genes,]
dim(all.fatty.acid.genes.ae.1.early.vs.late)

sum(all.fatty.acid.genes.ae.1.early.vs.late$logFC > 0) # 96 # 0 
sum(all.fatty.acid.genes.ae.1.early.vs.late$logFC < 0) # 147 # 2
```

# use mapman for fatty acid pathway analysis 
```{r}
### We would like to present the result in mapman, below I format my data for mapman visualization 
head(napus_vs_ara.final)
dim(napus_vs_ara.final)
napus_vs_ara.final$ara_locus_ID <- gsub("([[:print:]]+)(\\.)([[:digit:]]+)","\\1",napus_vs_ara.final$ara_ID)
head(napus_vs_ara.final)

list.between.gt <- list(DEgene.young, DEgene.bolting, DEgene.flowering, DEgene.early.silique, DEgene.late.silique)

list.between.stages.ae <- list(DEgene.young.vs.bolting.ae, DEgene.bolting.vs.flowering.ae, DEgene.flowering.vs.early.silique.ae, DEgene.early.vs.late.silique.ae)
 
list.between.stages.ol <- list(DEgene.young.vs.bolting.ol, DEgene.bolting.vs.flowering.ol, DEgene.flowering.vs.early.silique.ol, DEgene.early.vs.late.silique.ol)


DEG.between.gt <- lapply(list.between.gt, function(DEgene){
  DEG <- data.frame(napus_ID = rownames(DEgene),
                    fold_change = DEgene$logFC)  
}
)

DEG.between.stage.ae <- lapply(list.between.stages.ae, function(DEgene){
  DEG <- data.frame(napus_ID = rownames(DEgene),
                    fold_change = DEgene$logFC)  
}
)

DEG.between.stage.ol <- lapply(list.between.stages.ol, function(DEgene){
  DEG <- data.frame(napus_ID = rownames(DEgene),
                    fold_change = DEgene$logFC)  
}
)

between.gt.ID <- as.list(c("young", "bolting", "flowering", "early_silique", "late_silique"))
for (i in c(1:5)){
  colnames(DEG.between.gt[[i]])[2] <- between.gt.ID[[i]]
}

between.stage.ID.ae <- as.list(c("bolting_vs_young_ae", "flowering_vs_bolting_ae", "early_silique_vs_flowering_ae", "late_vs_early_silique_ae"))
for (i in c(1:4)){
  colnames(DEG.between.stage.ae[[i]])[2] <- between.stage.ID.ae[[i]]
}

between.stage.ID.ol <- as.list(c("bolting_vs_young_ol", "flowering_vs_bolting_ol", "early_silique_vs_flowering_ol", "late_vs_early_silique_ol"))
for (i in c(1:4)){
  colnames(DEG.between.stage.ol[[i]])[2] <- between.stage.ID.ol[[i]]
}

(DEG.between.gt[[1]])

tmp <- merge(DEG.between.gt[[1]], DEG.between.gt[[2]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.gt[[3]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.gt[[4]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.gt[[5]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.stage.ae[[1]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.stage.ae[[2]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.stage.ae[[3]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.stage.ae[[4]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.stage.ol[[1]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.stage.ol[[2]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.stage.ol[[3]], by="napus_ID", all=T)
tmp <- merge(tmp, DEG.between.stage.ol[[4]], by="napus_ID", all=T)
colnames(tmp)
# save the file for checking 
napus_vs_ara.full <- tmp
tmp <- merge(tmp, napus_vs_ara.final, by="napus_ID")[,c(2,3,4,5,6,7,8,9,10,11,12,13,14,17)]
head(tmp)

# double check I get everything right 
dim(DEgene.young); dim(DEgene.bolting); dim(DEgene.flowering); dim(DEgene.early.silique); dim(DEgene.late.silique); dim(DEgene.young.vs.bolting.ae); dim(DEgene.bolting.vs.flowering.ae); dim(DEgene.flowering.vs.early.silique.ae); dim(DEgene.early.vs.late.silique.ae); dim(DEgene.young.vs.bolting.ol); dim(DEgene.bolting.vs.flowering.ol); dim(DEgene.flowering.vs.early.silique.ol); dim(DEgene.early.vs.late.silique.ol) 

dim(tmp)
for (i in c(2:14)){
  print(sum(!is.na(tmp[,i])))
}

write.csv(tmp, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/output/mapman_input_all_final.csv")
# cat tmp.csv | sed 's/NA/X/g' > tmp_all_final_2.csv

# the list of genes 
# 11.1.1 AT1G36160 AT1G36180
# 11.1.2 AT2G38040 ATCG00500 AT3G15690 AT5G15530 AT5G16390 AT5G35360 AT3G56130 AT2G30200
# 11.1.3 AT1G62640 AT1G74960 AT2G04540 AT5G46290 
# 11.1.4 AT1G24360 
# 11.1.5 AT2G22230 AT5G10160 
# 11.1.6 AT2G05990 

Ara_fatty_acid_ID <- c("AT1G36160", "AT1G36180", "AT2G38040", "ATCG00500", "AT3G15690", "AT5G15530", "AT5G16390", "AT5G35360", "AT3G56130", "AT2G30200", "AT1G62640", "AT1G74960", "AT2G04540", "AT5G46290", "AT1G24360", "AT2G22230", "AT5G10160", "AT2G05990")

length(Ara_fatty_acid_ID)

Ara_fatty_acid_ID %in% napus_vs_ara.final$ara_locus_ID # only one is not absent in my data 
Ara_fatty_acid_ID %in% tmp$ara_locus_ID # however, only 4 are differentially expressed 

```

# which stage has the most stress related genes expressed/ differentially expressed 
```{r}

# 1) get stress related GO terms 
# GO:0006950 response to stress 
# GO:0009414 response to water deprivation
  # drought recovery: GO:0009819 
  # cellular response to water deprivation: GO:0042631
  # response to dessication: GO:0009269
  # behavioral response to water deprivation: GO:0042630
  # regulation of response to water deprivation: GO:2000070 
  # negative regulation of response to water deprivation: GO:0080148  
  # positive regulation of response to water deprivation: GO:1902584 
  # cellular response to dessication: GO:0071465 
  # trehalose transport in response to water deprivation: GO:0072514	
  # trehalose transport in response to desiccation: GO:0072515 
  # trehalose metabolism in response to water deprivation: GO:0070416


### the number of drought related genes are very small, see whether I can get more genes by using Ara GO term 
# awk 'NR==FNR{a[$1]=$0}NR>FNR{if ($2 in a) print $0 "\t" a[$2]}' ../KIAT_RNA_seq/stress_related/ATH_GO_GOSLIM.txt napus_vs_ara.final.tmp | awk 'BEGIN{FS="\t"}{print $1 "\t" $2 "\t" $3 "\t" $4 "\t" $10}' > napus_vs_ara.with.GO.final.final 


# 2) extract genes with those GO terms
stress.related.geneID <- read.delim("~/Desktop/Brassica_project/KIAT_RNA_seq/stress_related/stress_related_geneID", header = F)
drought.stress.related.geneID <- read.delim("~/Desktop/Brassica_project/KIAT_RNA_seq/stress_related/response_to_desiccation", header = F)

# get drought related genes by using Arabidopsis blast hit 
drought.stress.related.geneID.2 <- read.delim("~/Desktop/Brassica_project/KIAT_RNA_seq/stress_related/drought_related_geneID", header = F)

drought.stress.related.geneID.3 <- read.delim("~/Desktop/Brassica_project/KIAT_RNA_seq/drought_related/drought_related_napus_ID", header = F)

length(unique(c(as.character(stress.related.geneID$V1), as.character(drought.stress.related.geneID.2)))) # 243 genes in total 

length(stress.related.geneID$V1) # 242 genes 
length(drought.stress.related.geneID.2$V1) # 13 genes
length(drought.stress.related.geneID.3$V1) # 13694 genes 

# 3) check expression pattern of these genes for all different comparisons
## All of the dataset that I need to check with 

### extract genes expression pattern 


# stress.related.geneID.between.gt <- lapply(list.between.gt, function(DEgene){
#   stress.related <- DEgene[rownames(DEgene) %in% stress.related.geneID$V1,]
# }
# )
# 
# # calculate 
# # total 
# lapply(stress.related.geneID.between.gt, function(DEgene){
#   dim(DEgene)
# })
# 
# # up 
# lapply(stress.related.geneID.between.gt, function(DEgene){
#   sum(DEgene$logFC > 0)
# })
#   
# # down 
# lapply(stress.related.geneID.between.gt, function(DEgene){
#   sum(DEgene$logFC < 0)
# })

drought.stress.related.geneID.between.gt <- lapply(list.between.gt, function(DEgene){
  drought.related <- DEgene[rownames(DEgene) %in% drought.stress.related.geneID.3$V1,]
}
)

# total 
lapply(drought.stress.related.geneID.between.gt, function(DEgene){
  total <- dim(DEgene) 
})

# up 
lapply(drought.stress.related.geneID.between.gt, function(DEgene){
  sum(DEgene$logFC > 0)
})
  
# down 
lapply(drought.stress.related.geneID.between.gt, function(DEgene){
  sum(DEgene$logFC < 0)
}) 

# drought.stress.related.geneID.2.between.gt <- lapply(list.between.gt, function(DEgene){
#   drought.related <- DEgene[rownames(DEgene) %in% drought.stress.related.geneID.2$V1,]
# }
# )
# 
# lapply(drought.stress.related.geneID.2.between.gt, function(DEgene){
#   total <- dim(DEgene) 
# })
# 
# # up 
# lapply(drought.stress.related.geneID.2.between.gt, function(DEgene){
#   sum(DEgene$logFC > 0)
# })
#   
# # down 
# lapply(drought.stress.related.geneID.2.between.gt, function(DEgene){
#   sum(DEgene$logFC < 0)
# }) 

# stress.related.geneID.between.stages.ae <- lapply(list.between.stages.ae, function(DEgene){
#   stress.related <- DEgene[rownames(DEgene) %in% stress.related.geneID$V1,]
# }
# )
# 
# # calculate 
# # total 
# lapply(stress.related.geneID.between.stages.ae, function(DEgene){
#   dim(DEgene)
# })
# 
# # up 
# lapply(stress.related.geneID.between.stages.ae, function(DEgene){
#   sum(DEgene$logFC > 0)
# })
#   
# # down 
# lapply(stress.related.geneID.between.stages.ae, function(DEgene){
#   sum(DEgene$logFC < 0)
# })

drought.related.geneID.between.stages.ae <- lapply(list.between.stages.ae, function(DEgene){
  drought.related <- DEgene[rownames(DEgene) %in% drought.stress.related.geneID.3$V1,]
}
)

# total 
lapply(drought.related.geneID.between.stages.ae, function(DEgene){
  dim(DEgene)
})

# up 
lapply(drought.related.geneID.between.stages.ae, function(DEgene){
  sum(DEgene$logFC > 0)
})
  
# down 
lapply(drought.related.geneID.between.stages.ae, function(DEgene){
  sum(DEgene$logFC < 0)
})   
# 
# drought.related.geneID.2.between.stages.ae <- lapply(list.between.stages.ae, function(DEgene){
#   drought.related <- DEgene[rownames(DEgene) %in% drought.stress.related.geneID.2$V1,]
# }
# )
# 

# 
# stress.related.geneID.between.stages.ol <- lapply(list.between.stages.ol, function(DEgene){
#   stress.related <- DEgene[rownames(DEgene) %in% stress.related.geneID$V1,]
# }
# )

# calculate 
# total 
# lapply(stress.related.geneID.between.stages.ol, function(DEgene){
#   dim(DEgene)
# })
# 
# # up 
# lapply(stress.related.geneID.between.stages.ol, function(DEgene){
#   sum(DEgene$logFC > 0)
# })
#   
# # down 
# lapply(stress.related.geneID.between.stages.ol, function(DEgene){
#   sum(DEgene$logFC < 0)
# })

drought.related.geneID.between.stages.ol <- lapply(list.between.stages.ol, function(DEgene){
  drought.related <- DEgene[rownames(DEgene) %in% drought.stress.related.geneID.3$V1,]
}
)

# total 
lapply(drought.related.geneID.between.stages.ol, function(DEgene){
  dim(DEgene)
})

# up 
lapply(drought.related.geneID.between.stages.ol, function(DEgene){
  sum(DEgene$logFC > 0)
})
  
# down 
lapply(drought.related.geneID.between.stages.ol, function(DEgene){
  sum(DEgene$logFC < 0)
})

# drought.related.geneID.2.between.stages.ol <- lapply(list.between.stages.ol, function(DEgene){
#   drought.related <- DEgene[rownames(DEgene) %in% drought.stress.related.geneID.2$V1,]
# }
# )

# 4) get output tables with best Arabidopsis hit annotated 
# stress.related.between.gt.w.anno <- 
# lapply(stress.related.geneID.between.gt, function(DEgene){
#   merge(DEgene, napus_vs_ara.final, by="row.names")
# })
# 
# stress.related.between.gt.w.anno[[1]]$class <- "young"
# stress.related.between.gt.w.anno[[2]]$class <- "bolting"
# stress.related.between.gt.w.anno[[3]]$class <- "flowering"
# stress.related.between.gt.w.anno[[4]]$class <- "early_silique"
# stress.related.between.gt.w.anno[[5]]$class <- "late_silique"
# 
# stress.related.between.gt.w.anno.bind <- rbind(stress.related.between.gt.w.anno[[1]], stress.related.between.gt.w.anno[[2]],stress.related.between.gt.w.anno[[3]], stress.related.between.gt.w.anno[[4]],stress.related.between.gt.w.anno[[5]])
# 
# write.csv(stress.related.between.gt.w.anno.bind, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/stress_related/stress.related.between.gt.w.anno.bind.csv")

# stress.related.geneID.between.stages.ae.w.anno <- 
#   lapply(stress.related.geneID.between.stages.ae, function(DEgene){
#   merge(DEgene, napus_vs_ara.final, by="row.names")
# }) 
# 
# stress.related.geneID.between.stages.ae.w.anno[[1]]$class <- "young_vs_bolting_ae"
# stress.related.geneID.between.stages.ae.w.anno[[2]]$class <- "bolting_vs_flowering_ae"
# stress.related.geneID.between.stages.ae.w.anno[[3]]$class <- "flowering_vs_earlysilique_ae"
# stress.related.geneID.between.stages.ae.w.anno[[4]]$class <- "earlysilique_vs_latesilique_ae"
# 
# stress.related.geneID.between.stages.ol.w.anno <- 
#   lapply(stress.related.geneID.between.stages.ol, function(DEgene){
#   merge(DEgene, napus_vs_ara.final, by="row.names")
# }) 
# 
# stress.related.geneID.between.stages.ol.w.anno[[1]]$class <- "young_vs_bolting_ol"
# stress.related.geneID.between.stages.ol.w.anno[[2]]$class <- "bolting_vs_flowering_ol"
# stress.related.geneID.between.stages.ol.w.anno[[3]]$class <- "flowering_vs_earlysilique_ol"
# stress.related.geneID.between.stages.ol.w.anno[[4]]$class <- "earlysilique_vs_latesilique_ol"
# 
# stress.related.geneID.between.stages.w.anno.bind <- rbind(stress.related.geneID.between.stages.ae.w.anno[[1]], stress.related.geneID.between.stages.ae.w.anno[[2]], stress.related.geneID.between.stages.ae.w.anno[[3]], stress.related.geneID.between.stages.ae.w.anno[[4]], stress.related.geneID.between.stages.ol.w.anno[[1]], stress.related.geneID.between.stages.ol.w.anno[[2]], stress.related.geneID.between.stages.ol.w.anno[[3]], stress.related.geneID.between.stages.ol.w.anno[[4]])  
# 
# write.csv(stress.related.geneID.between.stages.w.anno.bind, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/stress_related/stress.related.geneID.between.stages.w.anno.bind.csv")

# drought related 
drought.related.between.gt.w.anno <- 
lapply(drought.stress.related.geneID.between.gt, function(DEgene){
  merge(DEgene, napus_vs_ara.final, by="row.names")
})

drought.related.between.gt.w.anno[[1]]$class <- "young"
drought.related.between.gt.w.anno[[2]]$class <- "bolting"
drought.related.between.gt.w.anno[[3]]$class <- "flowering"
drought.related.between.gt.w.anno[[4]]$class <- "early_silique"
drought.related.between.gt.w.anno[[5]]$class <- "late_silique"

drought.related.between.gt.w.anno.bind <- rbind(drought.related.between.gt.w.anno[[1]], drought.related.between.gt.w.anno[[2]],drought.related.between.gt.w.anno[[3]], drought.related.between.gt.w.anno[[4]],drought.related.between.gt.w.anno[[5]])

write.csv(drought.related.between.gt.w.anno.bind, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/stress_related/drought.related.between.gt.w.anno.bind.csv")

drought.related.geneID.between.stages.ae.w.anno <- 
  lapply(drought.related.geneID.between.stages.ae, function(DEgene){
  merge(DEgene, napus_vs_ara.final, by="row.names")
}) 

drought.related.geneID.between.stages.ae.w.anno[[1]]$class <- "young_vs_bolting_ae"
drought.related.geneID.between.stages.ae.w.anno[[2]]$class <- "bolting_vs_flowering_ae"
drought.related.geneID.between.stages.ae.w.anno[[3]]$class <- "flowering_vs_earlysilique_ae"
drought.related.geneID.between.stages.ae.w.anno[[4]]$class <- "earlysilique_vs_latesilique_ae"

drought.related.geneID.between.stages.ol.w.anno <- 
  lapply(drought.related.geneID.between.stages.ol, function(DEgene){
  merge(DEgene, napus_vs_ara.final, by="row.names")
}) 

drought.related.geneID.between.stages.ol.w.anno[[1]]$class <- "young_vs_bolting_ol"
drought.related.geneID.between.stages.ol.w.anno[[2]]$class <- "bolting_vs_flowering_ol"
drought.related.geneID.between.stages.ol.w.anno[[3]]$class <- "flowering_vs_earlysilique_ol"
drought.related.geneID.between.stages.ol.w.anno[[4]]$class <- "earlysilique_vs_latesilique_ol"

drought.related.geneID.between.stages.w.anno.bind <- rbind(drought.related.geneID.between.stages.ae.w.anno[[1]], drought.related.geneID.between.stages.ae.w.anno[[2]], drought.related.geneID.between.stages.ae.w.anno[[3]], drought.related.geneID.between.stages.ae.w.anno[[4]], drought.related.geneID.between.stages.ol.w.anno[[1]], drought.related.geneID.between.stages.ol.w.anno[[2]], drought.related.geneID.between.stages.ol.w.anno[[3]], drought.related.geneID.between.stages.ol.w.anno[[4]])  

dim(drought.related.geneID.between.stages.w.anno.bind)

write.csv(drought.related.geneID.between.stages.w.anno.bind, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/stress_related/drought.related.geneID.between.stages.w.anno.bind.csv") 

```

# generate DEGs with Arabidopsis hit & up down regulated genes with GO enriched terms 
```{r}
# 5) Also make gene list with Arabidopsis hit from different comparisons tomorrow...
dim(DEgene.early.vs.late.silique.ol) # 2776 5 
dim(DEgene.early.vs.late.silique.ae) # 545 5 
dim(DEgene.late.silique) # 7413 5 
dim(DEgene.early.silique) # 6695 5 
rownames(napus_vs_ara.final) <- napus_vs_ara.final$napus_ID
colnames(napus_vs_ara.final)[2:4] <- c("Arabidopsis_ID", "blast_evalue", "description")

DEgene.early.vs.late.silique.ol.list <- (merge(DEgene.early.vs.late.silique.ol, napus_vs_ara.final, by="row.names"))[,-7]
rownames(DEgene.early.vs.late.silique.ol.list) <- DEgene.early.vs.late.silique.ol.list$Row.names
DEgene.early.vs.late.silique.ol.list <- DEgene.early.vs.late.silique.ol.list[,-1]
head(DEgene.early.vs.late.silique.ol.list)
write.csv(DEgene.early.vs.late.silique.ol.list, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/output/early.vs.late.silique.ol.list.csv")

DEgene.early.vs.late.silique.ae.list <- merge(DEgene.early.vs.late.silique.ae, napus_vs_ara.final, by="row.names")
rownames(DEgene.early.vs.late.silique.ae.list) <- DEgene.early.vs.late.silique.ae.list$Row.names
DEgene.early.vs.late.silique.ae.list <- DEgene.early.vs.late.silique.ae.list[,-1]
head(DEgene.early.vs.late.silique.ae.list)
write.csv(DEgene.early.vs.late.silique.ae.list, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/output/early.vs.late.silique.ae.list.csv")


DEgene.early.silique.list <- merge(DEgene.early.silique, napus_vs_ara.final, by="row.names")
rownames(DEgene.early.silique.list) <- DEgene.early.silique.list$Row.names
DEgene.early.silique.list <- DEgene.early.silique.list[,-1]
head(DEgene.early.silique.list)
write.csv(DEgene.early.silique.list, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/output/early.silique.list.csv")


DEgene.late.silique.list <- merge(DEgene.late.silique, napus_vs_ara.final, by="row.names")
rownames(DEgene.late.silique.list) <- DEgene.late.silique.list$Row.names
DEgene.late.silique.list <- DEgene.late.silique.list[,-1]
head(DEgene.late.silique.list)
write.csv(DEgene.late.silique.list, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/output/late.silique.list.csv")


## Go enrichment of up & down regulated genes separately 
DEgene.early.vs.late.silique.ae.up.GO.out <- GOseq.Bn.ORA(rownames(DEgene.early.vs.late.silique.ae[DEgene.early.vs.late.silique.ae$logFC > 0,]))
DEgene.early.vs.late.silique.ae.up.GO.out$class <- rep("early.vs.late.silique.ae.up", nrow(DEgene.early.vs.late.silique.ae.up.GO.out))


DEgene.early.vs.late.silique.ae.down.GO.out <- GOseq.Bn.ORA(rownames(DEgene.early.vs.late.silique.ae[DEgene.early.vs.late.silique.ae$logFC < 0,]))
DEgene.early.vs.late.silique.ae.down.GO.out$class <- rep("early.vs.late.silique.ae.down", nrow(DEgene.early.vs.late.silique.ae.down.GO.out))

DEgene.early.vs.late.silique.ol.up.GO.out <- GOseq.Bn.ORA(rownames(DEgene.early.vs.late.silique.ol[DEgene.early.vs.late.silique.ol$logFC > 0,]))
DEgene.early.vs.late.silique.ol.up.GO.out$class <- rep("early.vs.late.silique.ol.up", nrow(DEgene.early.vs.late.silique.ol.up.GO.out))

DEgene.early.vs.late.silique.ol.down.GO.out <- GOseq.Bn.ORA(rownames(DEgene.early.vs.late.silique.ol[DEgene.early.vs.late.silique.ol$logFC < 0,]))
DEgene.early.vs.late.silique.ol.down.GO.out$class <- rep("early.vs.late.silique.ol.down", nrow(DEgene.early.vs.late.silique.ol.down.GO.out))

DEgene.early.silique.up.GO.out <- GOseq.Bn.ORA(rownames(DEgene.early.silique[DEgene.early.silique$logFC > 0,])) # no enriched GO

DEgene.early.silique.down.GO.out <- GOseq.Bn.ORA(rownames(DEgene.early.silique[DEgene.early.silique$logFC < 0,]))
DEgene.early.silique.down.GO.out$class <- rep("early.silique.down", nrow(DEgene.early.silique.down.GO.out))

DEgene.late.silique.up.GO.out <- GOseq.Bn.ORA(rownames(DEgene.late.silique[DEgene.late.silique$logFC > 0,]))
DEgene.late.silique.up.GO.out$class <- rep("late.silique.up", nrow(DEgene.late.silique.up.GO.out)) 

DEgene.late.silique.down.GO.out <- GOseq.Bn.ORA(rownames(DEgene.late.silique[DEgene.late.silique$logFC < 0,])) # no enriched GO 

all.up.down.GO.out <- rbind(DEgene.early.vs.late.silique.ol.up.GO.out, DEgene.early.vs.late.silique.ol.down.GO.out, DEgene.early.vs.late.silique.ae.up.GO.out, DEgene.early.vs.late.silique.ae.down.GO.out, DEgene.late.silique.up.GO.out, DEgene.early.silique.down.GO.out)

write.csv(all.up.down.GO.out, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/output/up.down.GO.out.csv")

```

# fatty acid related analysis, display the result in mapman 
```{r}
# 1) get reciprocal hit for B.rapa gene in Arabidopsis 
# 1.1) blast B.napus against Arabidopsis get only the top hit (already have) 
# 1.2) blast Arabidopsis against B.napus get only the top hit 
# blastn -db /Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/Reference/B.napus/Brassica_napus.annotation_v5.cds.fa -query TAIR10_cds_20110103_representative_gene_model_updated -out Arabidopsis_vs_napus -outfmt 6 -max_target_seqs 1 
# 1.3) if the top hit of for the two files are the same, write out the result 

# 2) learn mapman
# 3) display the fatty acid genes in mapman 

```

## coverage test 
```{r}
######### all trimmed data 
flowering.read.count.all <- read.table("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/read_count/read.count.flowering.all.tsv", header = T, check.names = F)
rownames(flowering.read.count.all) <- flowering.read.count.all[,1]
flowering.read.count.all <- flowering.read.count.all[,-1]

# format data 
head(flowering.read.count.all)
dim(flowering.read.count.all) # 101040     6 
colnames(flowering.read.count.all)

# replace sample ID 
colnames(flowering.read.count.all) <- c("Da-Ol-1_flowering_3","Da-Ae_flowering_3","Da-Ae_flowering_1","Da-Ae_flowering_2","Da-Ol-1_flowering_1","Da-Ol-1_flowering_2")
head(flowering.read.count.all)

library(edgeR)
library(ggplot2)

# filter based on read count, assign group, normalize, design matrix, calculate dispersion   
# set up group 
flowering.read.count.all <- flowering.read.count.all[,colSums(flowering.read.count.all) > 1000000]  
dim(flowering.read.count.all) # 101040     6 
flowering.read.count.all.sample <-data.frame(file=colnames(flowering.read.count.all),
                             batch=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\5",colnames(flowering.read.count.all))),  
                             genotype=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\1",colnames(flowering.read.count.all))),	
                             stage=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\3",colnames(flowering.read.count.all))),	
                             group=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\1\\3",colnames(flowering.read.count.all)))
)

flowering.read.count.all.sample
ftable(flowering.read.count.all.sample,row.vars="stage",col.vars=c("batch","genotype"))

# filter based on read count 
flowering.read.count.all.small <- flowering.read.count.all[rowSums(flowering.read.count.all > 10) >= 3,]
dim(flowering.read.count.all.small) # 48968     6 

# 
# flowering.read.count.all.sample$genotype <- as.factor(flowering.read.count.all.sample$genotype)
# is.factor(flowering.read.count.all.sample$genotype)
# flowering.read.count.all.sample$genotype <- relevel(flowering.read.count.all.sample$genotype,ref="Da-Ol-1")

# normalize 
dge.data.flowering.all <- DGEList(counts=flowering.read.count.all.small, group=flowering.read.count.all.sample$group)
dim(dge.data.flowering.all)
dge.data.flowering.all <- calcNormFactors(dge.data.flowering.all, method = "TMM") 
dge.data.flowering.all$sample
hist(dge.data.flowering.all$samples$norm.factors)

# MDS to check sample seperation, also using dengrogram.  
plotMDS(dge.data.flowering.all, method = "bcv")

# pairwise comparison using GLM model 
# design matrix # w/o batch effect
design.flowering.all <- model.matrix(~0+group, data = flowering.read.count.all.sample)
dge.data.flowering.all <- estimateGLMCommonDisp(dge.data.flowering.all, design.flowering.all,verbose = TRUE) # Disp = 0.14725 , BCV = 0.3837
dge.data.flowering.all <- estimateGLMTrendedDisp(dge.data.flowering.all,design.flowering.all)
dge.data.flowering.all <- estimateGLMTagwiseDisp(dge.data.flowering.all,design.flowering.all)
plotBCV(dge.data.flowering.all)

## w/o batch effect
fit.flowering.all <- glmFit(dge.data.flowering.all, design.flowering.all)
# lrt.tmp <- glmLRT(fit.flowering.all, coef = "genotypeDa-Ae")
# topTags(lrt.tmp)
# summary(de.flowering.all.tmp <- decideTestsDGE(lrt.tmp, p=0.05))

colnames(fit.flowering.all$coefficients)
lrt.flowering.all <- glmLRT(fit.flowering.all, contrast = c(1, -1))
topTags(lrt.flowering.all)
number.all <- summary(de.flowering.all <- decideTestsDGE(lrt.flowering.all, p=0.05))
DEgene.flowering.all <- topTags(lrt.flowering.all,n = Inf)$table[topTags(lrt.flowering.all,n = Inf)$table$FDR<0.05,]

dim(DEgene.flowering.all)
head(DEgene.flowering.all)

##### half of the reads 
flowering.read.count.half <- read.table("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/read_count/read.count.flowering.half.tsv", header = T, check.names = F)
rownames(flowering.read.count.half) <- flowering.read.count.half[,1]
flowering.read.count.half <- flowering.read.count.half[,-1]

# format data 
head(flowering.read.count.half)
dim(flowering.read.count.half) # 101040     6 
colnames(flowering.read.count.half)

# replace sample ID 
colnames(flowering.read.count.half) <- c("Da-Ol-1_flowering_3","Da-Ae_flowering_3","Da-Ae_flowering_1","Da-Ae_flowering_2","Da-Ol-1_flowering_1","Da-Ol-1_flowering_2")
head(flowering.read.count.half)

# filter based on read count, assign group, normalize, design matrix, calculate dispersion   
# set up group 
flowering.read.count.half <- flowering.read.count.half[,colSums(flowering.read.count.half) > 1000000]  
dim(flowering.read.count.half) # 101040     6 
flowering.read.count.half.sample <-data.frame(file=colnames(flowering.read.count.half),
                             batch=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\5",colnames(flowering.read.count.half))),  
                             genotype=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\1",colnames(flowering.read.count.half))),	
                             stage=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\3",colnames(flowering.read.count.half))),	
                             group=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\1\\3",colnames(flowering.read.count.half)))
)

flowering.read.count.half.sample
ftable(flowering.read.count.half.sample,row.vars="stage",col.vars=c("batch","genotype"))

# filter based on read count 
flowering.read.count.half.small <- flowering.read.count.half[rowSums(flowering.read.count.half > 10) >= 3,]
dim(flowering.read.count.half.small) # 41412     6 

# normalize 
dge.data.flowering.half <- DGEList(counts=flowering.read.count.half.small, group=flowering.read.count.half.sample$group)
dim(dge.data.flowering.half)
dge.data.flowering.half <- calcNormFactors(dge.data.flowering.half, method = "TMM") 
dge.data.flowering.half$sample
hist(dge.data.flowering.half$samples$norm.factors)

# MDS to check sample seperation, also using dengrogram.  
plotMDS(dge.data.flowering.half, method = "bcv")

# pairwise comparison using GLM model 
# design matrix # w/o batch effect
design.flowering.half <- model.matrix(~0+group, data = flowering.read.count.half.sample)
dge.data.flowering.half <- estimateGLMCommonDisp(dge.data.flowering.half, design.flowering.half,verbose = TRUE) # Disp = 0.14364 , BCV = 0.379 
dge.data.flowering.half <- estimateGLMTrendedDisp(dge.data.flowering.half,design.flowering.half)
dge.data.flowering.half <- estimateGLMTagwiseDisp(dge.data.flowering.half,design.flowering.half)
plotBCV(dge.data.flowering.half)

## w/o batch effect
fit.flowering.half <- glmFit(dge.data.flowering.half, design.flowering.half)

colnames(fit.flowering.half$coefficients)
lrt.flowering.half <- glmLRT(fit.flowering.half, contrast = c(1, -1))
topTags(lrt.flowering.half) 
number.half <- summary(de.flowering.half <- decideTestsDGE(lrt.flowering.half, p=0.05))
DEgene.flowering.half <- topTags(lrt.flowering.half,n = Inf)$table[topTags(lrt.flowering.half,n = Inf)$table$FDR<0.05,]

dim(DEgene.flowering.half)
head(DEgene.flowering.half) 

##### 1/4 of the reads 
flowering.read.count.halfhalf <- read.table("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/read_count/read.count.flowering.halfhalf.tsv", header = T, check.names = F)
rownames(flowering.read.count.halfhalf) <- flowering.read.count.halfhalf[,1]
flowering.read.count.halfhalf <- flowering.read.count.halfhalf[,-1]

# format data 
head(flowering.read.count.halfhalf)
dim(flowering.read.count.halfhalf) # 101040     6 
colnames(flowering.read.count.halfhalf)

# replace sample ID 
colnames(flowering.read.count.halfhalf) <- c("Da-Ol-1_flowering_3","Da-Ae_flowering_3","Da-Ae_flowering_1","Da-Ae_flowering_2","Da-Ol-1_flowering_1","Da-Ol-1_flowering_2")
head(flowering.read.count.halfhalf)

# filter based on read count, assign group, normalize, design matrix, calculate dispersion   
# set up group 
flowering.read.count.halfhalf <- flowering.read.count.halfhalf[,colSums(flowering.read.count.halfhalf) > 1000000]  
dim(flowering.read.count.halfhalf) # 101040     6 
flowering.read.count.halfhalf.sample <-data.frame(file=colnames(flowering.read.count.halfhalf),
                             batch=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\5",colnames(flowering.read.count.halfhalf))),  
                             genotype=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\1",colnames(flowering.read.count.halfhalf))),	
                             stage=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\3",colnames(flowering.read.count.halfhalf))),	
                             group=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\1\\3",colnames(flowering.read.count.halfhalf)))
)

flowering.read.count.halfhalf.sample
ftable(flowering.read.count.halfhalf.sample,row.vars="stage",col.vars=c("batch","genotype"))

# filter based on read count 
flowering.read.count.halfhalf.small <- flowering.read.count.halfhalf[rowSums(flowering.read.count.halfhalf > 10) >= 3,]
dim(flowering.read.count.halfhalf.small) # 32179     6 

# normalize 
dge.data.flowering.halfhalf <- DGEList(counts=flowering.read.count.halfhalf.small, group=flowering.read.count.halfhalf.sample$group)
dim(dge.data.flowering.halfhalf)
dge.data.flowering.halfhalf <- calcNormFactors(dge.data.flowering.halfhalf, method = "TMM") 
dge.data.flowering.halfhalf$sample
hist(dge.data.flowering.halfhalf$samples$norm.factors)

# MDS to check sample seperation, also using dengrogram.  
plotMDS(dge.data.flowering.halfhalf, method = "bcv")

# pairwise comparison using GLM model 
# design matrix # w/o batch effect
design.flowering.halfhalf <- model.matrix(~0+group, data = flowering.read.count.halfhalf.sample)
dge.data.flowering.halfhalf <- estimateGLMCommonDisp(dge.data.flowering.halfhalf, design.flowering.halfhalf,verbose = TRUE) # Disp = 0.13675 , BCV = 0.3698
dge.data.flowering.halfhalf <- estimateGLMTrendedDisp(dge.data.flowering.halfhalf,design.flowering.halfhalf)
dge.data.flowering.halfhalf <- estimateGLMTagwiseDisp(dge.data.flowering.halfhalf,design.flowering.halfhalf)
plotBCV(dge.data.flowering.halfhalf)

## w/o batch effect
fit.flowering.halfhalf <- glmFit(dge.data.flowering.halfhalf, design.flowering.halfhalf)

colnames(fit.flowering.halfhalf$coefficients)
lrt.flowering.halfhalf <- glmLRT(fit.flowering.halfhalf, contrast = c(1, -1))
topTags(lrt.flowering.halfhalf) 
number.halfhalf <- summary(de.flowering.halfhalf <- decideTestsDGE(lrt.flowering.halfhalf, p=0.05))
DEgene.flowering.halfhalf <- topTags(lrt.flowering.halfhalf,n = Inf)$table[topTags(lrt.flowering.halfhalf,n = Inf)$table$FDR<0.05,]

dim(DEgene.flowering.halfhalf)
head(DEgene.flowering.halfhalf) 

### 3/4 
##### 1/4 of the reads 
flowering.read.count.3_4 <- read.table("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/read_count/read.count.flowering.3_4.tsv", header = T, check.names = F)
rownames(flowering.read.count.3_4) <- flowering.read.count.3_4[,1]
flowering.read.count.3_4 <- flowering.read.count.3_4[,-1]

# format data 
head(flowering.read.count.3_4)
dim(flowering.read.count.3_4) # 101040     6 
colnames(flowering.read.count.3_4)

# replace sample ID 
colnames(flowering.read.count.3_4) <- c("Da-Ol-1_flowering_3","Da-Ae_flowering_3","Da-Ae_flowering_1","Da-Ae_flowering_2","Da-Ol-1_flowering_1","Da-Ol-1_flowering_2")
head(flowering.read.count.3_4)

# filter based on read count, assign group, normalize, design matrix, calculate dispersion   
# set up group 
flowering.read.count.3_4 <- flowering.read.count.3_4[,colSums(flowering.read.count.3_4) > 1000000]  
dim(flowering.read.count.3_4) # 101040     6 
flowering.read.count.3_4.sample <-data.frame(file=colnames(flowering.read.count.3_4),
                             batch=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\5",colnames(flowering.read.count.3_4))),  
                             genotype=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\1",colnames(flowering.read.count.3_4))),	
                             stage=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\3",colnames(flowering.read.count.3_4))),	
                             group=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)(1|2|3)","\\1\\3",colnames(flowering.read.count.3_4)))
)

flowering.read.count.3_4.sample
ftable(flowering.read.count.3_4.sample,row.vars="stage",col.vars=c("batch","genotype"))

# filter based on read count 
flowering.read.count.3_4.small <- flowering.read.count.3_4[rowSums(flowering.read.count.3_4 > 10) >= 3,]
dim(flowering.read.count.3_4.small)  

# normalize 
dge.data.flowering.3_4 <- DGEList(counts=flowering.read.count.3_4.small, group=flowering.read.count.3_4.sample$group)
dim(dge.data.flowering.3_4)
dge.data.flowering.3_4 <- calcNormFactors(dge.data.flowering.3_4, method = "TMM") 
dge.data.flowering.3_4$sample
hist(dge.data.flowering.3_4$samples$norm.factors)

# MDS to check sample seperation, also using dengrogram.  
plotMDS(dge.data.flowering.3_4, method = "bcv")

# pairwise comparison using GLM model 
# design matrix # w/o batch effect
design.flowering.3_4 <- model.matrix(~0+group, data = flowering.read.count.3_4.sample)
dge.data.flowering.3_4 <- estimateGLMCommonDisp(dge.data.flowering.3_4, design.flowering.3_4,verbose = TRUE) # Disp = 0.15922 , BCV = 0.399 
dge.data.flowering.3_4 <- estimateGLMTrendedDisp(dge.data.flowering.3_4,design.flowering.3_4)
dge.data.flowering.3_4 <- estimateGLMTagwiseDisp(dge.data.flowering.3_4,design.flowering.3_4)
plotBCV(dge.data.flowering.3_4)

## w/o batch effect
fit.flowering.3_4 <- glmFit(dge.data.flowering.3_4, design.flowering.3_4)

colnames(fit.flowering.3_4$coefficients)
lrt.flowering.3_4 <- glmLRT(fit.flowering.3_4, contrast = c(1, -1))
topTags(lrt.flowering.3_4) 
number.3_4 <- summary(de.flowering.3_4 <- decideTestsDGE(lrt.flowering.3_4, p=0.05))
DEgene.flowering.3_4 <- topTags(lrt.flowering.3_4,n = Inf)$table[topTags(lrt.flowering.3_4,n = Inf)$table$FDR<0.05,]

dim(DEgene.flowering.3_4)
head(DEgene.flowering.3_4)  

################################# 

# make graphs of results 
number.all; number.3_4; number.half; number.halfhalf
number.all[[1]]


DE.table.all <- data.frame(up = c(number.all[[3]], number.3_4[[3]], number.half[[3]], number.halfhalf[[3]]), 
                       down = c(number.all[[1]], number.3_4[[1]], number.half[[1]], number.halfhalf[[1]]),
                       total = c(sum(number.all[[1]], number.all[[3]]), sum(number.3_4[[1]], number.3_4[[3]]), sum(number.half[[1]], number.half[[3]]), sum(number.halfhalf[[1]], number.halfhalf[[3]])))

rownames(DE.table.all) <- c("all","threequarter","half","halfhalf")
DE.table.all 

# reshape for ggplot 
DE.table.all.melt <- melt(DE.table.all) 
DE.table.all.melt$type <- rep(c("all","threequarter","half","halfhalf"), 3)
DE.table.all.melt
DE.table.all.melt$type <- factor(DE.table.all.melt$type, levels = DE.table.all.melt$type)
DE.table.all.melt$label <- factor(DE.table.all.melt$value)

p.DE.all <- ggplot(data = DE.table.all.melt)
p.DE.all <- p.DE.all + geom_line(aes(x = factor(type), y = value, group=variable, color=variable))
p.DE.all <- p.DE.all + ggplot(data = DE.table.all.melt, aes(x=factor(type), y=values*1.05, label= DE.table.all.melt$label))  
p.DE.all <- p.DE.all + theme(axis.text.x=element_text(angle=90),strip.text.y = element_text(angle=0))
p.DE.all <- p.DE.all + labs(y = "number of differentially expressed genes", x = "coverage type")

p.DE.all
ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/DEGs_number_between_diff_coverage_level.png", width = 11, height = 8)  

# number of overlap genes
head(DEgene.flowering.halfhalf)
dim(DEgene.flowering.all); dim(DEgene.flowering.3_4); dim(DEgene.flowering.half); dim(DEgene.flowering.halfhalf)
number.all; number.half; number.halfhalf 

dim(DEgene.flowering.all[rownames(DEgene.flowering.half),]) 
dim(DEgene.flowering.half[rownames(DEgene.flowering.halfhalf),])
dim(DEgene.flowering.halfhalf) 

# install.packages("gplots")
library(gplots)

# all genes overlapped 
DE.flowering.all <- rownames(DEgene.flowering.all)
DE.flowering.3_4 <- rownames(DEgene.flowering.3_4)
DE.flowering.half <- rownames(DEgene.flowering.half)
DE.flowering.halfhalf <- rownames(DEgene.flowering.halfhalf)

png(filename = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/venn.between.diff.coverage.all.png")
venn(list(all=DE.flowering.all, threequarter=DE.flowering.3_4, half=DE.flowering.half, halfhalf=DE.flowering.halfhalf))
dev.off()

# up regulated genes 
DE.flowering.all.up <- rownames(DEgene.flowering.all[DEgene.flowering.all$logFC>=0,])
DE.flowering.3_4.up <- rownames(DEgene.flowering.3_4[DEgene.flowering.3_4$logFC>=0,])
DE.flowering.half.up <- rownames(DEgene.flowering.half[DEgene.flowering.half$logFC>=0,])
DE.flowering.halfhalf.up <- rownames(DEgene.flowering.halfhalf[DEgene.flowering.halfhalf$logFC>=0,])

png(filename = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/venn.between.diff.coverage.up.png")
venn(list(all=DE.flowering.all.up, threequarter=DE.flowering.3_4.up, half=DE.flowering.half.up, halfhalf=DE.flowering.halfhalf.up))
dev.off()

# down regulated genes 
DE.flowering.all.down <- rownames(DEgene.flowering.all[DEgene.flowering.all$logFC<0,])
DE.flowering.3_4.down <- rownames(DEgene.flowering.3_4[DEgene.flowering.3_4$logFC<0,])
DE.flowering.half.down <- rownames(DEgene.flowering.half[DEgene.flowering.half$logFC<0,])
DE.flowering.halfhalf.down <- rownames(DEgene.flowering.halfhalf[DEgene.flowering.halfhalf$logFC<0,])

png(filename = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/venn.between.diff.coverage.down.png")
venn(list(all=DE.flowering.all.down, threequarter=DE.flowering.3_4.down, half=DE.flowering.half.down, halfhalf=DE.flowering.halfhalf.down))
dev.off()
```

##################################################################
# anaylsis using genome mapped reads 
##################################################################
# Expression analysis using genome mapped reads 1) formatting data    
```{r}
parent.read.count.2 <- read.table("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/read_count/read.count.genome.mapped.tsv", header = T, check.names = F)
rownames(parent.read.count.2) <- parent.read.count.2[,1]
parent.read.count.2 <- parent.read.count.2[,-1]

# format data 
head(parent.read.count.2)
dim(parent.read.count.2) # 101040     27  
colnames(parent.read.count.2)
colnames(parent.read.count.2) <- sub("_paired","",colnames(parent.read.count.2),fixed = TRUE) # remove _paired.fq 
colnames(parent.read.count.2)

# sample description  
sample_des <- read.csv("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/parent_summary_corrected.csv")
dim(sample_des) # 27 22
sorted_sample_des <- sample_des[order(sample_des$SampleID),]
sorted_sample_des$SampleID <- as.character(sorted_sample_des$SampleID)
sorted_sample_des$SampleID

sorted_sample_des[,4:7]
new_sample_ID <- paste(sorted_sample_des$Cultivar, sorted_sample_des$Stage, sorted_sample_des$rep, sep = "_")
new_sample_ID

# replace sample ID (if the sorted sample ID is the same as colname of read count file, replace colname of read count file with new sample ID)
if (identical(colnames(parent.read.count.2), sorted_sample_des$SampleID) == TRUE){
  colnames(parent.read.count.2) <- new_sample_ID 
}

head(parent.read.count.2)

### remove bolting data 
# parent.read.count.2 <- parent.read.count.2[,grep("bolting", colnames(parent.read.count.2), invert = TRUE)]
dim(parent.read.count.2) # [1] 101040     24
# save(parent.read.count.2,file="/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.2.Rdata") 
```

# expression analysis using genome mapped reads 2) set up sample description 
```{r}
# load necessary libs & functions 
library(edgeR)
library(ggplot2)

# filter based on read count, assign group, normalize, design matrix, calculate dispersion   
# set up group 
# load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.2.Rdata")
parent.read.count.2 <- parent.read.count.2[,colSums(parent.read.count.2) > 1000000]  
dim(parent.read.count.2) # 101040     24 
parent.read.count.2.sample<-data.frame(file=colnames(parent.read.count.2),
                             batch=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)([[:digit:]])","\\5",colnames(parent.read.count.2))),  
                             genotype=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)([[:digit:]])","\\1",colnames(parent.read.count.2))),	
                             tissue=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)([[:digit:]])","\\3",colnames(parent.read.count.2))),	
                             group=factor(gsub("([[:print:]]+)(_)([[:print:]]+)(_)([[:digit:]])","\\1\\3",colnames(parent.read.count.2)))
)

parent.read.count.2.sample
ftable(parent.read.count.2.sample,row.vars="tissue",col.vars=c("batch","genotype"))

# filter based on read count 
parent.read.count.2.small <- parent.read.count.2[rowSums(parent.read.count.2 > 10) >= 3,]
dim(parent.read.count.2.small) # 57925    24   # 58486    27 
# save(parent.read.count.2.small, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.2.small.Rdata")
```

# expression analysis using genome mapped reads 3) voom transformation for clustering analysis 
```{r}
# voom transformation 
# source("https://bioconductor.org/biocLite.R")
# biocLite("DESeq2")
# load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.2.small.Rdata")
library("DESeq2")

dds.parent.2 <- DESeqDataSetFromMatrix(countData = round(parent.read.count.2.small), colData = parent.read.count.2.sample, design = ~ batch + genotype*tissue)

vsd.parent.2 <- varianceStabilizingTransformation(dds.parent.2)
vstMat.parent.2 <- assay(vsd.parent.2)
colnames(vstMat.parent.2) <- colnames(parent.read.count.2)
# save(vstMat.parent.2, file = "/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/vstMat.parent.2.Rdata")
```

# expression analysis using genome mapped reads 4) MDS and hierachical clustering  
```{r}
# load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/vstMat.parent.2.Rdata")
# load("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/data/parent.read.count.2.small.Rdata")

# normalize 
dge.data.parent.2 <- DGEList(counts=parent.read.count.2.small, group=parent.read.count.2.sample$group)
dge.data.parent.2 <- calcNormFactors(dge.data.parent.2, method = "TMM") 
dge.data.parent.2$sample
hist(dge.data.parent.2$samples$norm.factors)

# MDS to check sample seperation, also using dengrogram.  
mds.2 <- plotMDS(dge.data.parent.2, method = "bcv",labels = dge.data.parent.2$samples$group)

x.2 <- as.data.frame(mds.2$x)
y.2 <- as.data.frame(mds.2$y)

distance_matrix.2 <- merge(x.2, y.2, by="row.names")

distance_matrix.2$group <- gsub("(Da-Ae|Da-Ol-1)(_)([[:print:]]+)(_)(1|2|3)","\\1\\3",distance_matrix.2$Row.names) 
distance_matrix.2$gt <- gsub("(Da-Ae|Da-Ol-1)(_)([[:print:]]+)(_)(1|2|3)","\\1",distance_matrix.2$Row.names)
distance_matrix.2$tissue <- gsub("(Da-Ae|Da-Ol-1)(_)([[:print:]]+)(_)(1|2|3)","\\3",distance_matrix.2$Row.names)

colnames(distance_matrix.2) <- c("lib","x","y","group","gt","tissue")
head(distance_matrix.2)

# making color MDS figure 
p.mds.2 <- ggplot(data = distance_matrix.2) + theme_grey() 
p.mds.2 <- p.mds.2 + geom_point(aes(x, y, color=factor(gt)), size=3) 
p.mds.2 <- p.mds.2 + labs(y = "BCV distance 2", x="BCV distance 1")
# p.mds.2 <- p.mds.2 + facet_grid(~gt)
p.mds.2     

ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/genome_mapped/MDS_nobolting.by.gt.png", width = 11, height = 8)

# clustering to check sample seperation 
library(ggdendro)

hc.parent.2 <- hclust(dist(t(vstMat.parent.2)))

ggdata.parent.2 <- dendro_data(as.dendrogram(hc.parent.2))
ggdata.parent.2$labels$gt <- gsub("(Da-Ae|Da-Ol-1)(_)(Young|flowering|early-silique|late-silique)(_)(1|2|3)","\\1",ggdata.parent.2$labels$label) 
ggdata.parent.2$labels$tissue <- gsub("(Da-Ae|Da-Ol-1)(_)(Young|flowering|early-silique|late-silique|bolting)(_)(1|2|3)","\\3",ggdata.parent.2$labels$label)
ggdata.parent.2$labels$group <- paste(ggdata.parent.2$labels$gt, ggdata.parent.2$labels$tissue, sep = "_")
ggdata.parent.2$labels

# start ggplot here 
p.dengro.2 <- ggplot(data = segment(ggdata.parent.2))
p.dendro.2 <- p.dengro.2 + geom_segment(aes(x=x, y=y, xend=xend, yend=yend)) + theme_dendro()
p.dendro.2 <- p.dendro.2 + geom_text(data = label(ggdata.parent.2), aes(x = x, y = y, label = label, hjust=0, color=group)) + coord_flip() + scale_y_reverse(expand=c(0.2, 0))
p.dendro.2
ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/genome_mapped/clustering_nobolting.png", width = 15, height = 8) 
```

# expression analysis using genome mapped reads 5) pairwise expression analysis 
```{r}
####
# pairwise comparison using GLM model  
# design matrix
design.parent.2 <- model.matrix(~0+group, data = parent.read.count.2.sample)

# w/o batch effect
dge.data.parent.2 <- estimateGLMCommonDisp(dge.data.parent.2, design.parent.2,verbose = TRUE) # Disp = 0.27855 , BCV = 0.5278 # Disp = 0.26782 , BCV = 0.5175 (w/ bolting) 
dge.data.parent.2 <- estimateGLMTrendedDisp(dge.data.parent.2,design.parent.2)
dge.data.parent.2 <- estimateGLMTagwiseDisp(dge.data.parent.2,design.parent.2)
plotBCV(dge.data.parent.2)

## w/o batch effect
fit.parent.2 <- glmFit(dge.data.parent.2, design.parent.2)

lrt.young.2 <- glmLRT(fit.parent.2, contrast = c(0,0,0,0,1,0,0,0,0,-1))
topTags(lrt.young.2)
summary(de.young.2 <- decideTestsDGE(lrt.young.2, p=0.05))
DEgene.young.2 <- topTags(lrt.young.2,n = Inf)$table[topTags(lrt.young.2,n = Inf)$table$FDR<0.05,]
# -1  3319
# 0  50820 
# 1   3786

# -1  3420
# 0  51115
# 1   3951

lrt.bolting.2 <- glmLRT(fit.parent.2, contrast = c(1,0,0,0,0,-1,0,0,0,0))
topTags(lrt.bolting.2)
summary(de.bolting.2 <- decideTestsDGE(lrt.bolting.2, p=0.05))
DEgene.bolting.2 <- topTags(lrt.bolting.2,n = Inf)$table[topTags(lrt.bolting.2,n = Inf)$table$FDR<0.05,]

# -1  2795
# 0  54153
# 1   1538

lrt.flowering.2 <- glmLRT(fit.parent.2, contrast = c(0,0,1,0,0,0,0,-1,0,0))
topTags(lrt.flowering.2)
summary(de.flowering.2 <- decideTestsDGE(lrt.flowering.2, p=0.05))
DEgene.flowering.2 <- topTags(lrt.flowering.2,n = Inf)$table[topTags(lrt.flowering.2,n = Inf)$table$FDR<0.05,]
# -1  2180
# 0  53340
# 1   2405
# 
# -1  2245
# 0  53752
# 1   2489

lrt.early.silique.2 <- glmLRT(fit.parent.2, contrast = c(0,1,0,0,0,0,-1,0,0,0))
topTags(lrt.early.silique.2)
summary(de.early.silique.2 <- decideTestsDGE(lrt.early.silique.2, p=0.05))
DEgene.early.silique <- topTags(lrt.early.silique,n = Inf)$table[topTags(lrt.early.silique,n = Inf)$table$FDR<0.05,]
# -1  2510
# 0  52799
# 1   2616
# 
# -1  2594
# 0  53194
# 1   2698 

lrt.late.silique.2 <- glmLRT(fit.parent.2, contrast = c(0,0,0,1,0,0,0,0,-1,0))
topTags(lrt.late.silique.2)
summary(de.late.silique.2 <- decideTestsDGE(lrt.late.silique.2, p=0.05))
DEgene.late.silique.2 <- topTags(lrt.late.silique.2,n = Inf)$table[topTags(lrt.late.silique.2,n = Inf)$table$FDR<0.05,]
# -1  2865
# 0  52077
# 1   2983 

# -1  2970
# 0  52462
# 1   3054 

### barplot for DEGs number 
### reformat for ggplot barplot 
DEGs_number_between_gt.parent.2 <- data.frame(young = as.data.frame(summary(de.young.2 <- decideTestsDGE(lrt.young.2, p=0.05)))$Freq,
                                              bolting = as.data.frame(summary(de.bolting.2 <- decideTestsDGE(lrt.bolting.2, p=0.05)))$Freq,
                                     flowering = as.data.frame(summary(de.flowering.2 <- decideTestsDGE(lrt.flowering.2, p=0.05)))$Freq, 
                                     early.silique = as.data.frame(summary(de.early.silique.2 <- decideTestsDGE(lrt.early.silique.2, p=0.05)))$Freq,
                                     late.silique = as.data.frame(summary(de.late.silique.2 <- decideTestsDGE(lrt.late.silique.2, p=0.05)))$Freq)

rownames(DEGs_number_between_gt.parent.2) <- c("down", "no", "up")
DEGs_number_between_gt.parent.2 <- DEGs_number_between_gt.parent.2[c("down", "up"),]
DEGs_number_between_gt.parent.2

library(reshape2)
DEGs_number_between_gt.melt.parent.2 <- melt(DEGs_number_between_gt.parent.2)
DEGs_number_between_gt.melt.parent.2$DE <- rep(c("down", "up"), 5)
colnames(DEGs_number_between_gt.melt.parent.2) <- c("tissue", "number", "DE")
DEGs_number_between_gt.melt.parent.2

# reorder: up 1st down 2nd 
DEGs_number_between_gt.melt.parent.2$DE <- factor(DEGs_number_between_gt.melt.parent.2$DE, levels = c("up", "down"))

DEGs_number_between_gt.melt.parent.2 <- DEGs_number_between_gt.melt.parent.2[order(DEGs_number_between_gt.melt.parent.2$DE),]
DEGs_number_between_gt.melt.parent.2

### making ggplot for DEGs 
p.DEGs_number_between_gt.parent.2 <- ggplot(data = DEGs_number_between_gt.melt.parent.2) + theme_grey()
p.DEGs_number_between_gt.parent.2 <- p.DEGs_number_between_gt.parent.2 + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number), stat = "identity")
p.DEGs_number_between_gt.parent.2 <- p.DEGs_number_between_gt.parent.2 + facet_grid(~tissue)
p.DEGs_number_between_gt.parent.2 <- p.DEGs_number_between_gt.parent.2 + geom_text(data=DEGs_number_between_gt.melt.parent.2,aes(x=DE,y=number*1.05,label=factor(number),color=DE)) 
p.DEGs_number_between_gt.parent.2 <- p.DEGs_number_between_gt.parent.2 + labs(y = "number of differentially expressed genes", x = "")

p.DEGs_number_between_gt.parent.2
ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/genome_mapped/DEGs_number_between_gt.parent.bolting.png", width = 11, height = 8)

## between developmental stages 
# Da-Ae 
lrt.young.vs.bolting.ae.2 <- glmLRT(fit.parent.2, contrast = c(1,0,0,0,-1,0,0,0,0,0))
topTags(lrt.young.vs.bolting.ae.2)
summary(de.young.vs.bolting.ae.2 <- decideTestsDGE(lrt.young.vs.bolting.ae.2, p=0.05))
DEgene.young.vs.bolting.ae.2 <- topTags(lrt.young.vs.flowering.ae.2,n = Inf)$table[topTags(lrt.young.vs.flowering.ae.2,n = Inf)$table$FDR<0.05,]
# -1   442
# 0  59872
# 1    661
# 
# -1   385
# 0  57369
# 1    732 

lrt.bolting.vs.flowering.ae.2 <- glmLRT(fit.parent.2, contrast = c(-1,0,1,0,0,0,0,0,0,0))
topTags(lrt.bolting.vs.flowering.ae.2)
summary(de.bolting.vs.flowering.ae.2 <- decideTestsDGE(lrt.bolting.vs.flowering.ae.2, p=0.05))
DEgene.bolting.vs.flowering.ae.2 <- topTags(lrt.bolting.vs.flowering.ae.2,n = Inf)$table[topTags(lrt.bolting.vs.flowering.ae.2,n = Inf)$table$FDR<0.05,]
# -1   763
# 0  54039
# 1   3684 

lrt.flowering.vs.early.silique.ae.2 <- glmLRT(fit.parent.2, contrast = c(0,1,-1,0,0,0,0,0,0,0))
topTags(lrt.flowering.vs.early.silique.ae.2)
summary(de.flowering.vs.early.silique.ae.2 <- decideTestsDGE(lrt.flowering.vs.early.silique.ae.2, p=0.05))
DEgene.flowering.vs.early.silique.ae.2 <- topTags(lrt.flowering.vs.early.silique.ae.2,n = Inf)$table[topTags(lrt.flowering.vs.early.silique.ae.2,n = Inf)$table$FDR<0.05,]
# -1  4696
# 0  52202
# 1   1588

lrt.early.vs.late.silique.ae.2 <- glmLRT(fit.parent.2, contrast = c(0,-1,0,1,0,0,0,0,0,0))
topTags(lrt.early.vs.late.silique.ae.2)
summary(de.early.vs.late.silique.ae.2 <- decideTestsDGE(lrt.early.vs.late.silique.ae.2, p=0.05))
DEgene.early.vs.late.silique.ae.2 <- topTags(lrt.early.vs.late.silique.ae.2,n = Inf)$table[topTags(lrt.early.vs.late.silique.ae.2,n = Inf)$table$FDR<0.05,]
# -1   416
# 0  57721 
# 1    349

### ggplot 
### reformat for ggplot barplot 
DEGs_number_between_tissue.ae.2 <- data.frame("bolting-vs-young" = as.data.frame(summary(de.young.vs.bolting.2.ae <- decideTestsDGE(lrt.young.vs.bolting.ae.2, p=0.05)))$Freq, 
                                            "flowering-vs-bolting" = as.data.frame(summary(de.bolting.vs.flowering.ae.2 <- decideTestsDGE(lrt.bolting.vs.flowering.ae.2, p=0.05)))$Freq,
                                     "early.slique-vs-flowering"= as.data.frame(summary(de.flowering.vs.early.silique.ae.2 <- decideTestsDGE(lrt.flowering.vs.early.silique.ae.2, p=0.05)))$Freq, 
                                     "late.silique-vs-early.silique" = as.data.frame(summary(de.early.vs.late.silique.ae.2 <- decideTestsDGE(lrt.early.vs.late.silique.ae.2, p=0.05)))$Freq)

rownames(DEGs_number_between_tissue.ae.2) <- c("down", "no", "up")
DEGs_number_between_tissue.ae.2 <- DEGs_number_between_tissue.ae.2[c("down", "up"),]
DEGs_number_between_tissue.ae.2

DEGs_number_between_tissue.ae.melt.parent.2 <- melt(DEGs_number_between_tissue.ae.2)
DEGs_number_between_tissue.ae.melt.parent.2$DE <- rep(c("down", "up"), 4)
colnames(DEGs_number_between_tissue.ae.melt.parent.2) <- c("tissue", "number", "DE")
DEGs_number_between_tissue.ae.melt.parent

# reorder: up 1st down 2nd 
DEGs_number_between_tissue.ae.melt.parent.2$DE <- factor(DEGs_number_between_tissue.ae.melt.parent.2$DE, levels = c("up", "down"))

DEGs_number_between_tissue.ae.melt.parent.2 <- DEGs_number_between_tissue.ae.melt.parent.2[order(DEGs_number_between_tissue.ae.melt.parent.2$DE),]
DEGs_number_between_tissue.ae.melt.parent.2

### making ggplot for DEGs 
p.DEGs_number_between_tissue.ae.parent.2 <- ggplot(data = DEGs_number_between_tissue.ae.melt.parent.2) + theme_grey()
p.DEGs_number_between_tissue.ae.parent.2 <- p.DEGs_number_between_tissue.ae.parent.2 + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number), stat = "identity")
p.DEGs_number_between_tissue.ae.parent.2 <- p.DEGs_number_between_tissue.ae.parent.2 + facet_grid(~tissue)
p.DEGs_number_between_tissue.ae.parent.2 <- p.DEGs_number_between_tissue.ae.parent.2 + geom_text(data=DEGs_number_between_tissue.ae.melt.parent.2,aes(x=DE,y=number*1.05,label=factor(number),color=DE)) 
p.DEGs_number_between_tissue.ae.parent.2 <- p.DEGs_number_between_tissue.ae.parent.2 + labs(y = "number of differentially expressed genes", x = "")

p.DEGs_number_between_tissue.ae.parent.2 
ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/genome_mapped/DEGs_number_between_tissue.ae.parent.png", width = 11, height = 8)

### Da-Ol-1 
lrt.young.vs.bolting.ol.2 <- glmLRT(fit.parent.2, contrast = c(0,0,0,0,0,1,0,0,0,-1))
topTags(lrt.young.vs.bolting.ol.2)
summary(de.young.vs.bolting.ol.2 <- decideTestsDGE(lrt.young.vs.bolting.ol.2, p=0.05))
DEgene.young.vs.bolting.ol.2 <- topTags(lrt.young.vs.bolting.ol.2,n = Inf)$table[topTags(lrt.young.vs.bolting.ol.2,n = Inf)$table$FDR<0.05,]
# -1  2350
# 0  51596
# 1   4540 

lrt.bolting.vs.flowering.ol.2 <- glmLRT(fit.parent.2, contrast = c(0,0,0,0,0,-1,0,1,0,0))
topTags(lrt.bolting.vs.flowering.ol.2)
summary(de.bolting.vs.flowering.ol.2 <- decideTestsDGE(lrt.bolting.vs.flowering.ol.2, p=0.05))
DEgene.bolting.vs.flowering.ol.2 <- topTags(lrt.bolting.vs.flowering.ol.2,n = Inf)$table[topTags(lrt.bolting.vs.flowering.ol.2,n = Inf)$table$FDR<0.05,]
# -1   840
# 0  55623
# 1   2023  

lrt.flowering.vs.early.silique.ol.2 <- glmLRT(fit.parent.2, contrast = c(0,0,0,0,0,0,1,-1,0,0))
topTags(lrt.flowering.vs.early.silique.ol.2)
summary(de.flowering.vs.early.silique.ol.2 <- decideTestsDGE(lrt.flowering.vs.early.silique.ol.2, p=0.05))
DEgene.flowering.vs.early.silique.ol.2 <- topTags(lrt.flowering.vs.early.silique.ol.2,n = Inf)$table[topTags(lrt.flowering.vs.early.silique.ol.2,n = Inf)$table$FDR<0.05,]
# -1  5393
# 0  50586
# 1   2507 

lrt.early.vs.late.silique.ol.2 <- glmLRT(fit.parent.2, contrast = c(0,0,0,0,0,0,-1,0,1,0))
topTags(lrt.early.vs.late.silique.ol.2)
summary(de.early.vs.late.silique.ol.2 <- decideTestsDGE(lrt.early.vs.late.silique.ol.2, p=0.05))
DEgene.early.vs.late.silique.ol.2 <- topTags(lrt.early.vs.late.silique.ol.2,n = Inf)$table[topTags(lrt.early.vs.late.silique.ol.2,n = Inf)$table$FDR<0.05,]
# -1  1388
# 0  55090
# 1   2008

### ggplot 
### reformat for ggplot barplot 
DEGs_number_between_tissue.ol.2 <- data.frame("bolting-vs-young" = as.data.frame(summary(de.young.vs.bolting.ol.2 <- decideTestsDGE(lrt.young.vs.bolting.ol.2, p=0.05)))$Freq, 
                                            "flowering-vs-bolting" = as.data.frame(summary(de.bolting.vs.flowering.ol.2 <- decideTestsDGE(lrt.bolting.vs.flowering.ol.2, p=0.05)))$Freq,
                                     "early.slique-vs-flowering"= as.data.frame(summary(de.flowering.vs.early.silique.ol.2 <- decideTestsDGE(lrt.flowering.vs.early.silique.ol.2, p=0.05)))$Freq, 
                                     "late.silique-vs-early.silique" = as.data.frame(summary(de.early.vs.late.silique.ol.2 <- decideTestsDGE(lrt.early.vs.late.silique.ol.2, p=0.05)))$Freq)

rownames(DEGs_number_between_tissue.ol.2) <- c("down", "no", "up")
DEGs_number_between_tissue.ol.2 <- DEGs_number_between_tissue.ol.2[c("down", "up"),]
DEGs_number_between_tissue.ol.2

DEGs_number_between_tissue.ol.melt.parent.2 <- melt(DEGs_number_between_tissue.ol.2)
DEGs_number_between_tissue.ol.melt.parent.2$DE <- rep(c("down", "up"), 4)
colnames(DEGs_number_between_tissue.ol.melt.parent.2) <- c("tissue", "number", "DE")
DEGs_number_between_tissue.ol.melt.parent.2

# reorder: up 1st down 2nd 
DEGs_number_between_tissue.ol.melt.parent.2$DE <- factor(DEGs_number_between_tissue.ol.melt.parent.2$DE, levels = c("up", "down"))

DEGs_number_between_tissue.ol.melt.parent.2 <- DEGs_number_between_tissue.ol.melt.parent.2[order(DEGs_number_between_tissue.ol.melt.parent.2$DE),]
DEGs_number_between_tissue.ol.melt.parent.2

### making ggplot for DEGs 
p.DEGs_number_between_tissue.ol.parent.2 <- ggplot(data = DEGs_number_between_tissue.ol.melt.parent.2) + theme_grey()
p.DEGs_number_between_tissue.ol.parent.2 <- p.DEGs_number_between_tissue.ol.parent.2 + geom_bar(mapping = aes(fill=DE, x = factor(DE), y = number), stat = "identity")
p.DEGs_number_between_tissue.ol.parent.2 <- p.DEGs_number_between_tissue.ol.parent.2 + facet_grid(~tissue) 
p.DEGs_number_between_tissue.ol.parent.2 <- p.DEGs_number_between_tissue.ol.parent.2 + geom_text(data=DEGs_number_between_tissue.ol.melt.parent.2,aes(x=DE,y=number*1.05,label=factor(number),color=DE)) 
p.DEGs_number_between_tissue.ol.parent.2 <- p.DEGs_number_between_tissue.ol.parent.2 + labs(y = "number of differentially expressed genes", x = "")

p.DEGs_number_between_tissue.ol.parent.2
ggsave("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/figure/genome_mapped/DEGs_number_between_tissue.ol.parent.png", width = 11, height = 8) 
```

# new design to see how many fatty acid genes' expression are differentially expressed between genotypes and during developmental stages  
```{r}
source("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/function_BnRNAseq.R")
# assign group 
ftable(parent.read.count.2.sample,row.vars=c("tissue"),col.vars=c("batch","genotype"))

# normalize 
dge.new.2 <- DGEList(counts=parent.read.count.2.small, group=parent.read.count.2.sample$group)
dim(dge.new.2) # [1] 58486    27
dge.new.2 <- calcNormFactors(dge.new.2, method = "TMM") 
dge.new.2$sample 
# design matrix 
parent.read.count.2.sample$genotype <- as.factor(parent.read.count.2.sample$genotype)
parent.read.count.2.sample$tissue <- as.factor(parent.read.count.2.sample$tissue)
parent.read.count.2.sample$genotype <- relevel(parent.read.count.2.sample$genotype,ref="Da-Ol-1")
parent.read.count.2.sample$tissue <- relevel(parent.read.count.2.sample$tissue,ref="Young")

design.new.2 <- model.matrix(~tissue*genotype,data = parent.read.count.2.sample) 
colnames(design.new.2)

# calculate dispersion
dge.new.2 <- estimateGLMCommonDisp(dge.new.2, design.new.2,verbose = TRUE) # Disp = 0.25646 , BCV = 0.5064 
dge.new.2 <- estimateGLMTrendedDisp(dge.new.2,design.new.2)
dge.new.2 <- estimateGLMTagwiseDisp(dge.new.2,design.new.2)
# plotBCV(dge.new)

# use interaction as coefficient and see how many fatty acid genes are important for interaction term... 
fit.new.2 <- glmFit(dge.new.2, design.new.2)
# gt:tissue 
lrt.new.interaction.2 <- glmLRT(fit.new.2,coef = c("tissuebolting:genotypeDa-Ae", "tissueearly-silique:genotypeDa-Ae", "tissueflowering:genotypeDa-Ae", "tissuelate-silique:genotypeDa-Ae"))
topTags(lrt.new.interaction.2)
DEgene.new.interaction.2 <- topTags(lrt.new.interaction.2,n = Inf)$table[topTags(lrt.new.interaction.2,n = Inf)$table$FDR<0.05,]
nrow(DEgene.new.interaction.2) # number of genes with interaction effect in any tissue types # 1512

# check expression pattern
# reformat the data for expression pattern figure 
tmp <- data.frame(row.names = c(1:1512),
                  V1 =rownames(DEgene.new.interaction.2))
head(tmp)
expression.pattern.Bn.parent(head(tmp)) 

# genes for tissue types 
lrt.new.tissue.2 <- glmLRT(fit.new.2,coef = c("tissuebolting", "tissueearly-silique", "tissueflowering", "tissuelate-silique"))
topTags(lrt.new.tissue.2)
DEgene.new.tissue.2 <- topTags(lrt.new.tissue.2,n = Inf)$table[topTags(lrt.new.tissue.2,n = Inf)$table$FDR<0.05,]
nrow(DEgene.new.tissue.2) # number of genes for tissue effect # 23244  

### 
tmp2 <- data.frame(row.names = c(1:23244),
                  V1 =rownames(DEgene.new.tissue.2))
head(tmp2)
expression.pattern.Bn.parent(head(tmp2)) 
###
 
# genes for gt in any tissue type 
lrt.new.gt.2 <- glmLRT(fit.new.2,coef = c("genotypeDa-Ae", "tissuebolting:genotypeDa-Ae", "tissueearly-silique:genotypeDa-Ae", "tissueflowering:genotypeDa-Ae", "tissuelate-silique:genotypeDa-Ae"))
topTags(lrt.new.gt.2)
DEgene.new.gt.2 <- topTags(lrt.new.gt.2,n = Inf)$table[topTags(lrt.new.gt.2,n = Inf)$table$FDR<0.05,]
nrow(DEgene.new.gt.2) # number of genes for gt effect # 12842 

#####
tmp3 <- data.frame(row.names = c(1:12842),
                  V1 =rownames(DEgene.new.gt.2))
head(tmp3)
expression.pattern.Bn.parent(tail(tmp3)) 
#####

``` 









