---
title: "eQTL_batch_corrected_flipped"
output: html_document
---

* Purpose of this script is to repeat the eQTL analysis with several LG flipped 

### import lib
```{r} 
library(tidyverse) 
library(qtl)
library(snowfall) 
library(ggrepel) 
library(Biostrings) 
source("~/KIAT/function_BnRNAseq.R") 
```

### scanone 

### with flipped LG for eQTL anlaysis just for the plot 

```{r}
cross.F2 <- read.cross("csvsr", genfile ="~/F2/data/QTL_analysis/LG.f2.madmapper.final_gen_revised_flipped.csv", 
                         phefile = "~/F2/output/network_analysis/vstMat.f2.batch.corrected_revised.csv",
                         genotypes = c("AA", "AB", "BB"))   # although the sample IDs are not matched in the original phe and gen file, I still get the right result. 

cross.F2 <- sim.geno(cross.F2,step=1,n.draws=32) # imputation?  
cross.F2 <- calc.genoprob(cross.F2,step=1) 

scanone_eQTL.F2 <- scanone(cross.F2, pheno.col = 2:ncol(cross.F2$pheno), 
	         method = "imp", use = "all.obs")  

### permutation...  
set.seed(12345)   
permtest.F2 <- scanone(cross.F2, pheno.col = 2, method = "imp", n.perm = 10000)  

alphas <- seq(0.01, 0.10, by = 0.01) 
lod.thrs <- summary(permtest.F2, alphas) 
lod.thrs   

save(cross.F2, scanone_eQTL.F2, permtest.F2, lod.thrs, file = "~/F2/output/eQTL/scanone-eqtl_F2_flipped.RData")  
``` 

### determine cis- and trans- eQTL 
```{r}
load("~/F2/output/eQTL/scanone-eqtl_F2_flipped.RData")
scanone_eQTL.F2 %>% dim() # 4887 56182 

scanone_eQTL.F2$chr <- as.character(scanone_eQTL.F2$chr)

# get threshold 
threshold.95 <- lod.thrs[5,] 
threshold.95 # 4.13 

# get all eQTL based on this threshold  
eQTL_sign <- 
sapply(colnames(scanone_eQTL.F2), function(gene) {
  sum(scanone_eQTL.F2[,gene] > threshold.95) > 0 
}) 

sum(eQTL_sign) # 22,868 genes with eQTL 
scanone_eQTL.F2 <- scanone_eQTL.F2[,eQTL_sign]  
dim(scanone_eQTL.F2) # 4887 22868 

# get bayesint result for every gene 
scanone.gather <-  
scanone_eQTL.F2 %>% 
  gather(key = trait, value = LOD, -chr, -pos) 

sig.chrs <- scanone.gather %>% filter(LOD > threshold.95) %>%
  group_by(trait,chr) %>% 
  dplyr::summarise(count = n()) # this is to get the significant chr ID for each trait 

sig.chrs 
sig.chrs %>% dim() # 26581        3

bayesint.list <- apply(sig.chrs,1,function(hit) { # for every row("trait, chr, count") in eigengene module 
    result <- bayesint(scanone_eQTL.F2[c("chr","pos",hit["trait"])],  
                     chr=hit["chr"], 
                     lodcolumn = 1, 
                     expandtomarkers = TRUE 
  )
  colnames(result)[3] <- "LOD" 
  result
}) 

names(bayesint.list) <- sig.chrs$trait

bayesint.list <- lapply(bayesint.list,function(x) 
                          x %>% 
                          as.data.frame() %>%
                          rownames_to_column(var="markername")  %>% # make rownames to column and use "markername" as the colname for the new colomn  
                          mutate(chr=as.character(chr))
) 

bayesint.result <- as.tibble(bind_rows(bayesint.list,.id="trait")) %>% # combine list into tibble 
  dplyr::select(trait,chr,pos,markername,LOD) %>%  
  separate(markername,into=c("chr1","Mbp"),sep="_", convert=TRUE) %>% 
  group_by(trait,chr) %>% 
  dplyr::summarize(start=min(Mbp, na.rm = T),end=max(Mbp, na.rm = T), pos=median(pos, na.rm = T), min_eQTL_LOD=min(LOD),max_eQTL_LOD=max(LOD)) %>% 
  #for the high QTL peaks the interval width is 0.  That is overly precise and need to widen those.
  mutate(start=ifelse(start==end, start-20000,start), end=ifelse(start==end,end+20000,end)) 
  # mutate(start=start-1000000, end=end+1000000) 
# the original code (below) doesn't work correctly for some reason 
# mutate(start=ifelse(start==end,max(0,start-20000),start), end=ifelse(start==end,end+20000,end)) 

bayesint.result %>% dim() # 26582     7 

# get genome range 
library(IRanges)
library(GenomicRanges)
library(GenomicFeatures)
library("rtracklayer")

### get gff file with gene chrom & pos info, gff3 file must be sorted 
gff.mRNA <- read.table("~/Reference/B.napus/gff.mRNA")
dim(gff.mRNA) # 101040      4 
head(gff.mRNA) 
colnames(gff.mRNA) <- c("CHROM", "start", "end", "gene_ID") 
gff.mRNA %>% head()  

# look for cis-eQTL 
bayesint.result.2 <- 
bayesint.result %>% 
  mutate(gene_ID = trait, eQTL_chr = chr, eQTL_start = start, eQTL_end = end) %>% 
  dplyr::select(trait, gene_ID, eQTL_chr, pos, eQTL_start, eQTL_end, min_eQTL_LOD, max_eQTL_LOD) %>% 
  left_join(gff.mRNA, by = "gene_ID") 

bayesint.result.2$eQTL_chr <- paste("chr", bayesint.result.2$eQTL_chr, sep = "")
bayesint.result.2 %>% head()

cis_eQTL <- 
bayesint.result.2 %>% 
  filter(eQTL_chr == CHROM) %>% 
  filter((start < eQTL_start & end > eQTL_start) | # also need SNP pos... 
         (start >= eQTL_start & end <= eQTL_end) |
         (start < eQTL_end & end > eQTL_end)) 

dim(cis_eQTL) # 8894   11 

trans_eQTL <- 
bayesint.result.2 %>% 
  anti_join(cis_eQTL) 

save(cis_eQTL, trans_eQTL, file = "~/F2/output/eQTL/cis_trans_result_new_flipped.Rdata")
```

### make a plot of cis- and trans- eQTL, x-axis being eQTL physical position, y-axis being transcript start position (omit this part for now because I need to plot genetic position VS physical postion but for some the LG I flipped the physical order, need to discuss this with Julin)

* problem of this part: 1) flipped direction of genetic map VS physical map (flip it back and redo the analysis?); 2) many cis-like genes were identified as trans-eQTL because gene location is a little bit away from QTL loci (add 500kb flanking bp to the QTL loci?); 3) the cis-like distribution of trans-eQTL on homeologous chromosomes (the truth or misassembly?). 

```{r} 
library(ggplot2)
load("~/F2/output/eQTL/cis_trans_result_new_flipped.Rdata") 

cis_eQTL %>% dim() # 8894   11 
trans_eQTL %>% dim() # 17687    11 

cis_eQTL$class <- rep("cis", nrow(cis_eQTL)) 
trans_eQTL$class <- rep("trans", nrow(trans_eQTL)) 

# get only main chromosome stuff   
eQTL <- rbind(cis_eQTL, trans_eQTL) 
eQTL <- eQTL[grep("random", eQTL$CHROM, value = FALSE, invert = T),] 
eQTL$class <- factor(eQTL$class, levels = c("cis", "trans"))
eQTL$CHROM <- gsub("(chr)([[:print:]]+)", "\\2", eQTL$CHROM)
eQTL$eQTL_chr <- gsub("(chr)([[:print:]]+)", "\\2", eQTL$eQTL_chr)
eQTL$eQTL_chr <- factor(eQTL$eQTL_chr, levels = c("C09", "C08", "C07", "C06", "C05", "C04", "C03", "C02", "C01", "A10", "A09", "A08", "A07", "A06", "A05", "A04", "A03", "A02", "A01"))

# plot out the stuff 
p.eQTL <- 
eQTL %>% 
  ggplot() + 
  geom_point(aes(x = pos, y = start, color = class), size = 0.5) +
  facet_grid(eQTL_chr ~ CHROM, scales = "free", switch="both") + 
  theme_classic() +
  theme(panel.spacing = unit(0, "lines"), panel.border = element_rect(colour = "black", fill=NA, size=0.1)) +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank()) + 
  labs(x = "eQTL genetic position", y = "gene start")

p.eQTL     

p.cis_eQTL <- 
eQTL %>% 
  dplyr::filter(class == "cis") %>%
  ggplot() + 
  geom_point(aes(x = pos, y = start), size = 0.5) +
  facet_grid(eQTL_chr ~ CHROM, scales = "free", switch="both") + 
  theme_classic() +
  theme(panel.spacing = unit(0, "lines"), panel.border = element_rect(colour = "black", fill=NA, size=0.1)) +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank()) + 
  labs(x = "eQTL genetic position", y = "gene start")

p.cis_eQTL

p.trans_eQTL <- 
eQTL %>% 
  dplyr::filter(class == "trans") %>%
  ggplot() + 
  geom_point(aes(x = pos, y = start), size = 0.5) +
  facet_grid(eQTL_chr ~ CHROM, scales = "free", switch="both") + 
  theme_classic() +
  theme(panel.spacing = unit(0, "lines"), panel.border = element_rect(colour = "black", fill=NA, size=0.1)) +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(), 
        axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank()) + 
  labs(x = "eQTL genetic position", y = "gene start") 

p.trans_eQTL 

ggsave(p.eQTL, filename = "~/F2/output/eQTL/figure/p.eQTL.flipped.png", height = 10, width = 10) 
ggsave(p.cis_eQTL, filename = "~/F2/output/eQTL/figure/p.cis_eQTL.flipped.png", height = 10, width = 10)
ggsave(p.trans_eQTL, filename = "~/F2/output/eQTL/figure/p.trans_eQTL.flipped.png", height = 10, width = 10) 
```

 




