---
title: "F2_map_construction_maxrf0.55"
author: "Ruijuan Li"
date: "10/4/2017"
output: html_document
---

### Goal
The goal of this analysis is to build genetic map using F2 genotyping information, F2 genotyping was done based on the SNPs we identified from its parents 
difference compared to the other script, this time I am using max.rf = 0.55  

1) calculate rf using max rf = 0.55 and LOD score = 3 (one map) # https://github.com/leejimmy93/KIAT/blob/master/F2/F2_map_construction_LOD3_rf0.55_missing_0.10.R
2) assign marker to different LG based on madmapper result 
3) order marker within each LG (one map) https://github.com/leejimmy93/KIAT/blob/master/F2/map_AC_maxrf_0.55.R 

```{r}
F2.data <- read.mapmaker(file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_for_one_map_final.txt") 

# two-point rf data  
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.10/twopts.LOD3_rf0.55.Rdata")

# order data 
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.10/madmapper/LG.f2.ord.AC.0.55.Rdata")

# get order for LGs
set.map.fun(type = "kosambi")
LG.f2 <- list() 

for (i in 1:length(LG.f2.ord.0.55)){
  LG.f2[[i]] <- make.seq(LG.f2.ord.0.55[[i]], "safe") 
}

LG.f2
# remove null LGs 
rf.graph.table(LG.f2[[10]])
draw.map(LG.f2, names= F, grid=F, cex.mrk=0.5, cex.grp=0.75, horizontal=T)  
write.map(LG.f2, "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.0.55.map") 
```

# remove marker with double crossover 
```{r}
########### use R/qtl ########################  
# import data 
map <- read.cross("mm", file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_for_one_map_final.txt", mapfile = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.0.55.map")

# get rf figure & map figure for mapping order based map 
summary(map) # 2016 
```

# only 3 more markers are added, doen't really help... 
