---
title: "F2_map_construction"
output: html_document
---

# Goal
The goal of this analysis is to build genetic map using F2 genotyping information, F2 genotyping was done based on the SNPs we identified from its parents 

# install & load package 
```{r}
# install.packages("onemap")
# install.packages("tkrplot", type="source") 
library(tcltk)
library(tkrplot)
library(onemap)  
library(ggplot2)
library(reshape) 
library(tidyverse) 
library("qtl") 
library(snowfall)
```

# import & format data (MAC version)
```{r} 
F2_geno_data <- read.table("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/data/F2_Final_SNP_Calls", header = T)
head(F2_geno_data) 
dim(F2_geno_data) # 18226   172 
class(F2_geno_data)

rownames(F2_geno_data) <- paste(F2_geno_data$CHROM, F2_geno_data$POS, sep = "_")
F2_geno_data <- F2_geno_data[, -c(1:6)]
dim(F2_geno_data) # 18226   166 
colnames(F2_geno_data)
colnames(F2_geno_data) <- gsub("([[:print:]]+)(\\.)([[:digit:]])", "\\3", colnames(F2_geno_data))
colnames(F2_geno_data)  
```

# check missing rate & remove data with high missing rate above 0.85 
```{r}
F2_geno_data_with_missing_rate <- mutate(F2_geno_data, 
       missing_rate = round(apply(F2_geno_data, 1, function(x) sum(is.na(x)))/166, 2)) 

rownames(F2_geno_data_with_missing_rate) <- rownames(F2_geno_data) 
hist(F2_geno_data_with_missing_rate$missing_rate)
max(F2_geno_data_with_missing_rate$missing_rate) # 0.47, the highest missing rate is 0.47 

# remove markers with missing rate greater than 0.85 
F2_geno_data_with_missing_rate_filtered <-  F2_geno_data_with_missing_rate[F2_geno_data_with_missing_rate$missing_rate < 0.15,]
rownames(F2_geno_data_with_missing_rate_filtered) 

dim(F2_geno_data_with_missing_rate_filtered) # 3608 167 # 5797  167 
hist(F2_geno_data_with_missing_rate_filtered$missing_rate)

F2_geno_data_with_missing_rate_filtered.final <- subset(F2_geno_data_with_missing_rate_filtered, select=-missing_rate)
rownames(F2_geno_data_with_missing_rate_filtered.final)

dim(F2_geno_data_with_missing_rate_filtered.final) # 3608 166 # 5797  166 
rownames(F2_geno_data_with_missing_rate_filtered.final) 
```

# calculate pairwise corrlelation and remove SNPs with corrlelation greater or equal to 0.9
some SNPs should have exactly the same genotype across all individuals, they are redundant in terms of genetic map construction, so they should be removed. find those SNPs by doing correlation test. 
```{r}
F2_geno_data_t <- as.data.frame(t(F2_geno_data_with_missing_rate_filtered.final))
F2_geno_data_t.numeric <- data.matrix(F2_geno_data_t)

# delete markers with all same genotypes across individuals 
test <- apply(F2_geno_data_t.numeric, 2, function(x) length(unique(x[!is.na(x)])))
filter.polymorphsm <- test != 1 

F2_geno_data_t.numeric.2 <- F2_geno_data_t.numeric[,filter.polymorphsm]
dim(F2_geno_data_t.numeric.2) # 166 3606  # 166 5795 

# also output save markers with same genotype across individuals
polymorphism <- test == 1
SNP.not.poly <- colnames(F2_geno_data_t.numeric)[polymorphism]
length(SNP.not.poly) # two markers are not polymorphism across individuals 
save(SNP.not.poly, file="SNP.not.poly.Rdata")

# correlation test
options(warn=-1) # suppress warning message
F2_SNP_correlation <- cor(F2_geno_data_t.numeric.2, use = "pairwise.complete.obs")
save(F2_SNP_correlation, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_SNP_correlation.Rdata")
options(warn=0) # unsuppress warning message
dim(F2_SNP_correlation) # 3606 3606 #  5795 5795 

# calcualte number of SNPs with correlation of 1
nrow(which(F2_SNP_correlation == 1 & lower.tri(F2_SNP_correlation), arr.ind = T, useNames = T)) # 172 # 251 pairs of SNPs have identical genotype data 

# find SNPs with correlation of 1 and remove them 
dup.cordinate <- which(F2_SNP_correlation == 1 & lower.tri(F2_SNP_correlation), arr.ind = T, useNames = F)
length(dup.cordinate) # 344 # 502
dup.cordinate.df <- as.data.frame(dup.cordinate)
sample.ID <- colnames(F2_SNP_correlation)

# extract duplicate pair information based on their coordicate
dup.pair <- data.frame(matrix(nrow = nrow(dup.cordinate), ncol = 2))
for (i in 1:nrow(dup.cordinate)){
 dup.pair[i,1] <- sample.ID[dup.cordinate[i,1]]
 dup.pair[i,2] <- sample.ID[dup.cordinate[i,2]]
}

length(dup.pair[,2]) # 171 # 251 
rem <- unique(dup.pair[,2])
length(rem) # 163 markers are removed due to pairwise correlation of 1 # 239 

# save data
save(dup.pair, file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/dup.pair.Rdata")
save(rem, file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/rem.Rdata") 

### 
F2_geno_data_1 <- F2_geno_data_with_missing_rate_filtered.final[!(rownames(F2_geno_data_with_missing_rate_filtered.final) %in% rem),]  

# remove non polymorphism SNPs 
F2_geno_data_2 <- F2_geno_data_1[!(rownames(F2_geno_data_1) %in% SNP.not.poly),]

dim(F2_geno_data_2) # 3443  166, end up with 2932 SNPs for map construction # 5556  166 

save(F2_geno_data_2, file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_data_2.Rdata") # this is non-redundant SNPs with missing data less than 0.1
``` 

# format data for onemap  
```{r}
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_data_2.Rdata")
dim(F2_geno_data_2) # 3443 166  # 5556 166 
head(F2_geno_data_2) 

########### below should be corrected
# read parent data 
Ae_Ol <- read.csv("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/data/vcf.Ae.Ol.intersect.final.csv", stringsAsFactors = F)

# left join to filter Ae_Ol SNP based on F2 genotypes
Ae_Ol$index <- paste(Ae_Ol$CHROM, Ae_Ol$POS, sep = "_")
F2_geno_data_2$index <- rownames(F2_geno_data_2)
F2_geno_data_2_Ae_Ol <- 
left_join(F2_geno_data_2, Ae_Ol, by="index") %>% 
  dplyr::select(-(X:ALT)) 

F2_geno_data_2_Ae_Ol <- as.matrix(F2_geno_data_2_Ae_Ol) 

# reassign genotype according to parents genotypes
F2_geno_data_2_Ae_Ol_new <- data.frame()

for (i in colnames(F2_geno_data_2_Ae_Ol)[1:166]) {
  print(i)
  for (j in 1:nrow(F2_geno_data_2_Ae_Ol)){
    if (is.na(F2_geno_data_2_Ae_Ol[j,i])){
    F2_geno_data_2_Ae_Ol_new[j,i] = "-"
    } else if (F2_geno_data_2_Ae_Ol[j,i] == "0/1"){
      F2_geno_data_2_Ae_Ol_new[j,i] = "H"
    } else if (F2_geno_data_2_Ae_Ol[j,i] == F2_geno_data_2_Ae_Ol[j,"Ae.gt"]){
      F2_geno_data_2_Ae_Ol_new[j,i] = "A"
    } else {
      F2_geno_data_2_Ae_Ol_new[j,i] = "B"
    }
  }
}

rownames(F2_geno_data_2_Ae_Ol_new) <- F2_geno_data_2_Ae_Ol[,"index"]
F2_geno_data_2_Ae_Ol_new[1:10, 1:10]
############ start working from here ############### 
dim(F2_geno_data_2_Ae_Ol_new) # 3443 166  # 5556 166 
write.table(F2_geno_data_2_Ae_Ol_new, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.15_5556/F2_geno_data_2_Ae_Ol_new_5556.txt")
# change file format in linux 
# cat F2_geno_data_2_Ae_Ol_new_5556.txt | sed 's/"//g' | awk '{first = $1; $1 = ""; print $0}' | sed 's/ //g' > tmp
# tail -3443 tmp > tmp.1 

write.table(rownames(F2_geno_data_2_Ae_Ol_new), file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/marker_info_5556.txt" )
# cat marker_info.txt | awk '{print "*"$2}' | tail -5556 | sed 's/"//g' > marker_info_reform.txt

# paste marker_info_reform.txt tmp.1 | awk '{print $1 " " $2}' > F2_geno_for_one_map.txt 
# cat header_one_map_input F2_geno_for_one_map.txt > F2_geno_for_one_map_final.txt 
# change header info: maker number to the right marker number 
```

# calculate two point rf & assign to linkage groups (on whitney) 
```{r}
# F2_map_construction_LOD3_rf0.5_missing_0.10.R 
```

While ripple doesn't help to condense map distance, I decided to increase min LOD and max rf to see whether this way can help reduce map distance 
```{r}
# on whitney # takes about 45 mins 
# min LOD 4.0 and max rf 5.0 
# min LOD 3.0 and max rf 4.0 
# min LOD 4.0 and max rf 4.0
```

### use madmapper to assign markers to clusters/LGs 
[http://cgpdb.ucdavis.edu/XLinkage/MadMapper/Genetic_Map_MadMapper_Arabidopsis.html]

### running madmapper 
### 1) format marker data to madmapper accepted format (on MAC)
```{r}
# cat F2_geno_data_2_Ae_Ol_new.txt | sed 's/"//g' | tr " " "\t" > F2_geno_data_m.txt 
# vim edit add ";" to the 1st line

# The above 3 lines I think, was not correct... 

# mv F2_geno_data_m.txt F2_geno_data_m.loc 
# wc -l F2_geno_data_m.loc 5557 
# cat F2_geno_data_m.loc | awk '{print NF}' | sort | uniq # 167 
# scp to whitney 
```

* madmapper (on whitney)
http://cgpdb.ucdavis.edu/XLinkage/MadMapper/#Part_2
```{r} 
# https://github.com/leejimmy93/KIAT/blob/master/F2/madmapper_5556.sh # the output file we should check is tree.clust file 
``` 

check madmapper output (export *.tree.clust file to Excel, and decide which SNP belong to which LG based on tree value) /Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.15_5556/F2_geno_data_m_loc_mmout.x_tree_clust.xlsx 

* get linkage group from madmapper generated tree.clust file 
```{r}
madmapper.result <- read.csv("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.15_5556/F2_geno_data_m_loc_mmout.x_tree_clust.csv", header = F)

LG.madmapper <- 
lapply(c(1:19), function(index) {
  madmapper.result[madmapper.result$V34==index,]$V26
})

LG.madmapper[[19]]

length(LG.madmapper) # 19 
save(LG.madmapper, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.15_5556/LG.madmapper.R")
```   

# get index for the above markers in different linkage group 
```{r}
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.15_5556/twopts.LOD3_rf0.5_5556.Rdata")
twopts.f2.LOD3_rf0.5$marnames

group.AC <- list()

for (i in 1:length(LG.madmapper)) {
  group.AC[[i]] <-  which(twopts.f2.LOD3_rf0.5$marnames %in% LG.madmapper[[i]])
}
group.AC 
save(group.AC, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.15_5556/group.AC.Rdata") 
``` 

#### order within each chromosome 
https://github.com/leejimmy93/KIAT/blob/master/F2/map_AC_LOD3_rf0.5_5556.R

### load data and check map 
```{r}
F2.data <- read.mapmaker(file="~/F2/output/missing_rate_0.15_5556/F2_geno_for_one_map_final_5556.txt")  

# two-point rf data
load("/Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/F2/output/missing_rate_0.15_5556/twopts.LOD3_rf0.5_5556.Rdata")

# order data 
load("/Network/Servers/avalanche.plb.ucdavis.edu/Volumes/Mammoth/Users/ruijuanli/F2/output/missing_rate_0.15_5556/LG.f2.ord.AC.LOD3_rf0.5.Rdata")

# get order 
LG.f2 <- lapply(seq_along(LG.f2.ord), function(LG) make.seq(LG.f2.ord[[LG]], "safe"))
draw.map(LG.f2, names= F, grid=F, cex.mrk=0.5, cex.grp=0.75, horizontal=T)

LG.f2[[13]] 
LG.f2[[2]]
write.map(LG.f2, "~/F2/output/missing_rate_0.15_5556/LG.f2.madmapper.map")
```

# R/qtl package to plot LOD & rf heatmap to check marker order & assignment to LG correctness
```{r}
LG.f2.madmapper <- read.cross("mm", file = "~/F2/output/missing_rate_0.15_5556/F2_geno_for_one_map_final_5556.txt", mapfile = "~/F2/output/missing_rate_0.15_5556/LG.f2.madmapper.map")

summary(LG.f2.madmapper)
plotMap(LG.f2.madmapper) 
summaryMap(LG.f2.madmapper)
```

```{r}
LG.f2.madmapper.before.crossover <- LG.f2.madmapper

for (chr in names(LG.f2.madmapper$geno)) { # for each chromosome in cross genotype data
  my.chr <- get(chr,LG.f2.madmapper$geno) # return the genotype data, including data & map
  print(paste(chr,"NA before",sum(is.na(my.chr$data)))) 
  if(ncol(my.chr$data) > 3) { 
    my.chr$data[,2:(ncol(my.chr$data)-1)] <- sapply(2:(ncol(my.chr$data)-1),function(i) {
      apply(my.chr$data[,(i-1):(i+1)],1,function(gt) {
        if (any(is.na(gt))) return(gt[2]) #technically should be looking at the next genotyped marker.
        if ( (length(unique(gt)) == 2) & (gt[1] == gt[3])) return(NA)
        if ( length(unique(gt))  == 3) return(NA)
        return(gt[2])
      })
    })
  }
  LG.f2.madmapper$geno <- within(LG.f2.madmapper$geno,assign(chr,my.chr))
  print(paste(chr,"NA after",sum(is.na(get(chr,LG.f2.madmapper$geno)$data))))
}  
 
f2.map.new <- est.map(LG.f2.madmapper,verbose=T,error.prob=.01)  
LG.f2.madmapper <- replace.map(LG.f2.madmapper, f2.map.new)
LG.f2.madmapper.after.crossover <- LG.f2.madmapper
plot.map(LG.f2.madmapper.after.crossover, alternate.chrid = T) # the old genetic map
plot.map(LG.f2.madmapper.before.crossover,LG.f2.madmapper.after.crossover, alternate.chrid = T) # genetic map comparison
```

give up... 



















