---
title: "F2_map_construction"
author: "Ruijuan Li"
date: "4/18/2017"
output: html_document 
---

# Goal
The goal of this analysis is to build genetic map using F2 genotyping information, F2 genotyping was done based on the SNPs we identified from its parents 

# install & load package 
```{r}
# install.packages("onemap")
# install.packages("tkrplot", type="source") 
library(tcltk)
library(tkrplot)
library(onemap)  
library(ggplot2)
library(reshape) 
library(tidyverse) 
library("qtl") 
library(snowfall)
```

# import & format data (MAC version)
```{r} 
F2_geno_data <- read.table("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/data/F2_Final_SNP_Calls", header = T)
head(F2_geno_data) 
dim(F2_geno_data) # 18226   172 
class(F2_geno_data)

rownames(F2_geno_data) <- paste(F2_geno_data$CHROM, F2_geno_data$POS, sep = "_")
F2_geno_data <- F2_geno_data[, -c(1:6)]
dim(F2_geno_data) # 18226   166 
colnames(F2_geno_data)
colnames(F2_geno_data) <- gsub("([[:print:]]+)(\\.)([[:digit:]])", "\\3", colnames(F2_geno_data))
colnames(F2_geno_data)  
```

# check missing rate & remove data with high missing rate above 0.85 
```{r}
F2_geno_data_with_missing_rate <- mutate(F2_geno_data, 
       missing_rate = round(apply(F2_geno_data, 1, function(x) sum(is.na(x)))/166, 2)) 

rownames(F2_geno_data_with_missing_rate) <- rownames(F2_geno_data) 
hist(F2_geno_data_with_missing_rate$missing_rate)
max(F2_geno_data_with_missing_rate$missing_rate) # 0.47, the highest missing rate is 0.47 

# remove markers with missing rate greater than 0.85 
F2_geno_data_with_missing_rate_filtered <-  F2_geno_data_with_missing_rate[F2_geno_data_with_missing_rate$missing_rate < 0.10,]
rownames(F2_geno_data_with_missing_rate_filtered)

dim(F2_geno_data_with_missing_rate_filtered) # 3608 167
hist(F2_geno_data_with_missing_rate_filtered$missing_rate)

F2_geno_data_with_missing_rate_filtered.final <- subset(F2_geno_data_with_missing_rate_filtered, select=-missing_rate)
rownames(F2_geno_data_with_missing_rate_filtered.final)

dim(F2_geno_data_with_missing_rate_filtered.final) # 3608 166 
rownames(F2_geno_data_with_missing_rate_filtered.final) 
```

# calculate pairwise corrlelation and remove SNPs with corrlelation greater or equal to 0.9
some SNPs should have exactly the same genotype across all individuals, they are redundant in terms of genetic map construction, so they should be removed. find those SNPs by doing correlation test. 
```{r}
F2_geno_data_t <- as.data.frame(t(F2_geno_data_with_missing_rate_filtered.final))
F2_geno_data_t.numeric <- data.matrix(F2_geno_data_t)

# delete markers with all same genotypes across individuals 
test <- apply(F2_geno_data_t.numeric, 2, function(x) length(unique(x[!is.na(x)])))
filter.polymorphsm <- test != 1 

F2_geno_data_t.numeric.2 <- F2_geno_data_t.numeric[,filter.polymorphsm]
dim(F2_geno_data_t.numeric.2) # 166 3606  

# also output save markers with same genotype across individuals
polymorphism <- test == 1
SNP.not.poly <- colnames(F2_geno_data_t.numeric)[polymorphism]
length(SNP.not.poly) # two markers are not polymorphism across individuals 
save(SNP.not.poly, file="SNP.not.poly.Rdata")

# correlation test
options(warn=-1) # suppress warning message
F2_SNP_correlation <- cor(F2_geno_data_t.numeric.2, use = "pairwise.complete.obs")
save(F2_SNP_correlation, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_SNP_correlation.Rdata")
options(warn=0) # unsuppress warning message
dim(F2_SNP_correlation) # 3606 3606 

# calcualte number of SNPs with correlation of 1
nrow(which(F2_SNP_correlation == 1 & lower.tri(F2_SNP_correlation), arr.ind = T, useNames = T)) # 172 pairs of SNPs have identical genotype data 

# find SNPs with correlation of 1 and remove them 
dup.cordinate <- which(F2_SNP_correlation == 1 & lower.tri(F2_SNP_correlation), arr.ind = T, useNames = F)
length(dup.cordinate) # 344
dup.cordinate.df <- as.data.frame(dup.cordinate)
sample.ID <- colnames(F2_SNP_correlation)

# extract duplicate pair information based on their coordicate
dup.pair <- data.frame(matrix(nrow = nrow(dup.cordinate), ncol = 2))
for (i in 1:nrow(dup.cordinate)){
 dup.pair[i,1] <- sample.ID[dup.cordinate[i,1]]
 dup.pair[i,2] <- sample.ID[dup.cordinate[i,2]]
}

length(dup.pair[,2]) # 171 
rem <- unique(dup.pair[,2])
length(rem) # 163 markers are removed due to pairwise correlation of 1 

# save data
save(dup.pair, file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/dup.pair.Rdata")
save(rem, file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/rem.Rdata") 

### 
F2_geno_data_1 <- F2_geno_data_with_missing_rate_filtered.final[!(rownames(F2_geno_data_with_missing_rate_filtered.final) %in% rem),]  

# remove non polymorphism SNPs 
F2_geno_data_2 <- F2_geno_data_1[!(rownames(F2_geno_data_1) %in% SNP.not.poly),]

dim(F2_geno_data_2) # 3443  166, end up with 2932 SNPs for map construction 

save(F2_geno_data_2, file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_data_2.Rdata") # this is non-redundant SNPs with missing data less than 0.1
``` 

# format data for onemap  
```{r}
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_data_2.Rdata")
dim(F2_geno_data_2) # 3443 166 
head(F2_geno_data_2) 

########### below should be corrected
# read parent data 
Ae_Ol <- read.csv("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/data/vcf.Ae.Ol.intersect.final.csv", stringsAsFactors = F)

# left join to filter Ae_Ol SNP based on F2 genotypes
Ae_Ol$index <- paste(Ae_Ol$CHROM, Ae_Ol$POS, sep = "_")
F2_geno_data_2$index <- rownames(F2_geno_data_2)
F2_geno_data_2_Ae_Ol <- 
left_join(F2_geno_data_2, Ae_Ol, by="index") %>% 
  select(-(X:ALT)) 

F2_geno_data_2_Ae_Ol <- as.matrix(F2_geno_data_2_Ae_Ol) 

# reassign genotype according to parents genotypes
F2_geno_data_2_Ae_Ol_new <- data.frame()

for (i in colnames(F2_geno_data_2_Ae_Ol)[1:166]) {
  print(i)
  for (j in 1:nrow(F2_geno_data_2_Ae_Ol)){
    if (is.na(F2_geno_data_2_Ae_Ol[j,i])){
    F2_geno_data_2_Ae_Ol_new[j,i] = "-"
    } else if (F2_geno_data_2_Ae_Ol[j,i] == "0/1"){
      F2_geno_data_2_Ae_Ol_new[j,i] = "H"
    } else if (F2_geno_data_2_Ae_Ol[j,i] == F2_geno_data_2_Ae_Ol[j,"Ae.gt"]){
      F2_geno_data_2_Ae_Ol_new[j,i] = "A"
    } else {
      F2_geno_data_2_Ae_Ol_new[j,i] = "B"
    }
  }
}

rownames(F2_geno_data_2_Ae_Ol_new) <- F2_geno_data_2_Ae_Ol[,"index"]
F2_geno_data_2_Ae_Ol_new[1:10, 1:10]
############ start working from here ############### 
dim(F2_geno_data_2_Ae_Ol_new) # 3443 166  
write.table(F2_geno_data_2_Ae_Ol_new, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_data_2_Ae_Ol_new.txt")
# change file format in linux 
# cat F2_geno_data_2_Ae_Ol_new.txt | sed 's/"//g' | awk '{first = $1; $1 = ""; print $0}' | sed 's/ //g' > tmp
# tail -3443 tmp > tmp.1 

write.table(rownames(F2_geno_data_2_Ae_Ol_new), file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/marker_info.txt" )
# cat marker_info.txt | awk '{print "*"$2}' | tail -5556 | sed 's/"//g' > marker_info_reform.txt

# paste marker_info_reform.txt tmp.1 | awk '{print $1 " " $2}' > F2_geno_for_one_map.txt 
# cat header_one_map_input F2_geno_for_one_map.txt > F2_geno_for_one_map_final.txt 
# change header info: maker number to the right marker number 
```

# calculate two point rf & assign to linkage groups (on whitney) 
```{r}
# F2_map_construction_LOD3_rf0.5_missing_0.10.R 
```

### assign markers to different linkage groups based on their chromsome assignment 
use this method to test if I get correct result, test it on one chromosome with the least number of markers 
```{r}
# get the number of markers for each chromosome 
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_data_2.Rdata")
dim(F2_geno_data_2) # 3443 166

F2_geno_data_2$Chr <- gsub("([[:print:]]+)(_)([[:digit:]]+)", "\\1", rownames(F2_geno_data_2)) 
F2_geno_data_2$Pos <- gsub("([[:print:]]+)(_)([[:digit:]]+)", "\\3", rownames(F2_geno_data_2)) 
F2_geno_data_2$index <- seq(1: nrow(F2_geno_data_2))

head(F2_geno_data_2[F2_geno_data_2$Chr=="chrA04",],1)
tail(F2_geno_data_2[F2_geno_data_2$Chr=="chrC09",],1)

sum(F2_geno_data_2$Chr=="chrA01") # 221 1-221  
sum(F2_geno_data_2$Chr=="chrA02") # 92 222-313
sum(F2_geno_data_2$Chr=="chrA03") # 303 314-616
sum(F2_geno_data_2$Chr=="chrA04") # 165 617-781
sum(F2_geno_data_2$Chr=="chrA05") # 156 782-937
sum(F2_geno_data_2$Chr=="chrA06") # 439 938-1376
sum(F2_geno_data_2$Chr=="chrA07") # 290 1377-1666
sum(F2_geno_data_2$Chr=="chrA08") # 178 1667-1844
sum(F2_geno_data_2$Chr=="chrA09") # 352 1845-2196
sum(F2_geno_data_2$Chr=="chrA10") # 295 2197-2491
sum(F2_geno_data_2$Chr=="chrC01") # 135 2492-2626
sum(F2_geno_data_2$Chr=="chrC02") # 67 2627-2693
sum(F2_geno_data_2$Chr=="chrC03") # 188 2694-2881
sum(F2_geno_data_2$Chr=="chrC04") # 161 2882-3042
sum(F2_geno_data_2$Chr=="chrC05") # 51 3043-3093
sum(F2_geno_data_2$Chr=="chrC06") # 105 3094-3198
sum(F2_geno_data_2$Chr=="chrC07") # 124 3199-3322
sum(F2_geno_data_2$Chr=="chrC08") # 73 3323-3395
sum(F2_geno_data_2$Chr=="chrC09") # 48 3396-3443  

i = 1
index = vector()
for (ch in unique(F2_geno_data_2$Chr)){
  while (F2_geno_data_2$Chr[i]==ch){
      i = i + 1 
  }
  print (i-1)
  index[ch] = i-1
}

index 
# test on C09 
# https://github.com/leejimmy93/KIAT/map_A01.R .... mapC09.R 
```

### use madmapper to assign markers to clusters/LGs 
[http://cgpdb.ucdavis.edu/XLinkage/MadMapper/Genetic_Map_MadMapper_Arabidopsis.html]

### running madmapper 
### 1) format marker data to madmapper accepted format (on MAC)
```{r}
# cat F2_geno_data_2_Ae_Ol_new.txt | sed 's/"//g' | tr " " "\t" > F2_geno_data_m.txt 
# vim edit add ";" to the 1st line 
# seq 166 | tr "\n" "\t" > number_line_header 
# vim edit add ";" to number_line_header
# cat number_line_header F2_geno_data_m.txt > mv F2_geno_data_m.txt 
# mv F2_geno_data_m.txt F2_geno_data_m.loc 
# wc -l F2_geno_data_m.loc 3444 
# cat F2_geno_data_m.loc | awk '{print NF}' | sort | uniq 167 
# scp to whitney 
```

* madmapper (on whitney)
http://cgpdb.ucdavis.edu/XLinkage/MadMapper/#Part_2
```{r} 
# https://github.com/leejimmy93/KIAT/blob/master/madmapper.sh
# the output file we should check is tree.clust file 
```

check madmapper output (export *.tree.clust file to Excel, and decide which SNP belong to which LG based on tree value) /Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.10/F2_geno_data_m_loc_mmout.x_tree_clust.xlsx 

* get linkage group from madmapper generated tree.clust file 
```{r}
madmapper.result <- read.csv("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.10/F2_geno_data_m_loc_mmout.x_tree_clust.csv", header = F)

LG.A01 <- madmapper.result$V26[1:212] # 212 
LG.C01 <- madmapper.result$V26[225:355] # 131
LG.A03 <- madmapper.result$V26[356:621] # 266
LG.C03 <- madmapper.result$V26[659:844] # 186
LG.A09 <- madmapper.result$V26[845:1186] # 342
LG.C08 <- madmapper.result$V26[1187:1248] # 62
LG.A06 <- madmapper.result$V26[1261:1699] # 439
LG.A05 <- madmapper.result$V26[1717:1871] # 155
LG.C05 <- madmapper.result$V26[1872:1909] # 38
LG.A02 <- madmapper.result$V26[1924:2015] # 92
LG.A04 <- madmapper.result$V26[2016:2178] # 163
LG.C04 <- madmapper.result$V26[2179:2335] # 157
LG.A07 <- madmapper.result$V26[2336:2626] # 291
LG.A10 <- madmapper.result$V26[2627:2921] # 295
LG.C06 <- madmapper.result$V26[2922:3026] # 105
LG.C09 <- madmapper.result$V26[3027:3062] # 36
LG.A08 <- madmapper.result$V26[3075:3251] # 177
LG.C02 <- madmapper.result$V26[3252:3319] # 68
LG.C07 <- madmapper.result$V26[3320:3442] # 123

######### unincorported LGs
LG.A01.2 <- madmapper.result$V26[213:224] # 12
LG.A03.2 <- madmapper.result$V26[622:658] # 37
LG.A09.2 <- madmapper.result$V26[1251:1259] # 9
LG.C04.C05 <- madmapper.result$V26[1703:1716] # 14
LG.C05.2 <- madmapper.result$V26[1910:1913] # 4
LG.C08.2 <- madmapper.result$V26[1914:1923] # 10 
LG.C09.2 <- madmapper.result$V26[3063:3073] # 11
#########

LG.A <- list(LG.A01, LG.A01.2, LG.A02, LG.A03, LG.A03.2, LG.A04, LG.A05, LG.A06, LG.A07, LG.A08, LG.A09, LG.A09.2, LG.A10)

LG.C <- list(LG.C01, LG.C02, LG.C03, LG.C04, LG.C04.C05, LG.C05, LG.C05.2, LG.C06, LG.C07, LG.C08, LG.C08.2, LG.C09, LG.C09.2)
length(LG.A) # 13 linkage groups 
length(LG.C) # 13 linkage groups 

LG.madmapper <- list(LG.A01, LG.A02, LG.A03, LG.A04, LG.A05, LG.A06, LG.A07, LG.A08, LG.A09, LG.A10, LG.C01, LG.C02, LG.C03, LG.C04, LG.C05, LG.C06, LG.C07, LG.C08, LG.C09) 
length(LG.madmapper) # 19 
```  

# get index for the above markers in different linkage group 
```{r}
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.10/twopts.LOD3_rf0.5.Rdata")
twopts.f2.LOD3_rf0.5$marnames

group.A <- list()

for (i in 1:length(LG.A)) {
  group.A[[i]] <-  which(twopts.f2.LOD3_rf0.5$marnames %in% LG.A[[i]])
}
group.A

group.C <- list()

for (i in 1:length(LG.C)) {
  group.C[[i]] <-  which(twopts.f2.LOD3_rf0.5$marnames %in% LG.C[[i]])
}

group.C
save(group.A, group.C, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/group.Rdata")

group.AC <- list()

for (i in 1:length(LG.madmapper)) {
  group.AC[[i]] <-  which(twopts.f2.LOD3_rf0.5$marnames %in% LG.madmapper[[i]])
}
group.AC 
save(group.AC, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/group.AC.Rdata")  
```

#### order within each chromosome 
```{r}
# scp group.A & group.C to whitney 
# https://github.com/leejimmy93/KIAT/blob/master/F2/map_A.R
# https://github.com/leejimmy93/KIAT/blob/master/F2/map_C.R 
# https://github.com/leejimmy93/KIAT/blob/master/F2/map_AC.R 
``` 

### combine A & C subgenome 
```{r}
F2.data <- read.mapmaker(file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_for_one_map_final.txt") 
F2.data

# two-point rf data  
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.10/twopts.LOD3_rf0.5.Rdata")

# order data 
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.10/madmapper/LG.f2.ord.A.Rdata")
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/missing_rate_0.10/madmapper/LG.f2.ord.C.Rdata")

# get order for A sub
LG.f2 <- list() 

for (i in c(1, 3, 4, 6, 7, 8, 9, 10, 11, 13)){
  LG.f2[[i]] <- make.seq(LG.f2.ord.A[[i]], "safe") 
}

for (i in c(1, 2, 3, 4, 6, 8, 9, 10, 12)){
  LG.f2[[13+i]] <- make.seq(LG.f2.ord.C[[i]], "safe") 
} 

# remove null LGs 
LG.f2 = LG.f2[-which(sapply(LG.f2, is.null))] 
draw.map(LG.f2, names= F, grid=F, cex.mrk=0.5, cex.grp=0.75, horizontal=T)  
write.map(LG.f2, "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.map") 
```

# R/qtl package to plot LOD & rf headmap to check marker order & assignment to LG correctness  
```{r}
LG.f2.C <- read.cross("mm", file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_for_one_map_final.txt", mapfile = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.C.map")

summary(LG.f2.C)
plotMap(LG.f2.C)
# plot.rf(LG.f2.C) 

totmar(LG.f2.C) # 945

LG.f2.C.safe <- read.cross("mm", file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_for_one_map_final.txt", mapfile = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.C.safe.map")

summary(LG.f2.C.safe)
plotMap(LG.f2.C.safe)
# plot.rf(LG.f2.C.safe) 

totmar(LG.f2.C.safe) # 631 

# geno.image(LG.f2.C, alternate.chrid = T) # grid with color pixels for different genotypes
# geno.image(LG.f2.A, alternate.chrid = T) # grid with color pixels for different genotypes

# save(newmap.drop.marker, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/newmap.drop.marker.Rdata")

####### A sub
LG.f2.A <- read.cross("mm", file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_for_one_map_final.txt", mapfile = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.A.map")

summary(LG.f2.A)
plotMap(LG.f2.A)
# plot.rf(LG.f2.A)

totmar(LG.f2.C) # 945

LG.f2.A.safe <- read.cross("mm", file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_for_one_map_final.txt", mapfile = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.A.safe.map")

summary(LG.f2.A.safe) 
plotMap(LG.f2.A.safe) 
totmar(LG.f2.C.safe) # 
# plot.rf(LG.f2.A.safe) 

## all markers on A & C 
LG.f2.madmapper <- read.cross("mm", file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_for_one_map_final.txt", mapfile = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.map")

summary(LG.f2.madmapper) # 2013 
plotMap(LG.f2.madmapper) 
totmar(LG.f2.madmapper)  
# plot.rf(LG.f2.madmapper) 

# write.csv(pull.geno(LG.f2.madmapper),file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.before.doublecrossover.csv") 
```

### refine map 
### ripple 
ripple uses two methods for comparing orders: maximium likelihood and minimal obligate crossovers. It is simpler to count the number of obligate crossover in the data, for a given order, and then choose the order for which the number of obligate crossover is minimized. It is generally recommended using the crossover count method with a large window, followed by maximum likelihood with a smaller window. 
```{r} 
load("~/F2/output/LG.f2.madmapper.before.crossover.Rdata") 

# LG1 looks good, no better order based on ripple 
# LG2 
rip2 <- ripple(cross = LG.f2.madmapper.before.crossover, window = 5, chr = 2, method = "countxo")
summary(rip2)

LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip2[2,], chr = 2) # second order has the lowest crossover, replace with that order 

rip2.1 <- ripple(LG.f2.madmapper.before.crossover, method = "likelihood", 
                 error.prob = 0.001, verbose = FALSE, chr = 2, window = 3)

summary(rip2.1) # original order has the lowest LOD, keep it 

# iterate over all LGs 
# rename to the right name 
names(LG.f2.madmapper.before.crossover$geno)[1:10] <- paste("A", c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10"), sep = "")
names(LG.f2.madmapper.before.crossover$geno)[11:19] <- paste("C", c("01", "02", "03", "04", "05", "06", "07", "08", "09"), sep = "")

sfInit(parallel = TRUE, cpus = 12)
sfLibrary(qtl)
sfExport("LG.f2.madmapper.before.crossover")

rip <- 
  sfLapply(names(LG.f2.madmapper.before.crossover$geno), function(LG) {
    ripple(cross = LG.f2.madmapper.before.crossover, window = 7, chr = LG, method = "countxo")
  })

sfStop()

save(rip, file = "~/F2/output/missing_rate_0.10/rip.Rdata")
 
### pick a better order based on double crossover 
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[2]][2,], chr = 2)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[4]][2,], chr = 4)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[5]][2,], chr = 5)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[6]][2,], chr = 6)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[7]][2,], chr = 7)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[8]][2,], chr = 8)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[9]][2,], chr = 9)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[10]][2,], chr = 10)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[12]][2,], chr = 12)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[13]][2,], chr = 13)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[14]][2,], chr = 14)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[16]][2,], chr = 16)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[17]][2,], chr = 17)
LG.f2.madmapper.before.crossover <- switch.order(LG.f2.madmapper.before.crossover, rip[[18]][2,], chr = 18)

save(LG.f2.madmapper.before.crossover, file = "~/F2/output/LG.f2.madmapper.before.crossover.ripple1.Rdata")

# ripple using maximum likelihood method, window size 3 
# ripple_maximum_likelihood.R 

# after ripple with maximum likelihood, check the result and select orders which gives me the smallest LOD and chrlen --> remove double crossver --> ...  
```

### remove double crossover 
```{r} 
# remove double crossover 
LG.f2.madmapper <- read.cross("mm", file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/F2_geno_for_one_map_final.txt", mapfile = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.map")

LG.f2.madmapper.before.crossover <- LG.f2.madmapper
save(LG.f2.madmapper.before.crossover, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.before.crossover.Rdata")

for (chr in names(LG.f2.madmapper$geno)) { # for each chromosome in cross genotype data
  my.chr <- get(chr,LG.f2.madmapper$geno) # return the genotype data, including data & map
  print(paste(chr,"NA before",sum(is.na(my.chr$data)))) 
  if(ncol(my.chr$data) > 3) { 
    my.chr$data[,2:(ncol(my.chr$data)-1)] <- sapply(2:(ncol(my.chr$data)-1),function(i) {
      apply(my.chr$data[,(i-1):(i+1)],1,function(gt) {
        if (any(is.na(gt))) return(gt[2]) #technically should be looking at the next genotyped marker.
        if ( (length(unique(gt)) == 2) & (gt[1] == gt[3])) return(NA)
        if ( length(unique(gt))  == 3) return(NA)
        return(gt[2])
      })
    })
  }
  LG.f2.madmapper$geno <- within(LG.f2.madmapper$geno,assign(chr,my.chr))
  print(paste(chr,"NA after",sum(is.na(get(chr,LG.f2.madmapper$geno)$data))))
}  

write.csv(pull.geno(LG.f2.madmapper),file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.rm.doublecrossover.csv") 
```

#### restimate map with double crossover removed 
```{r}
f2.map.new <- est.map(LG.f2.madmapper,verbose=T,error.prob=.01)  
LG.f2.madmapper <- replace.map(LG.f2.madmapper, f2.map.new)
LG.f2.madmapper.after.crossover <- LG.f2.madmapper
plot.map(LG.f2.madmapper.after.crossover, alternate.chrid = T) # the old genetic map
plot.map(LG.f2.madmapper.before.crossover,LG.f2.madmapper.after.crossover, alternate.chrid = T) # genetic map comparison

# plot.rf(LG.f2.madmapper, col.scheme = "redblue", alternate.chrid = T) 
# par(mfrow=c(1,1)) # reset to 1:1
save(LG.f2.madmapper.after.crossover, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.after.crossover.Rdata")
save(f2.map.new, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/f2.map.new.Rdata") 
```

### on A10 
### check rf graph for map 
```{r}
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.before.crossover.Rdata")
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.after.crossover.Rdata")

png("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/figure /rf_chr10.png", width=8, height=8, units="in", res=300)
plot.rf(LG.f2.madmapper.before.crossover, chr = '10', main = "RF and LOD scores for chrom A10")
dev.off()

png("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/figure /rf_madmapper.png", width=8, height=8, units="in", res=300)
plot.rf(LG.f2.madmapper.before.crossover, main = "RF and LOD scores")
dev.off() 

png("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/figure /rf_chr10.png", width=8, height=8, units="in", res=300)
plot.rf(LG.f2.madmapper.after.crossover, chr = '10', main = "RF and LOD scores for chrom A10")
dev.off() 

png("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/figure /map_before_crossover.png", width=8, height=8, units="in", res=300)
plot.map(LG.f2.madmapper.before.crossover, alternate.chrid = T) # the old genetic map
dev.off()   

png("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/figure /map_after_crossover.png", width=8, height=8, units="in", res=300)
plot.map(LG.f2.madmapper.after.crossover, alternate.chrid = T) # genetic map comparison
dev.off()

png("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/figure /map_comparison_1.png", width=8, height=8, units="in", res=300)
plot.map(LG.f2.madmapper.before.crossover,LG.f2.madmapper.after.crossover, alternate.chrid = T) # genetic map comparison
dev.off() 
```

### continue refine map, ripple 
```{r}
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.after.crossover.Rdata")
summaryMap(LG.f2.madmapper.after.crossover)
plotMap(LG.f2.madmapper.after.crossover)

totmar(LG.f2.madmapper.after.crossover) # 2013 

# work on A10 
plot.rf(LG.f2.madmapper.after.crossover, chr = "10") # this 
plotMap(LG.f2.madmapper.after.crossover, chr = '10') 

set.seed(16)
LG.f2.madmapper.after.crossover <- orderMarkers(LG.f2.madmapper.after.crossover, chr = c(10), 
	                        window = 5, use.ripple = TRUE, maxit = 4000, 
	                        error.prob = 0.0001, verbose = T)

plotMap(LG.f2.madmapper.after.crossover, chr = '10') 
plot.rf(LG.f2.madmapper.after.crossover, chr = '10') # plot appears better but length increased... doesn't seem to be a good way. 
```

# follow Julin's suggestion, to remove markers with high double crossover (genotyping error)
```{r}
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.after.crossover.Rdata")
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.before.crossover.Rdata")

summaryMap(LG.f2.madmapper.after.crossover) 
plotMap(LG.f2.madmapper.after.crossover, LG.f2.madmapper.before.crossover)

totmar(LG.f2.madmapper.after.crossover) # 2013 

geno_new <- pull.geno(LG.f2.madmapper.after.crossover)
geno <- pull.geno(LG.f2.madmapper.before.crossover)
identical(geno_new, geno) 
dim(geno); dim(geno_new)
geno[1:10, 1:10]
geno_new[1:10, 1:10]

double.crossover.count <- sapply(colnames(geno), function(marker){
  sum(is.na(geno_new[,marker]))-sum(is.na(geno[,marker]))
} 
)

double.crossover.count <- as.data.frame(double.crossover.count)
double.crossover.count
hist(double.crossover.count$double.crossover.count)
marker.drop <- rownames(subset(double.crossover.count, double.crossover.count>10))
length(marker.drop) # 48 markers 

# drop marker with double crossover more than 10 times，  
LG.f2.madmapper.after.crossover.drop <- drop.markers(LG.f2.madmapper.after.crossover, marker.drop)
totmar(LG.f2.madmapper.after.crossover.drop) # 1965 

LG.f2.madmapper.after.crossover.drop <- 
                          orderMarkers(LG.f2.madmapper.after.crossover.drop, 
	                        window = 4, use.ripple = TRUE, maxit = 4000, 
	                        error.prob = 0.0001) # takes 20 mins 

map.new.new <- est.map(LG.f2.madmapper.after.crossover.drop) # takes 1 min 
plot.map(map.new.new, alternate.chrid = TRUE)
# even longer... bad.... 
save(map.new.new, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/map.new.new.Rdata")
```
conclussion: removing markers with high double crossover makes the map even longer... why? 

# continue refine 
```{r}
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/LG.f2.madmapper.after.crossover.Rdata")

# identify markers with possible segregation distortion
gt <- geno.table(LG.f2.madmapper.after.crossover)
seg_dist <- gt[gt$P.value < 0.05/totmar(LG.f2.madmapper.after.crossover),]
seg_dist
dim(seg_dist) # 67 8 

#                 chr missing AA  AB BB not.BB not.AA      P.value
# chrA01_684630     1      13  9 106 38      0      0 4.702676e-08
# chrA01_4318350    1      25  7  99 35      0      0 3.815686e-08
# chrA01_4081437    1      12 12 114 28      0      0 3.602786e-09
# chrA02_6523578    2       8 13  98 47      0      0 6.885616e-06
# chrA02_6264806    2      20  6  93 47      0      0 4.168391e-08 .... this many might be caused by missing data replace SNPs with double cross over 

# compare individuals's genotypes 
cg <- comparegeno(LG.f2.madmapper.after.crossover)
hist(cg, breaks = 200)
rug(cg)
which(cg > 0.9, arr.ind = T) 
# is it possible for two F2s with the same or very similar genotypes? 

# check marker order 
checkAlleles(LG.f2.madmapper.after.crossover) # no apparent problem 







```






### appendix 
### check double cross over 
```{r}
rm.double.crossover <- function(LG){
for (chr in names(LG$geno)) { # for each chromosome in cross genotype data
  my.chr <- get(chr,LG$geno) # return the genotype data, including data & map
  print(paste(chr,"NA before",sum(is.na(my.chr$data)))) 
  if(ncol(my.chr$data) > 3) { 
    my.chr$data[,2:(ncol(my.chr$data)-1)] <- sapply(2:(ncol(my.chr$data)-1),function(i) {
      apply(my.chr$data[,(i-1):(i+1)],1,function(gt) {
        if (any(is.na(gt))) return(gt[2]) #technically should be looking at the next genotyped marker.
        if ( (length(unique(gt)) == 2) & (gt[1] == gt[3])) return(NA)
        if ( length(unique(gt))  == 3) return(NA)
        return(gt[2])
      })
    })
  }
  LG$geno <- within(LG$geno,assign(chr,my.chr))
  print(paste(chr,"NA after",sum(is.na(get(chr,LG$geno)$data))))
} 
}


# write.csv(pull.geno(cross.drop.marker),file="~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/cross.drop.marker.csv")

# newmap.drop.marker <- est.map(cross.drop.marker,verbose=T,error.prob=.01) # why going through 
# # the last section of code doesn't assign new map 
# plot.map(map.F2, alternate.chrid = T) # the old genetic map
# plot.map(map.F2,newmap.drop.marker, alternate.chrid = T) # genetic map comparison
# 
# geno.image(cross.drop.marker, alternate.chrid = T) # grid with color pixels for different genotypes
# 
# save(newmap.drop.marker, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/newmap.drop.marker.Rdata")
``` 
