---
title: "pheno_trait_analysis"
author: "Ruijuan Li"
date: "1/22/2018"
output: html_document
---

Goal of this script is to summarize the trait values, caculate heritability, as well as correlation/heatmap for different traits 

### growth trait 
```{r}
library(tidyverse) 
library(ggplot2) 
library(reshape2) 

width <- read.csv("~/Desktop/F2_paper/data/width.added(indentical).best.fitted.summary.csv")
lobe_number <- read.csv("~/Desktop/F2_paper/data/lobe.n.added(identical).best.fitted.summary.csv")
height <- read.csv("~/Desktop/F2_paper/data/height.added(identical).best.fitted.summary.csv") 
leaf_number <- read.csv("~/Desktop/F2_paper/data/leaf.n.added(identical).best.fitted.summary.csv") 
height %>% dim() # 516 11 
height[,1:2] %>% head() 

ID <- gsub("(r_ID__I)(\\[)([[:print:]]+)(,Intercept])", "\\3", height$X)[347:515] 

### functions  
coef_extract <- function(model_summary, ID){
  coef_Hmax <- model_summary[grep("r_ID__Hmax", model_summary$X) ,"mean"]
  coef_k <- model_summary[grep("r_ID__k", model_summary$X) ,"mean"] 
  coef_I <- model_summary[grep("r_ID__I", model_summary$X) ,"mean"] 
  intercept_Hmax <- model_summary[model_summary$X == "b_Hmax_Intercept", "mean"]
  intercept_k <- model_summary[model_summary$X == "b_k_Intercept", "mean"]
  intercept_I <- model_summary[model_summary$X == "b_I_Intercept", "mean"]

  growth_model_trait <- data.frame(line_ID = ID,
                            Hmax = coef_Hmax + intercept_Hmax,
                            k = coef_k + intercept_k,
                            I = coef_I + intercept_I
                            )
  return(growth_model_trait)
}

coef_extract_width <- function(model_summary, ID){
  
  model_summary <- width
  
  coef_Hmax <- model_summary[grep("r_ID__Hmax", model_summary$X) ,"mean"]
  coef_k <- model_summary[grep("r_ID__k", model_summary$X) ,"mean"] 
  coef_delta <- model_summary[grep("r_ID__delta", model_summary$X) ,"mean"] 
  intercept_Hmax <- model_summary[model_summary$X == "b_Hmax_Intercept", "mean"]
  intercept_k <- model_summary[model_summary$X == "b_k_Intercept", "mean"]
  intercept_delta <- model_summary[model_summary$X == "b_delta_Intercept", "mean"]

  growth_model_trait <- data.frame(line_ID = ID,
                            Hmax = coef_Hmax + intercept_Hmax,
                            k = coef_k + intercept_k,
                            delta = coef_delta + intercept_delta
                            )
  return(growth_model_trait)
}

### load all data 
height <- coef_extract(height, ID = ID)  
lobe_number <- coef_extract(lobe_number, ID = ID)
leaf_number <- coef_extract(leaf_number, ID = ID)
width <- coef_extract_width(width, ID = ID)  

colnames(height)[2:4] <- paste("height", colnames(height)[2:4], sep = "_")
colnames(lobe_number)[2:4] <- paste("lobe_number", colnames(lobe_number)[2:4], sep = "_")
colnames(leaf_number)[2:4] <- paste("leaf_number", colnames(leaf_number)[2:4], sep = "_")
colnames(width)[2:4] <- paste("width", colnames(width)[2:4], sep = "_")

height_parent <- height[c(1:3),]
lobe_number_parent <- lobe_number[c(1:3),]
leaf_number_parent <- leaf_number[c(1:3),] 
width_parent <- width[c(1:3),]

height <- height[c(4:nrow(height)),]
lobe_number <- lobe_number[c(4:nrow(lobe_number)),]
leaf_number <- leaf_number[c(4:nrow(leaf_number)),] 
width <- width[c(4:nrow(width)),]

growth_model_trait <- 
height %>% 
  left_join(leaf_number) %>% 
  left_join(lobe_number) %>% 
  left_join(width) 

parent_model <- 
height_parent %>% 
  left_join(lobe_number_parent) %>%
  left_join(leaf_number_parent) %>%
  left_join(width_parent) 

parent_model %>% dim() # 3 13
test.melt <- parent_model %>% melt()
test.melt %>% head()

growth_model_trait %>% dim() # 166 13
growth_model_trait$id <- gsub("ID", "Sample_F2", growth_model_trait$line_ID) 
growth_model_trait <- growth_model_trait[,-1] 
save(growth_model_trait, file = "~/Desktop/F2_paper/data/growth_model_trait.Rdata")

# png(filename = "~/Desktop/F2_paper/model.png",width=6, height=7, units="in", res=300)
# par(mfrow=c(4,3), mai = c(0.4, 0.4, 0.4, 0.2))  
setEPS()
postscript(file = "~/Desktop/F2_paper/model.eps", width=5.7, height=6.9)
par(mfrow=c(4,3), mai = c(0.3, 0.3, 0.3, 0.3), ps = 8, cex = 1, cex.main = 1, cex.axis = 0.8, tcl = -0.3)

for (i in colnames(growth_model_trait)[1:12]) { 
  hist(growth_model_trait[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i, xlim = c(min(growth_model_trait[,i], test.melt[(test.melt$variable == i),]$value), max(growth_model_trait[,i], test.melt[(test.melt$variable == i),]$value)), mgp=c(.1,.5,.0)) 
  abline(v = test.melt[(test.melt$variable == i & test.melt$line_ID == "Da_Ae"),]$value, col = "red", lwd = 2)
  abline(v = test.melt[(test.melt$variable == i & test.melt$line_ID == "Da_Ol1"),]$value, col = "blue", lwd = 2) 
  abline(v = test.melt[(test.melt$variable == i & test.melt$line_ID == "F1"),]$value, col = "green", lty = 2, lwd = 2)  
  } 
 
dev.off()                
```  

### summary of trait values 
```{r}
phefile = read.csv("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/data/QTL_analysis/F2.pheno.csv", stringsAsFactors = F, row.names = 1, header = F) 
phefile %>% dim() # 44 166 

phefile <- as.data.frame(t(phefile))   

colnames(phefile)
rname <- phefile$id %>% as.character() 

phefile <- 
phefile %>% 
  dplyr::select(-Weight_of_survey, -Number_of_survey_seeds, -No__of_Transplanting_plants, -No__of_Survival_plant, -id)   

phefile <- sapply(colnames(phefile), function(trait) as.numeric(as.character(phefile[,trait])))
phefile <- as.data.frame(phefile)
rownames(phefile) <- rname

trait_sd <- sapply(colnames(phefile), function(trait) sd(phefile[,trait], na.rm = T))

### mean, range, sd, histogram 
trait_summary <- 
summary(phefile) %>% as.data.frame() %>% 
  separate(Freq, into=c("class", "value"), sep = ":") %>% 
  spread(key = class, value = value) %>% 
  dplyr::select(-Var1) 

colnames(trait_summary)  <- c("trait", "1st_Quantile", "3rd_Quantile", "max", "mean", "median", "min", "missing_data", "to_delete")

trait_summary <- 
trait_summary %>% 
  mutate(missing_rate = round(as.numeric(missing_data)/166, 2)) %>% 
  dplyr::select(mean, median, max, min, missing_rate)  

trait_summary <- 
as.data.frame(sapply(trait_summary, function(x)  
   round(as.numeric(x), 2)))

rownames(trait_summary) <- colnames(phefile)
trait_summary$missing_rate <- ifelse(is.na(trait_summary$missing_rate), 0, trait_summary$missing_rate) 
# write.csv(trait_summary, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/for_paper/trait_summary.csv")

### add branch data into phefile 
extra <- read.csv("~/Desktop/F2_paper/data/AdditionalF2Data_updated_20180222.csv")
extra_F2 <- 
extra %>% 
  filter(Name == "F2")  

extra_F2$Individual_.No. <- gsub("138-", "Sample_F2_", extra_F2$Individual_.No.)

phefile$ID <- rownames(phefile) 

phefile <- 
phefile %>% 
  left_join(extra_F2, c("ID" = "Individual_.No.")) %>% 
  mutate(length_of_branch_5 = X5_length_of_branch,
         length_of_branch_10 = X10_length_of_branch,
         length_of_branch_15 = X15_length_of_branch,
         id = ID) %>% 
  dplyr::select(-(Name:X15_length_of_branch), -ID) 

# save(phefile, file = "~/Desktop/F2_paper/data/phefile.Rdata")

### 
library(lubridate) 
flowering_data <- read.csv("~/Desktop/F2_paper/data/flowering_data.csv")
colnames(flowering_data)[c(5,6)] <- c("bolting_date", "flowering_date")

flowering_data <- 
flowering_data %>%  
  filter(Name == "Da-Ol1" | Name == "Da-Ae" | Name == "F1") %>% 
  group_by(Name) %>%
  dplyr::summarise(days_to_bolt = mean(ymd(bolting_date)-ymd(Day_of_sowing), na.rm = T), 
            days_to_flower = mean(ymd(flowering_date)-ymd(Day_of_sowing), na.rm=T),
            bolting_to_flowering = mean(ymd(flowering_date)-ymd(bolting_date), na.rm=T))  

extra_1 <- 
extra %>% 
  filter(Name == "Da-Ol1" | Name == "Da-Ae" | Name == "F1") %>% 
  group_by(Name) %>% 
  dplyr::summarise(plant_weight_2016_05_13 = mean(Plant_weight_2016_05_13, na.rm = T), 
            root_weight_2016_05_13 = mean(Root_weight_2016_05_13, na.rm = T),
            length_of_branch_5 = mean(X5_length_of_branch, na.rm = T), 
            length_of_branch_10 = mean(X10_length_of_branch, na.rm = T),
            length_of_branch_15 = mean(X15_length_of_branch, na.rm = T)) 

library(reshape2) 

extra_2 <- 
flowering_data %>% melt() %>%
  rbind(melt(extra_1)) %>%
  mutate(value = as.numeric(value))

colnames(extra_2) <- c("variable", "PHENOTYPE", "mean")
extra_2$variable <- gsub("Da-Ol1", "Da-Ol-1", extra_2$variable) 

# add oil data 
# oil trait 
pheno.parent.oil <- read.csv("~/Desktop/F2_paper/data/FA_parents_F1_20180416.csv", header = T, as.is = T, skip = 1)
head(pheno.parent.oil) 
pheno.parent.oil <- pheno.parent.oil[,-1] 

colnames(pheno.parent.oil) <- gsub("([[:print:]]+)(C)([[:print:]]+)", "\\1", colnames(pheno.parent.oil))
colnames(pheno.parent.oil) <- gsub("\\.$", "", colnames(pheno.parent.oil))
colnames(pheno.parent.oil) <- gsub("\\.$", "", colnames(pheno.parent.oil))
colnames(pheno.parent.oil) <- gsub("\\.$", "", colnames(pheno.parent.oil))
colnames(pheno.parent.oil) <- gsub("\\.", "\\_", colnames(pheno.parent.oil))  
colnames(pheno.parent.oil)[1] <- "Name" 
pheno.parent.oil$Name[1:6] <- gsub("(Da-Ae|Da-Ol1)([[:print:]]+)", "\\1", pheno.parent.oil$Name[1:6])
pheno.parent.oil$Name[7:9] <- "F1"  
str(pheno.parent.oil)
pheno.parent.oil$cis_11_Eicosenoic_acid[2] <- NA
pheno.parent.oil$cis_11_Eicosenoic_acid[3] <- NA
pheno.parent.oil$Arachidic_acid[2] <- NA

pheno.parent.oil <- 
pheno.parent.oil %>% 
  filter(Name == "Da-Ol1" | Name == "Da-Ae" | Name == "F1") %>% 
  melt() %>% 
  group_by(Name, variable) %>% 
  dplyr::summarise(value = mean(value, na.rm = T))

colnames(pheno.parent.oil) <- c("variable", "PHENOTYPE", "mean")

pheno.parent.oil$PHENOTYPE <- gsub("crude_oil", "Crude_oil_contents", pheno.parent.oil$PHENOTYPE)
pheno.parent.oil$PHENOTYPE <- gsub("Palmitoliec_aicd", "Palmitoliec_acid", pheno.parent.oil$PHENOTYPE)
pheno.parent.oil$PHENOTYPE <- gsub("length_of_branch_5", "height_to_branch_5", pheno.parent.oil$PHENOTYPE)
pheno.parent.oil$PHENOTYPE <- gsub("length_of_branch_10", "height_to_branch_10", pheno.parent.oil$PHENOTYPE)
pheno.parent.oil$PHENOTYPE <- gsub("length_of_branch_15", "height_to_branch_15", pheno.parent.oil$PHENOTYPE)

pheno.parent.oil$variable <- gsub("Da-Ol1", "Da-Ol-1", pheno.parent.oil$variable)  

load("~/Desktop/F2_paper/data/phefile.Rdata") # F2 data 
colnames(phefile) <- gsub("Palmitoliec_aicd", "Palmitoliec_acid", colnames(phefile)) 

pheno.parent.oil <-
extra_2[,c(2,1,3)] %>% 
  rbind(as.data.frame(pheno.parent.oil))  

phefile_oil <- phefile[,grep("oil|acid|seeds|rate|flower|bolt|root|weight|branch", colnames(phefile))] 
phefile_oil <- 
phefile_oil %>% 
  select(-Survival_rate)  
  
colnames(phefile_oil)[22:24] <- c("height_to_branch_5", "height_to_branch_10", "height_to_branch_15")
colnames(phefile_oil)[5] <- "Palmitoleic_acid"
pheno.parent.oil$PHENOTYPE <- gsub("Palmitoliec_acid", "Palmitoleic_acid", pheno.parent.oil$PHENOTYPE)

# png(filename = "~/Desktop/F2_paper/trait_oil.png",width=174, height=200, units="mm", res=300)
setEPS()
postscript(file = "~/Desktop/F2_paper/trait_oil.eps", width=7.4, height=7.9)
par(mfrow=c(6,4), mai = c(0.3, 0.3, 0.3, 0.3), ps = 8, cex = 1, cex.main = 1, cex.axis = 0.8)   

for (i in colnames(phefile_oil)) {
  if(i == "Erucic_acid" |i == "Oleic_acid"){
    hist(phefile_oil[,i], breaks = 30, col = "grey", border = "black", xlab = "", ylab = "", main = i, xlim = c(min(phefile_oil[,i]), 65), mgp=c(.1,.5,.0))
  } else { 
  hist(phefile_oil[,i], breaks = 30, col = "grey", border = "black", xlab = "", ylab = "", main = i, mgp=c(.1,.5,.0))}
  abline(v = pheno.parent.oil[(pheno.parent.oil$PHENOTYPE == i & pheno.parent.oil$variable == "Da-Ae"),]$mean, col = "red", lwd = 2)
  abline(v = pheno.parent.oil[(pheno.parent.oil$PHENOTYPE == i & pheno.parent.oil$variable == "Da-Ol-1"),]$mean, col = "blue", lwd = 2) 
  abline(v = pheno.parent.oil[(pheno.parent.oil$PHENOTYPE == i & pheno.parent.oil$variable == "F1"),]$mean, col = "green", lty= 2, lwd= 2)
  }  

dev.off()        

### raw data for growth traits 
raw_growth <- read.csv("~/Desktop/F2_paper/data/raw_growth_traits.csv")  
raw_growth %>% colnames()
colnames(raw_growth)[3:6] <- gsub("X", "plant_height_",colnames(raw_growth)[3:6])
colnames(raw_growth)[7:10] <- gsub("X", "plant_width_",colnames(raw_growth)[7:10])
colnames(raw_growth)[11:14] <- gsub("X", "lobe_number_",colnames(raw_growth)[11:14])
colnames(raw_growth)[15:18] <- gsub("X", "leaf_number_",colnames(raw_growth)[15:18])
colnames(raw_growth)[19] <- gsub("X", "plant_height_",colnames(raw_growth)[19])

colnames(raw_growth) <- gsub("([[:print:]]+)(\\.1$)", "\\1", colnames(raw_growth))
raw_growth %>% colnames()

raw_growth <- 
raw_growth %>% 
  filter(Name == "Da-Ol1" | Name == "Da-Ae" | Name == "F1") %>% 
  group_by(Name) %>% 
  dplyr::summarise(plant_height_2015.12.18 = mean(plant_height_2015.12.18, na.rm = T), 
            plant_height_2016.01.20 = mean(plant_height_2016.01.20, na.rm = T),
            plant_height_2016.02.17 = mean(plant_height_2016.02.17, na.rm = T), 
            plant_height_2016.03.16 = mean(plant_height_2016.03.16, na.rm = T),
            plant_width_2015.12.18 = mean(plant_width_2015.12.18, na.rm = T),
            plant_width_2016.01.20 = mean(plant_width_2016.01.20, na.rm = T),
            plant_width_2016.02.17 = mean(plant_width_2016.02.17, na.rm = T),
            plant_width_2016.03.16 = mean(plant_width_2016.03.16, na.rm = T),
            lobe_number_2015.12.28 = mean(lobe_number_2015.12.28, na.rm = T),
            lobe_number_2016.01.21 = mean(lobe_number_2016.01.21, na.rm = T),
            lobe_number_2016.03.21 = mean(lobe_number_2016.03.21, na.rm = T),
            lobe_number_2016.02.18 = mean(lobe_number_2016.02.18, na.rm = T),
            leaf_number_2015.12.28 = mean(leaf_number_2015.12.28, na.rm = T),
            leaf_number_2016.01.21 = mean(leaf_number_2016.01.21, na.rm = T),
            leaf_number_2016.02.18 = mean(leaf_number_2016.02.18, na.rm = T),
            leaf_number_2016.03.21 = mean(leaf_number_2016.03.21, na.rm = T),
            plant_height_2016.05.13 = mean(plant_height_2016.05.13, na.rm = T)
            )  

colnames(raw_growth) <- gsub("\\.", "_", colnames(raw_growth))
raw_growth <- melt(raw_growth)
colnames(raw_growth) <- c("variable", "PHENOTYPE", "mean")
raw_growth$variable <- gsub("Da-Ol1", "Da-Ol-1", raw_growth$variable)

phefile_growth_2 <- phefile[,grep("oil|acid|seeds|rate|flower|bolt|root|weight|aicd|branch", colnames(phefile), invert = T)]

reordered <- c("plant_height_2015_12_18", 
               "lobe_number_2015_12_28", 
               "leaf_number_2015_12_28", 
               "plant_width_2015_12_18", 
               "plant_height_2016_01_20",
               "lobe_number_2016_01_21", 
               "leaf_number_2016_01_21", 
               "plant_width_2016_01_20", 
               "plant_height_2016_02_17", 
               "lobe_number_2016_02_18", 
               "leaf_number_2016_02_18", 
               "plant_width_2016_02_17", 
               "plant_height_2016_03_16",
               "lobe_number_2016_03_21", 
               "leaf_number_2016_03_21",
               "plant_width_2016_03_16",
               "plant_height_2016_05_13")

phefile_growth_2 <- phefile_growth_2[,reordered] 
# load("~/Desktop/F2_paper/data/heritability.Rdata")
# 
# for(i in colnames(phefile_growth_2)) {
#   for(j in heritability$Row.names){
#     colnames(phefile_growth_2)[colnames(phefile_growth_2) == i] <- 
#       ifelse(i == j, heritability[heritability$Row.names == j,]$label, i)
#   }
# } 

colnames(phefile_growth_2) 
# png(filename = "~/Desktop/F2_paper/trait_growth.png",width=10, height=7, units="mm", res=300)
 
setEPS()
postscript(file = "~/Desktop/F2_paper/trait_growth.eps", width=6.7, height=7.9)
par(mfrow=c(5,4), mai = c(0.3, 0.3, 0.3, 0.3), ps = 8, cex = 1, cex.main = 1, cex.axis = 0.8, tcl = -0.3)
# par(mfrow=c(5,4), mai = c(0.4, 0.4, 0.4, 0.2)) 

for (i in colnames(phefile_growth_2)) { 
  hist(phefile_growth_2[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i, mgp=c(.1,.5,.0))
  abline(v = raw_growth[(raw_growth$PHENOTYPE == i & raw_growth$variable == "Da-Ae"),]$mean, col = "red", lwd= 2)
  abline(v = raw_growth[(raw_growth$PHENOTYPE == i & raw_growth$variable == "Da-Ol-1"),]$mean, col = "blue", lwd= 2)
  abline(v = raw_growth[(raw_growth$PHENOTYPE == i & raw_growth$variable == "F1"),]$mean, col = "green", lty = 2, lwd= 2)    
  }

dev.off()               
```

### correlation & heatmap for different traits 
```{r}
# pairs plot for different oil composition trait, flowering trait, and growth related trait 
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/trait/phefile.Rdata")

trait_cor <- round(cor(phefile, use = "pairwise.complete.obs"), digits = 2) 
# show pair of traits whose correlation are above 0.5 
high.cor.pair(trait_cor, threshold = 0.5)

write.csv(trait_cor, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/for_paper/trait_cor.csv")

# heatmap 
library(devtools)
# install_github("raivokolde/pheatmap") 
library(pheatmap)
library(RColorBrewer)
rownames(phefile) <- gsub("(Sample_)([[:print:]]+)", "\\2", rownames(phefile))

# change color scheme later on??? 
png(filename = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/for_paper/trait_heatmap.png",width=22, height=6, units="in", res=300)
par(mar = c(3, 3, 3, 3))
pheatmap(as.matrix(t(scale(phefile))),
     scale="row",
     clustering_method = "average", 
     # cellheight = 10,
     # cellwidth = 5,
     border_color=NA,
     fontsize_row = 6, 
     legend = F
)   
dev.off()    

### correlation for growth model trait 
load("~/F2/data/growth_model_trait.Rdata")  
dim(growth_model_trait) # 166 13 
growth_model_trait_cor <- round(cor(growth_model_trait[,2:13], use = "pairwise.complete.obs"), digits = 2)  
# show pair of traits whose correlation are above 0.5 
high.cor.pair(growth_model_trait_cor, threshold = 0.5) 

png(filename = "~/F2/for_paper/growth_model_trait_heatmap.png",width=8, height=4, units="in", res=300)
par(mar = c(3, 3, 3, 3))
pheatmap(as.matrix(t(scale(growth_model_trait[2:13]))),
     scale="row",
     clustering_method = "average", 
     # cellheight = 10,
     # cellwidth = 5,
     border_color=NA,
     fontsize_row = 6, 
     legend = F
)   
dev.off()      
```

### phenotype traits for Da-Ae & Da-Ol-1
get korean greenhouse data for Da-Ae & Da-Ol-1 from 505 dataset.  
Da-Ae (K200) Da-Ol-1 (K201)  

```{r}
library(reshape2)
pheno.505 <- read.csv("~/505/data/505_Phenotype_Bnapus_38_traits.csv")
tmp.1 <- pheno.505[,c("PHENOTYPE", "ID_505_K200", "ID_505_K201")] 
tmp.1 <- melt(tmp.1[grep("MG|field|Field", tmp.1$PHENOTYPE, invert = T),])
tmp.1$PHENOTYPE
tmp.1$class <- rep("growth_traits", nrow(tmp.1))
tmp.1$class[grep("Oil|acid", tmp.1$PHENOTYPE)] <- rep("oil_traits", 6)

pheno.505 %>% colnames()  
fatty_acid_oil_new <- read.csv("~/505/data/phenotype/505_fatty_acid_20171128.csv")
colnames(fatty_acid_oil_new) 

pheno.505.fatty.acid <- 
fatty_acid_oil_new %>% 
  dplyr::select(Line.number:Erucic.acid..C22.1n9.) 

colnames(pheno.505.fatty.acid)

tmp.2 <- 
pheno.505.fatty.acid %>% 
  dplyr::filter(Line.number == "K200" | Line.number == "K201") 

tmp.2 <- as.data.frame(t(tmp.2))   
colnames(tmp.2) <- as.character(unlist(tmp.2[1,])) 
tmp.2 <- tmp.2[-1,]
rownames(tmp.2) <- gsub("([[:print:]]+)(..C)([[:print:]]+)", "\\1", rownames(tmp.2))

rownames(tmp.2) <- gsub("\\.", "_", rownames(tmp.2))
rownames(tmp.2)[1] <- "Oil_content"
tmp.2$PHENOTYPE <- rownames(tmp.2)
tmp.2 <- melt(tmp.2, id.vars = "PHENOTYPE")  
tmp.2$class <- rep("oil_traits", nrow(tmp.2))

save(tmp.1, tmp.2, file = "~/F2/for_paper/heritability/parental_oil.Rdata")

pheno.parent <- rbind(tmp.1, tmp.2) 
pheno.parent$variable <- gsub("ID_505_K200", "Da-Ae", pheno.parent$variable)
pheno.parent$variable <- gsub("ID_505_K201", "Da-Ol-1", pheno.parent$variable)
pheno.parent$variable <- gsub("K200", "Da-Ae", pheno.parent$variable)
pheno.parent$variable <- gsub("K201", "Da-Ol-1", pheno.parent$variable) 

pheno.parent$value <- round(as.numeric(as.character(pheno.parent$value)), 2)
pheno.parent %>% head()  

pheno.parent$PHENOTYPE <- factor(pheno.parent$PHENOTYPE, levels = unique(pheno.parent[order(tolower(pheno.parent$PHENOTYPE)),]$PHENOTYPE))

p.oil.traits <- 
pheno.parent %>% 
  dplyr::filter(class == "oil_traits") %>% 
  dplyr::filter(PHENOTYPE != "Caprylic_acid" & 
                PHENOTYPE != "Capric_acid" & 
                PHENOTYPE != "Lauric_acid" & 
                PHENOTYPE != "Pentadecanoic_acid" & 
                PHENOTYPE != "Heptadecanoic_aci") %>% 
  ggplot() + 
  geom_bar(aes(x = variable, y = value, fill = variable), stat = "summary", fun.y = "mean", width = 0.5) + 
  facet_wrap(~PHENOTYPE, scales = "free") + 
  theme(axis.title = element_blank(), 
        axis.text.x = element_blank()) + 
  scale_fill_manual(values = alpha(c("blue", "red"), .8))

p.growth.traits <-  
pheno.parent %>% 
  dplyr::filter(class == "growth_traits") %>% 
  dplyr::filter(PHENOTYPE != "germ.rate_KR_GH") %>% 
  ggplot() + 
  geom_col(aes(x = variable, y = value, fill = variable), width = 0.5) + 
  facet_wrap(~PHENOTYPE, scales = "free", nrow = 3) + 
  theme(axis.title = element_blank(), 
        axis.text.x = element_blank()) + 
  scale_fill_manual(values = alpha(c("blue", "red1"), .8))

library(cowplot) 

plot.all.paper <-plot_grid(
  p.oil.traits+theme(legend.position="none",axis.text.x=element_blank())+labs(title="oil traits", ylab = "percentage"),
  p.growth.traits+theme(legend.position="bottom", legend.justification = 'center', axis.text.x=element_blank(),  legend.title = element_blank())+labs(title="growth traits", y="unit???"),
  ncol=1, nrow = 2,labels=c("","")) 

plot.all.paper 

ggsave(plot.all.paper, filename = "~/F2/paper/p.parent.pheno.png", width = 15, height = 12)    

### get the main of oil traits for two parents 
pheno.parent.oil <- 
pheno.parent %>% 
  dplyr::filter(class == "oil_traits") %>% 
  dplyr::filter(PHENOTYPE != "Capric_acid" & 
                PHENOTYPE != "Lauric_acid" & 
                PHENOTYPE != "Pentadecanoic_acid") 

pheno.parent.oil <- 
pheno.parent.oil %>% 
  group_by(PHENOTYPE, variable) %>% 
  summarise(mean = mean(value)) 

save(pheno.parent.oil, file = "~/Desktop/F2_paper/data/pheno.parent.oil.Rdata")
```

### prepare phefile for QTL mapping 
```{r}
load("~/Desktop/F2_paper/data/phefile.all.Rdata") 
phefile.all %>% colnames()  

phefile.branch <- 
phefile.all %>% 
  dplyr::select(leaf_number_k, length_of_branch_5, length_of_branch_10, length_of_branch_15, Erucic_acid) %>%
  t() %>% 
  as.data.frame() 

write.csv(phefile.branch, file = "~/Desktop/F2_paper/data/phefile.branch.csv") 
phefile.all <- 
phefile.all %>% 
  dplyr::select(-Survival_rate) 

write.csv(phefile.all, file = "~/Desktop/F2_paper/submission/Online_Resource_Table2.csv") 
```

### heritability 

heritability is formally defined as the proportion of phenotypic variation (VP) that is due to variation in genetic values (VG); Broad-sense heritability, defined as H2 = VG/VP, captures the proportion of phenotypic variation due to genetic values that may include effects due to dominance and epistasis; On the other hand, narrow-sense heritability, h2 = VA/VP, captures only that proportion of genetic variation that is due to additive genetic values (VA).  

```{r}
library(lme4) 
library(lmerTest)
library(reshape2)

# I see that Julin have repliates so that heritability can be caculated after fitting a linear model with fixed and random effect, with line effect as random effect. Then using VarCorr() function I can get the amount of variance due to line and the total amount of variance, then H2 can be calculated by deviding the 1st by the 2nd value. However, here I don't have replicates, so how to calculate this???  

# flowering time 
library(lubridate)
flowering_data <- read.csv("~/Desktop/F2_paper/data/flowering_data.csv")
colnames(flowering_data)[c(5,6)] <- c("bolting_date", "flowering_date")

flowering_data_parent <- 
flowering_data %>%  
  filter(Name == "Da-Ol1" | Name == "Da-Ae" | Name == "F1") %>% 
  group_by(Name) %>%
  mutate(days_to_bolt = ymd(bolting_date)-ymd(Day_of_sowing), 
          days_to_flower = ymd(flowering_date)-ymd(Day_of_sowing),
          bolting_to_flowering = ymd(flowering_date)-ymd(bolting_date)) %>% 
  dplyr::select(Name, Individual_.No., days_to_bolt, days_to_flower, bolting_to_flowering) %>% 
  melt() %>% 
  mutate(value = as.numeric(value))

# plant weight, root weight, and branch data 
extra <- read.csv("~/Desktop/F2_paper/data/AdditionalF2Data_updated_20180222.csv")

weight_root_branch_parent <- 
extra %>% 
  filter(Name == "Da-Ol1" | Name == "Da-Ae" | Name == "F1") %>% 
  group_by(Name) %>% 
  mutate(plant_weight_2016_05_13 = Plant_weight_2016_05_13, 
            root_weight_2016_05_13 = Root_weight_2016_05_13,
            length_of_branch_5 = X5_length_of_branch, 
            length_of_branch_10 = X10_length_of_branch,
            length_of_branch_15 = X15_length_of_branch) %>% 
  dplyr::select(-starts_with("X"), -Plant_weight_2016_05_13, -Root_weight_2016_05_13) %>% 
  melt() %>% 
  mutate(value = as.numeric(value)) 

# raw growth data 
raw_growth <- read.csv("~/Desktop/F2_paper/data/raw_growth_traits.csv")  
raw_growth %>% colnames()
colnames(raw_growth)[3:6] <- gsub("X", "plant_height_",colnames(raw_growth)[3:6])
colnames(raw_growth)[7:10] <- gsub("X", "plant_width_",colnames(raw_growth)[7:10])
colnames(raw_growth)[11:14] <- gsub("X", "lobe_number_",colnames(raw_growth)[11:14])
colnames(raw_growth)[15:18] <- gsub("X", "leaf_number_",colnames(raw_growth)[15:18])
colnames(raw_growth)[19] <- gsub("X", "plant_height_",colnames(raw_growth)[19])

colnames(raw_growth) <- gsub("([[:print:]]+)(\\.1$)", "\\1", colnames(raw_growth))
raw_growth %>% colnames() 

raw_growth_parent <- 
raw_growth <- 
raw_growth %>% 
  filter(Name == "Da-Ol1" | Name == "Da-Ae" | Name == "F1") %>% 
  group_by(Name) %>% 
  melt() %>% 
  mutate(value = as.numeric(value))

raw_growth_parent$variable <- gsub("\\.", "_", raw_growth_parent$variable)
raw_growth_parent$Name <- gsub("Da-Ol1", "Da-Ol-1", raw_growth_parent$Name)

# combine all parental data 
flowering_data_parent %>% colnames()
weight_root_branch_parent %>% colnames()
raw_growth_parent %>% colnames()  

final_parent_no_oil <- 
flowering_data_parent %>%
  rbind(weight_root_branch_parent) %>% 
  rbind(raw_growth_parent) 

colnames(final_parent_no_oil) <- c("genotype", "rep_ID", "PHENOTYPE", "value")

# oil trait 
######## 
pheno.parent.oil <- read.csv("~/Desktop/F2_paper/data/FA_parents_F1_20180416.csv", header = T, as.is = T, skip = 1)
head(pheno.parent.oil) 
pheno.parent.oil <- pheno.parent.oil[,-1] 

colnames(pheno.parent.oil) <- gsub("([[:print:]]+)(C)([[:print:]]+)", "\\1", colnames(pheno.parent.oil))
colnames(pheno.parent.oil) <- gsub("\\.$", "", colnames(pheno.parent.oil))
colnames(pheno.parent.oil) <- gsub("\\.$", "", colnames(pheno.parent.oil))
colnames(pheno.parent.oil) <- gsub("\\.$", "", colnames(pheno.parent.oil))
colnames(pheno.parent.oil) <- gsub("\\.", "\\_", colnames(pheno.parent.oil))  
colnames(pheno.parent.oil)[1] <- "Name" 
pheno.parent.oil$Name[1:6] <- gsub("(Da-Ae|Da-Ol1)([[:print:]]+)", "\\1", pheno.parent.oil$Name[1:6])
pheno.parent.oil$Name[7:9] <- "F1"  
str(pheno.parent.oil)
pheno.parent.oil$cis_11_Eicosenoic_acid[2] <- NA
pheno.parent.oil$cis_11_Eicosenoic_acid[3] <- NA
pheno.parent.oil$Arachidic_acid[2] <- NA

pheno.parent.oil.2 <- 
pheno.parent.oil %>% 
  filter(Name == "Da-Ol1" | Name == "Da-Ae" | Name == "F1") %>% 
  group_by(Name) %>% 
  melt() %>% 
  mutate(genotype = Name, rep_ID = date_of_analysis, PHENOTYPE = variable) %>% 
  dplyr::select(genotype, rep_ID, PHENOTYPE, value) %>% 
  mutate(value = as.numeric(value)) 

pheno.parent.oil.2$PHENOTYPE <- gsub("crude_oil", "Crude_oil_contents", pheno.parent.oil.2$PHENOTYPE)
pheno.parent.oil.2$PHENOTYPE <- gsub("Palmitoliec_aicd", "Palmitoliec_acid", pheno.parent.oil.2$PHENOTYPE)
pheno.parent.oil.2$genotype <- gsub("Da-Ol1", "Da-Ol-1", pheno.parent.oil.2$genotype)  

colnames(pheno.parent.oil.2) 
colnames(final_parent_no_oil)

final_parent <- 
  final_parent_no_oil %>% 
  rbind(pheno.parent.oil.2) 

# test for one trait 
test <- 
final_parent_no_oil %>% 
  filter(PHENOTYPE == "days_to_bolt")

lme <- lmer(as.numeric(value) ~ (1|genotype),data=test)
summary(lme) 
var <- as.data.frame(VarCorr(lme)) 
var$vcov[2] 
var$vcov[1]/sum(var$vcov) 
# residual is the environmental effect  

## calculate VarE
VarE <- 
sapply(unique(final_parent$PHENOTYPE), function(i) {
  test <- final_parent %>% 
  filter(PHENOTYPE == i) 
  
  lme <- lmer(as.numeric(value) ~ (1|genotype),data=test)
  var <- as.data.frame(VarCorr(lme)) 
  var$vcov[2]
})

VarE <- data.frame(row.names = unique(final_parent$PHENOTYPE),
                   VarE = VarE)
VarE

dim(VarE) # 43 

## collect F2 data and calculate variance 
load("~/Desktop/F2_paper/data/phefile.all.Rdata")
phefile.all %>% colnames() 

VarP <- 
sapply(colnames(phefile.all), function(i) var(phefile.all[[i]], na.rm = T)) %>% 
  as.data.frame()

colnames(VarP) <- "VarP" 

save(pheno.parent.H2, file = "~/F2/for_paper/heritability/pheno.parent.H2.Rdata")
dim(VarP) 

heritability <- 
merge(VarE, VarP, by = 0) %>% 
  mutate(heritability = (VarP - VarE)/VarP) %>% 
  arrange(desc(heritability)) 

heritability$heritability[3:40] <- round(heritability$heritability[3:40], digits = 2)
heritability[heritability$Row.names == "Caprylic_acid",]$heritability <- "NA"
heritability[heritability$Row.names == "Heptadecanoic_acid",]$heritability <- "NA"

heritability$label <- paste(heritability$Row.names, heritability$heritability, sep = ":") 
heritability 
save(heritability, file = "~/Desktop/F2_paper/data/heritability.Rdata")

# second method, use parental data only to calculate heritability, will not use this method  
heritability_2 <- 
sapply(unique(final_parent$PHENOTYPE), function(i) {
  test <- final_parent %>% 
  filter(PHENOTYPE == i) 
  
  lme <- lmer(as.numeric(value) ~ (1|genotype),data=test)
  var <- as.data.frame(VarCorr(lme)) 
  var$vcov[1]/sum(var$vcov) 
  
}) 

heritability_2 <- data.frame(Row.names = unique(final_parent$PHENOTYPE),
                   heritability_2 = heritability_2)
heritability_2

heritability_all <- 
heritability %>% 
  left_join(heritability_2) %>% 
  mutate(cor = cor(heritability, heritability_2)) %>% 
  arrange(desc(heritability_2)) 

library(ggrepel)
set.seed(111) 

heritability_all %>% 
  dplyr::filter(Row.names != "Heptadecanoic_acid") %>% 
  ggplot(aes(x = heritability, y = heritability_2)) + 
  geom_point() + 
  geom_text_repel(aes(label = Row.names)) + 
  geom_abline(a = 0, b = 1) 
  
heritability_all   
```  

### heatmap for correlation plot 
```{r}
library(WGCNA)
library(pheatmap)
# load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/trait/phefile.Rdata")
load("~/Desktop/F2_paper/data/phefile.Rdata")  
phefile %>% colnames()

# phefile.growth_model <- read.csv("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/trait/growth_model_trait.2.csv", header = F) 
load("~/Desktop/F2_paper/data/growth_model_trait.Rdata")
phefile.growth_model <- growth_model_trait

# phefile.growth_model <- as.data.frame(t(phefile.growth_model))
# colnames(phefile.growth_model) <- as.character(unlist(phefile.growth_model[1,]))
# phefile.growth_model <- phefile.growth_model[2:167,]

phefile.all <- 
phefile.growth_model %>% 
  left_join(phefile) 

rownames(phefile.all) <- phefile.all$id

phefile.all <- 
phefile.all %>% 
  dplyr::select(-id)  

colnames(phefile.all) <- gsub("aicd", "acid", colnames(phefile.all))
write.csv(as.data.frame(t(phefile.all)), "~/Desktop/F2_paper/data/F2.pheno.csv") 

load("~/Desktop/F2_paper/data/heritability.Rdata") 

for(i in colnames(phefile.all)) {
  for(j in heritability$Row.names){
    colnames(phefile.all)[colnames(phefile.all) == i] <- 
      ifelse(i == j, heritability[heritability$Row.names == j,]$label, i)
  } 
} 

phefile.non_time_series <- 
phefile.all[,grep("oil|acid|seeds|rate|flower|bolt|root|weight|Hmax|I|k|delta|branch|2016_03_21|2016_05_13", colnames(phefile.all))]  
save(phefile.all, file = "~/Desktop/F2_paper/data/phefile.all.Rdata")

trait_cor <- round(cor(phefile.non_time_series, use = "pairwise.complete.obs"), digits = 2) 
# show pair of traits whose correlation are above 0.5 
high.cor.pair(trait_cor, threshold = 0.5)
cor.P <- corPvalueStudent(trait_cor, 166) 
cor.sig <- trait_cor
cor.sig[cor.P>0.05] <- NA 
cor.sig 

colnames(trait_cor) <- gsub("Crude_oil_contents:-0.1", "Crude_oil_contents:0", colnames(trait_cor))
colnames(trait_cor) <- gsub("length_of_branch_5:0.4", "height_to_branch_5:0.4", colnames(trait_cor))
colnames(trait_cor) <- gsub("length_of_branch_10:0.52", "height_to_branch_10:0.52", colnames(trait_cor))
colnames(trait_cor) <- gsub("length_of_branch_15:0.79", "height_to_branch_15:0.79", colnames(trait_cor))
colnames(trait_cor) <- gsub(":NA", "", colnames(trait_cor))

rownames(trait_cor) <- gsub("Crude_oil_contents:-0.1", "Crude_oil_contents:0", rownames(trait_cor))
rownames(trait_cor) <- gsub("length_of_branch_5:0.4", "height_to_branch_5:0.4", rownames(trait_cor))
rownames(trait_cor) <- gsub("length_of_branch_10:0.52", "height_to_branch_10:0.52", rownames(trait_cor))
rownames(trait_cor) <- gsub("length_of_branch_15:0.79", "height_to_branch_15:0.79", rownames(trait_cor))
rownames(trait_cor) <- gsub("Palmitoliec_acid:0.86", "Palmitoleic_acid:0.86", rownames(trait_cor))
rownames(trait_cor) <- gsub(":NA", "", rownames(trait_cor))

trait_cor <- trait_cor[,colnames(trait_cor) != "Survival_rate"]
trait_cor <- trait_cor[rownames(trait_cor) != "Survival_rate",]

# png(filename = "~/Desktop/F2_paper/trait_cor_heatmap.png",width=170, height=200, units="mm", res=300)
setEPS()
postscript(file = "~/Desktop/F2_paper/trait_cor_heatmap.eps", width=7.4, height=7.9)
par(mar = c(3, 3, 3, 3), ps = 8, cex = 1, cex.main = 1, cex.axis = 1, tcl = -0.3)

pheatmap(as.matrix(trait_cor),  
     clustering_method = "average", 
     # cellheight = 10,
     # cellwidth = 5,
     border_color=NA,
     fontsize_col = 8, 
     legend = T 
)
dev.off()        
```  

### F2 mapping summary
```{r}
mapping_result_C <- read.table("~/Desktop/F2_paper/data/Star_Stats_C.tab", header = T) # ID is the original wrong ID
mapping_result_C$Sample <- gsub("Sample","Sample_F2", mapping_result_C$Sample)

batch_C <- read.csv("~/Desktop/F2_paper/data/F2_batch_C.csv") 
batch_C$Sample <- paste("Sample_F2_", batch_C$Sample.ID, sep = "") # get Sample with the sample wrong ID 
batch_C <- 
batch_C %>% 
  dplyr::select(Sample, No..of.Sowing) 
dim(batch_C) # 47 
batch_C$Sample
mapping_result_C$Sample

mapping_result_C <- 
mapping_result_C %>%
  left_join(batch_C) %>% 
  mutate(Sample = paste("Sample_F2_", No..of.Sowing, sep = "")) %>% 
  dplyr::select(-No..of.Sowing) 

mapping_result_AB <- read.table("~/Desktop/F2_paper/data/Star_Stats_AB.tab", header = T)
mapping_result <- rbind(mapping_result_AB, mapping_result_C)
mapping_result$Sample 
mapping_result$Sample <- gsub("138-", "F2_", mapping_result$Sample)
mapping_result$Sample <- gsub(".paired", "", mapping_result$Sample)
mapping_result <- 
mapping_result %>% 
  mutate(TotalReads = Number_Input_Reads) %>% 
  dplyr::select(Sample, TotalReads, Percent_Unique_Mapped, Percent_Multi_Mapped, Percent_Unmapped_Too_Short, Percent_Unmapped_Other)

write.csv(mapping_result, file = "~/Desktop/F2_paper/supplementary/mapping_result_F2.csv") 
```

### create SRA submission file 
```{r}
biosample <- read_tsv("~/Desktop/F2_paper/data/attributes.tsv")

parent_sample <- read.csv("/Users/ruijuanli/Desktop/Brassica_project/KIAT_RNA_seq/analysis/parent/parent_summary_corrected.csv")

parent_sample <- 
parent_sample %>% 
  mutate(library_ID = SampleID, filename = paste(SampleID, "_1.fq.gz", sep = ""), filename2 = paste(SampleID, "_2.fq.gz", sep = ""), cultivar = Cultivar, dev_stage = Stage) %>% 
  dplyr::select(library_ID, filename, filename2, cultivar, dev_stage)  

parent_sample$dev_stage <- gsub("Young", "young", parent_sample$dev_stage)
parent_sample$sample_name <- paste(parent_sample$cultivar, parent_sample$dev_stage, sep = "_")
parent_sample$sample_name <- gsub("Da-Ae_late-silique", "Da-Ae_late_silique", parent_sample$sample_name)

F2_sample <- data.frame(sample_name = mapping_result$Sample,
                        library_ID = mapping_result$Sample, 
                        filename = gsub("(Sample_F2_)([[:digit:]]+)", "\\2_1.fq.gz", mapping_result$Sample),
                        filename2 = gsub("(Sample_F2_)([[:digit:]]+)", "\\2_2.fq.gz", mapping_result$Sample))  

F2_sample$sample_name <- as.character(F2_sample$sample_name)
F2_sample$library_ID <- as.character(F2_sample$library_ID)
F2_sample$filename <- as.character(F2_sample$filename)
F2_sample$filename2 <- as.character(F2_sample$filename2)

biosample %>% 
  left_join(parent_sample) %>% 
  left_join(F2_sample) %>% 
  View() 

parent_sample <- 
parent_sample %>% 
  left_join(biosample) %>% 
  dplyr::select(bioproject_accession, accession, library_ID, filename, filename2) 

F2_sample <- 
F2_sample %>% 
  left_join(biosample) %>% 
  dplyr::select(bioproject_accession, accession, library_ID, filename, filename2) 

SRA_to_submit <- rbind(parent_sample, F2_sample) 
write.csv(SRA_to_submit, file = "~/Desktop/F2_paper/submission/SRA_to_submit.csv")

SRA_to_submit_final <- read.csv("~/Desktop/F2_paper/submission/Book3.csv")
SRA_to_submit_final$filename <- as.character(SRA_to_submit_final$filename)
SRA_to_submit_final$filename[c(20:25, 27)] <- as.character(sapply(SRA_to_submit_final$filename[c(20:25, 27)], function(i) paste("parent", i, sep = "_")))

SRA_to_submit_final$filename2 <- as.character(SRA_to_submit_final$filename2)
SRA_to_submit_final$filename2[c(20:25, 27)] <- as.character(sapply(SRA_to_submit_final$filename2[c(20:25, 27)], function(i) paste("parent", i, sep = "_")))

SRA_to_submit_final <- 
SRA_to_submit_final %>% 
  dplyr::select(-filename3, -filename4, -assembly)

write_tsv(SRA_to_submit_final, path = "~/Desktop/SRA_Maloof_lab_KIAT_2018.tsv")  
```

### appendix function for this script  
```{r}
### function to extract pair of traits with high correlation 
high.cor.pair <- function(correlation, threshold){ 

  dup.cordinate <- 
    which(abs(correlation) >= threshold & lower.tri(correlation), arr.ind = T, useNames = F) 
  cor.value <- 
    correlation[which(abs(correlation) >= threshold & lower.tri(correlation), 
                      arr.ind = T, useNames = F)] 

  # dup.number <- length(unique(c(dup.cordinate[,1], dup.cordinate[,2])))
  dup.cordinate.df <- as.data.frame(dup.cordinate)
  dup.cordinate.df$V3 <- cor.value
  sample.ID <- colnames(correlation) 

  # extract duplicate pair information based on their coordicate
  output <- data.frame(matrix(nrow = nrow(dup.cordinate.df), ncol = 3))
  for (i in 1:nrow(dup.cordinate.df)){ 
    output[i,1] <- sample.ID[dup.cordinate.df[i,1]]
    output[i,2] <- sample.ID[dup.cordinate.df[i,2]]
    output[i,3] <- dup.cordinate.df[i,3]
  }

  output   
}       
```  

