---
title: "pheno_trait_analysis"
author: "Ruijuan Li"
date: "1/22/2018"
output: html_document
---

Goal of this script is to summarize the trait values, caculate heritability, as well as correlation/heatmap for different traits 

### growth trait 
```{r}
library(tidyverse)
library(ggplot2)
library(reshape2)

height <- read.csv("~/414_growth_model/height/height.best.fitted.summary.csv")
height %>% dim() # 501 11 
height[,1:2] %>% head() 

ID <- gsub("(r_ID__I)(\\[)([[:print:]]+)(,Intercept])", "\\3", height$X)[341:506] 

### functions 
coef_extract <- function(model_summary, ID){
  coef_Hmax <- model_summary[grep("r_ID__Hmax", model_summary$X) ,"mean"]
  coef_k <- model_summary[grep("r_ID__k", model_summary$X) ,"mean"] 
  coef_I <- model_summary[grep("r_ID__I", model_summary$X) ,"mean"] 
  intercept_Hmax <- model_summary[model_summary$X == "b_Hmax_Intercept", "mean"]
  intercept_k <- model_summary[model_summary$X == "b_k_Intercept", "mean"]
  intercept_I <- model_summary[model_summary$X == "b_I_Intercept", "mean"]

  growth_model_trait <- data.frame(line_ID = ID,
                            Hmax = coef_Hmax + intercept_Hmax,
                            k = coef_k + intercept_k,
                            I = coef_I + intercept_I
                            )
  return(growth_model_trait)
}

coef_extract_width <- function(model_summary, ID){
  
  model_summary <- width
  
  coef_Hmax <- model_summary[grep("r_ID__Hmax", model_summary$X) ,"mean"]
  coef_k <- model_summary[grep("r_ID__k", model_summary$X) ,"mean"] 
  coef_delta <- model_summary[grep("r_ID__delta", model_summary$X) ,"mean"] 
  intercept_Hmax <- model_summary[model_summary$X == "b_Hmax_Intercept", "mean"]
  intercept_k <- model_summary[model_summary$X == "b_k_Intercept", "mean"]
  intercept_delta <- model_summary[model_summary$X == "b_delta_Intercept", "mean"]

  growth_model_trait <- data.frame(line_ID = ID,
                            Hmax = coef_Hmax + intercept_Hmax,
                            k = coef_k + intercept_k,
                            delta = coef_delta + intercept_delta
                            )
  return(growth_model_trait)
}

### load all data 
leaf_number <- read.csv("~/414_growth_model/leaf_number/leaf.n.best.fitted.summary.csv")
lobe_number <- read.csv("~/414_growth_model/lobe_number/lobe.n.best.fitted.summary.csv")
width <- read.csv("~/414_growth_model/width/width.best.fitted.summary.csv")

height <- coef_extract(height, ID = ID)
leaf_number <- coef_extract(leaf_number, ID = ID)
lobe_number <- coef_extract(lobe_number, ID = ID)
width <- coef_extract_width(width, ID = ID) 

colnames(height)[2:4] <- paste("height", colnames(height)[2:4], sep = "_")
colnames(leaf_number)[2:4] <- paste("leaf_number", colnames(leaf_number)[2:4], sep = "_")
colnames(lobe_number)[2:4] <- paste("lobe_number", colnames(lobe_number)[2:4], sep = "_")
colnames(width)[2:4] <- paste("width", colnames(width)[2:4], sep = "_")

growth_model_trait <- 
height %>% 
  left_join(leaf_number) %>% 
  left_join(lobe_number) %>% 
  left_join(width) 

save(growth_model_trait, file = "~/F2/data/growth_model_trait.Rdata")

p.growth_model_trait <- 
growth_model_trait %>% melt() %>% 
  ggplot() + 
  geom_histogram(mapping = aes(value), bins = 15, fill = "white", color = "black") +
  facet_wrap(~variable, scales = "free", ncol = 3, nrow = 4) + 
  labs(x = "", y = "number of lines")

ggsave(p.growth_model_trait, filename = "~/F2/for_paper/p.growth_model_trait.png", width = 7, height = 10)  

save(growth_model_trait, file = "~/F2/output/growth_model_trait.Rdata")
```

### summary of trait values 
```{r}
phefile = read.csv("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/data/QTL_analysis/F2.pheno.csv", stringsAsFactors = F, row.names = 1, header = F) 
phefile %>% dim() # 166 44 

phefile <- as.data.frame(t(phefile))  

colnames(phefile)
rname <- phefile$id %>% as.character()

phefile <- 
phefile %>% 
  dplyr::select(-Weight_of_survey, -Number_of_survey_seeds, -No__of_Transplanting_plants, -No__of_Survival_plant, -id)   

phefile <- sapply(colnames(phefile), function(trait) as.numeric(as.character(phefile[,trait])))
phefile <- as.data.frame(phefile)
rownames(phefile) <- rname

trait_sd <- sapply(colnames(phefile), function(trait) sd(phefile[,trait], na.rm = T))

### mean, range, sd, histogram 
trait_summary <- 
summary(phefile) %>% as.data.frame() %>% 
  separate(Freq, into=c("class", "value"), sep = ":") %>% 
  spread(key = class, value = value) %>% 
  dplyr::select(-Var1) 

colnames(trait_summary)  <- c("trait", "1st_Quantile", "3rd_Quantile", "max", "mean", "median", "min", "missing_data", "to_delete")

trait_summary <- 
trait_summary %>% 
  mutate(missing_rate = round(as.numeric(missing_data)/166, 2)) %>% 
  dplyr::select(mean, median, max, min, missing_rate)  

trait_summary <- 
as.data.frame(sapply(trait_summary, function(x) 
   round(as.numeric(x), 2)))

rownames(trait_summary) <- colnames(phefile)
trait_summary$missing_rate <- ifelse(is.na(trait_summary$missing_rate), 0, trait_summary$missing_rate) 
write.csv(trait_summary, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/for_paper/trait_summary.csv")

### add branch data into phefile 
extra <- read.csv("~/Desktop/F2_paper/data/AdditionalF2Data_updated_20180222.csv")
extra_F2 <- 
extra %>% 
  filter(Name == "F2")  

extra_F2$Individual_.No. <- gsub("138-", "Sample_F2_", extra_F2$Individual_.No.)

phefile$ID <- rownames(phefile) 

phefile <- 
phefile %>% 
  left_join(extra_F2, c("ID" = "Individual_.No.")) %>% 
  mutate(length_of_branch_5 = X5_length_of_branch,
         length_of_branch_10 = X10_length_of_branch,
         length_of_branch_15 = X15_length_of_branch) %>% 
  dplyr::select(-(ID:X15_length_of_branch)) 

load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/for_paper/pheno.parent.oil.Rdata")
pheno.parent.oil$PHENOTYPE <- gsub("Oil_content", "Crude_oil_contents", pheno.parent.oil$PHENOTYPE)
pheno.parent.oil$PHENOTYPE <- gsub("Heptadecanoic_aci", "Heptadecanoic_acid", pheno.parent.oil$PHENOTYPE)
pheno.parent.oil$PHENOTYPE <- gsub("Palmitoliec_aicd", "Palmitoliec_acid", pheno.parent.oil$PHENOTYPE) 

### 
library(lubridate)
flowering_data <- read.csv("~/Desktop/F2_paper/data/flowering_data.csv")
colnames(flowering_data)[c(5,6)] <- c("bolting_date", "flowering_date")

flowering_data <- 
flowering_data %>%  
  filter(Name == "Da-Ol1" | Name == "Da-Ae" | Name == "F1") %>% 
  group_by(Name) %>%
  summarise(days_to_bolt = mean(ymd(bolting_date)-ymd(Day_of_sowing), na.rm = T), 
            days_to_flower = mean(ymd(flowering_date)-ymd(Day_of_sowing), na.rm=T),
            bolting_to_flowering = mean(ymd(flowering_date)-ymd(bolting_date), na.rm=T))  

extra_1 <- 
extra %>% 
  filter(Name == "Da-Ol1" | Name == "Da-Ae" | Name == "F1") %>% 
  group_by(Name) %>% 
  summarise(plant_weight_2016_05_13 = mean(Plant_weight_2016_05_13, na.rm = T), 
            root_weight_2016_05_13 = mean(Root_weight_2016_05_13, na.rm = T),
            length_of_branch_5 = mean(X5_length_of_branch, na.rm = T), 
            length_of_branch_10 = mean(X10_length_of_branch, na.rm = T),
            length_of_branch_15 = mean(X15_length_of_branch, na.rm = T)) 

library(reshape2)

extra_2 <- 
flowering_data %>% melt() %>%
  rbind(melt(extra_1)) %>%
  mutate(value = as.numeric(value))

colnames(extra_2) <- c("variable", "PHENOTYPE", "mean")
extra_2$variable <- gsub("Da-Ol1", "Da-Ol-1", extra_2$variable)

pheno.parent.oil <-
extra_2[,c(2,1,3)] %>% 
  rbind(as.data.frame(pheno.parent.oil))  

phefile_oil <- phefile[,grep("oil|acid|seeds|rate|flower|bolt|root|weight|branch", colnames(phefile))]  
png(filename = "~/Desktop/F2_paper/trait_oil.png",width=15, height=7, units="in", res=300)
par(mfrow=c(4,6), mai = c(0.4, 0.4, 0.4, 0.2)) 

for (i in colnames(phefile_oil)) {
  hist(phefile_oil[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i)
  abline(v = pheno.parent.oil[(pheno.parent.oil$PHENOTYPE == i & pheno.parent.oil$variable == "Da-Ae"),]$mean, col = "red")
  abline(v = pheno.parent.oil[(pheno.parent.oil$PHENOTYPE == i & pheno.parent.oil$variable == "Da-Ol-1"),]$mean, col = "blue") 
  abline(v = pheno.parent.oil[(pheno.parent.oil$PHENOTYPE == i & pheno.parent.oil$variable == "F1"),]$mean, col = "purple")
  }

dev.off()     

### raw data for growth traits 
test <- test[,c(18, 19)]  
rownames(test) <- c("Da-Ol-1", "Da_Ae", "F1_136", "F1_137")
test <- read.csv("~/Desktop/test.csv")
test <- test[c(4, 7, 11, 15),] 
colnames(test) <- gsub("\\.", "_", colnames(test))
colnames(test)
colnames(test)[2:5] <- gsub("X","plant_height_", colnames(test)[2:5])
colnames(test)[6:9] <- gsub("X","plant_width_", colnames(test)[6:9])
colnames(test)[6:9] <- gsub("_1$","", colnames(test)[6:9])
colnames(test)[10:13] <- gsub("X","lobe_number_", colnames(test)[10:13])
colnames(test)[14:17] <- gsub("X","leaf_number_", colnames(test)[14:17])
colnames(test)[14:17] <- gsub("_1$","", colnames(test)[14:17])
colnames(test)[18] <- gsub("X","plant_weight_", colnames(test)[18])
colnames(test)[19] <- gsub("X","root_weight_", colnames(test)[19])
colnames(test)[19] <- gsub("_1$","", colnames(test)[19])
colnames(test)[20] <- gsub("X","plant_height_", colnames(test)[20])
colnames(test)[20] <- gsub("_2$","", colnames(test)[20])

test

test.melt <- test %>% melt() 

phefile_growth<- phefile[,grep("oil|acid|seeds|rate|flower|bolt|root|weight", colnames(phefile), invert = T)] %>% melt() 
phefile_growth$variable <- factor(phefile_growth$variable, levels =  unique(phefile_growth[order(tolower(phefile_growth$variable)),]$variable))

p.trait_growth <- 
  phefile_growth %>% 
  ggplot() + 
  geom_histogram(mapping = aes(value), bins = 15, fill = "white", color = "black") +
  facet_wrap(~variable, scales = "free", ncol = 4) + 
  labs(x = "", y = "number of lines")

ggsave(p.trait_growth, filename = "~/Desktop/F2_paper/p.trait_growth.png", width = 10, height = 7) 
save(phefile, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/trait/phefile.Rdata")

phefile_growth_2 <- phefile[,grep("oil|acid|seeds|rate|flower|bolt|root|weight|aicd", colnames(phefile), invert = T)]

png(filename = "~/Desktop/F2_paper/trait_growth.png",width=15, height=7, units="in", res=300)
par(mfrow=c(4,5), mai = c(0.4, 0.4, 0.4, 0.2)) 

for (i in colnames(phefile_growth_2)) { 
  hist(phefile_growth_2[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i)
  abline(v = test.melt[(test.melt$variable == i & test.melt$X == "Da-Ae"),]$value, col = "red")
  abline(v = test.melt[(test.melt$variable == i & test.melt$X == "Da-Ol-1"),]$value, col = "blue")
  abline(v = test.melt[(test.melt$variable == i & test.melt$X == "F1_136"),]$value, col = "green", lty = 3)
  abline(v = test.melt[(test.melt$variable == i & test.melt$X == "F1_137"),]$value, col = "purple", lty = 3)
  }

dev.off()      
```

### heritability 
```{r}
# heritability is formally defined as the proportion of phenotypic variation (VP) that is due to variation in genetic values (VG); Broad-sense heritability, defined as H2 = VG/VP, captures the proportion of phenotypic variation due to genetic values that may include effects due to dominance and epistasis; On the other hand, narrow-sense heritability, h2 = VA/VP, captures only that proportion of genetic variation that is due to additive genetic values (VA).  

library(lme4)
library(lmerTest)

# I see that Julin have repliates so that heritability can be caculated after fitting a linear model with fixed and random effect, with line effect as random effect. Then using VarCorr() function I can get the amount of variance due to line and the total amount of variance, then H2 can be calculated by deviding the 1st by the 2nd value. However, here I don't have replicates, so how to calculate this???  

```

### correlation & heatmap for different traits 
```{r}
# pairs plot for different oil composition trait, flowering trait, and growth related trait 
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/trait/phefile.Rdata")

trait_cor <- round(cor(phefile, use = "pairwise.complete.obs"), digits = 2) 
# show pair of traits whose correlation are above 0.5 
high.cor.pair(trait_cor, threshold = 0.5)

write.csv(trait_cor, file = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/for_paper/trait_cor.csv")

# heatmap 
library(devtools)
# install_github("raivokolde/pheatmap") 
library(pheatmap)
library(RColorBrewer)
rownames(phefile) <- gsub("(Sample_)([[:print:]]+)", "\\2", rownames(phefile))

# change color scheme later on??? 
png(filename = "~/Desktop/Brassica_project/KIAT_RNA_seq/F2/for_paper/trait_heatmap.png",width=22, height=6, units="in", res=300)
par(mar = c(3, 3, 3, 3))
pheatmap(as.matrix(t(scale(phefile))),
     scale="row",
     clustering_method = "average", 
     # cellheight = 10,
     # cellwidth = 5,
     border_color=NA,
     fontsize_row = 6, 
     legend = F
)   
dev.off()    

### correlation for growth model trait 
load("~/F2/data/growth_model_trait.Rdata")  
dim(growth_model_trait) # 166 13 
growth_model_trait_cor <- round(cor(growth_model_trait[,2:13], use = "pairwise.complete.obs"), digits = 2) 
# show pair of traits whose correlation are above 0.5 
high.cor.pair(growth_model_trait_cor, threshold = 0.5) 

png(filename = "~/F2/for_paper/growth_model_trait_heatmap.png",width=8, height=4, units="in", res=300)
par(mar = c(3, 3, 3, 3))
pheatmap(as.matrix(t(scale(growth_model_trait[2:13]))),
     scale="row",
     clustering_method = "average", 
     # cellheight = 10,
     # cellwidth = 5,
     border_color=NA,
     fontsize_row = 6, 
     legend = F
)   
dev.off()      
```

### heatmap for correlation plot 
```{r}
library(WGCNA)
load("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/trait/phefile.Rdata")
phefile.growth_model <- read.csv("~/Desktop/Brassica_project/KIAT_RNA_seq/F2/output/trait/growth_model_trait.2.csv", header = F)

phefile.growth_model <- as.data.frame(t(phefile.growth_model))
colnames(phefile.growth_model) <- as.character(unlist(phefile.growth_model[1,]))
phefile.growth_model <- phefile.growth_model[2:167,]

phefile$id <- rownames(phefile)

phefile.all <- 
phefile.growth_model %>% 
  left_join(phefile) 

rownames(phefile.all) <- phefile.all$id

phefile.all <- 
phefile.all %>% 
  dplyr::select(-id)  

colnames(phefile.all) <- gsub("aicd", "acid", colnames(phefile.all))

phefile.non_time_series <- 
phefile.all[,grep("oil|acid|seeds|rate|flower|bolt|root|weight|Hmax|I|k|delta", colnames(phefile.all))]  

trait_cor <- round(cor(phefile.non_time_series, use = "pairwise.complete.obs"), digits = 2) 
# show pair of traits whose correlation are above 0.5 
high.cor.pair(trait_cor, threshold = 0.5)
cor.P <- corPvalueStudent(trait_cor, 166)
cor.sig <- trait_cor
cor.sig[cor.P>0.05] <- NA 
cor.sig

png(filename = "~/Desktop/F2_paper/trait_cor_heatmap.png",width=8, height=7, units="in", res=300)
par(mar = c(3, 3, 3, 3))

pheatmap(as.matrix(trait_cor),
     clustering_method = "average", 
     # cellheight = 10,
     # cellwidth = 5,
     border_color=NA,
     fontsize_row = 6, 
     legend = T
)
dev.off()  
```

### phenotype traits for Da-Ae & Da-Ol-1
get korean greenhouse data for Da-Ae & Da-Ol-1 from 505 dataset.  
Da-Ae (K200) Da-Ol-1 (K201) 

There is another dataset which contains more detailed growth traits, I need to talk to Julin on whether that should be included or not. 

```{r}
library(reshape2)
pheno.505 <- read.csv("~/505/data/505_Phenotype_Bnapus_38_traits.csv")
tmp.1 <- pheno.505[,c("PHENOTYPE", "ID_505_K200", "ID_505_K201")] 
tmp.1 <- melt(tmp.1[grep("MG|field|Field", tmp.1$PHENOTYPE, invert = T),])
tmp.1$PHENOTYPE
tmp.1$class <- rep("growth_traits", nrow(tmp.1))
tmp.1$class[grep("Oil|acid", tmp.1$PHENOTYPE)] <- rep("oil_traits", 6)

pheno.505 %>% colnames()  
fatty_acid_oil_new <- read.csv("~/505/data/phenotype/505_fatty_acid_20171128.csv")
colnames(fatty_acid_oil_new) 

pheno.505.fatty.acid <- 
fatty_acid_oil_new %>% 
  dplyr::select(Line.number:Erucic.acid..C22.1n9.) 

colnames(pheno.505.fatty.acid)

tmp.2 <- 
pheno.505.fatty.acid %>% 
  dplyr::filter(Line.number == "K200" | Line.number == "K201") 

tmp.2 <- as.data.frame(t(tmp.2))   
colnames(tmp.2) <- as.character(unlist(tmp.2[1,])) 
tmp.2 <- tmp.2[-1,]
rownames(tmp.2) <- gsub("([[:print:]]+)(..C)([[:print:]]+)", "\\1", rownames(tmp.2))

rownames(tmp.2) <- gsub("\\.", "_", rownames(tmp.2))
rownames(tmp.2)[1] <- "Oil_content"
tmp.2$PHENOTYPE <- rownames(tmp.2)
tmp.2 <- melt(tmp.2, id.vars = "PHENOTYPE")  
tmp.2$class <- rep("oil_traits", nrow(tmp.2))

pheno.parent <- rbind(tmp.1, tmp.2) 
pheno.parent$variable <- gsub("ID_505_K200", "Da-Ae", pheno.parent$variable)
pheno.parent$variable <- gsub("ID_505_K201", "Da-Ol-1", pheno.parent$variable)
pheno.parent$variable <- gsub("K200", "Da-Ae", pheno.parent$variable)
pheno.parent$variable <- gsub("K201", "Da-Ol-1", pheno.parent$variable)

pheno.parent$value <- round(as.numeric(as.character(pheno.parent$value)), 2)
pheno.parent %>% head()  

pheno.parent$PHENOTYPE <- factor(pheno.parent$PHENOTYPE, levels = unique(pheno.parent[order(tolower(pheno.parent$PHENOTYPE)),]$PHENOTYPE))

p.oil.traits <- 
pheno.parent %>% 
  dplyr::filter(class == "oil_traits") %>% 
  dplyr::filter(PHENOTYPE != "Caprylic_acid" & 
                PHENOTYPE != "Capric_acid" & 
                PHENOTYPE != "Lauric_acid" & 
                PHENOTYPE != "Pentadecanoic_acid" & 
                PHENOTYPE != "Heptadecanoic_aci") %>% 
  ggplot() + 
  geom_bar(aes(x = variable, y = value, fill = variable), stat = "summary", fun.y = "mean", width = 0.5) + 
  facet_wrap(~PHENOTYPE, scales = "free") + 
  theme(axis.title = element_blank(), 
        axis.text.x = element_blank()) + 
  scale_fill_manual(values = alpha(c("blue", "red"), .8))

p.growth.traits <-  
pheno.parent %>% 
  dplyr::filter(class == "growth_traits") %>% 
  dplyr::filter(PHENOTYPE != "germ.rate_KR_GH") %>% 
  ggplot() + 
  geom_col(aes(x = variable, y = value, fill = variable), width = 0.5) + 
  facet_wrap(~PHENOTYPE, scales = "free", nrow = 3) + 
  theme(axis.title = element_blank(), 
        axis.text.x = element_blank()) + 
  scale_fill_manual(values = alpha(c("blue", "red1"), .8))

library(cowplot) 

plot.all.paper <-plot_grid(
  p.oil.traits+theme(legend.position="none",axis.text.x=element_blank())+labs(title="oil traits", ylab = "percentage"),
  p.growth.traits+theme(legend.position="bottom", legend.justification = 'center', axis.text.x=element_blank(),  legend.title = element_blank())+labs(title="growth traits", y="unit???"),
  ncol=1, nrow = 2,labels=c("","")) 

plot.all.paper 

ggsave(plot.all.paper, filename = "~/F2/paper/p.parent.pheno.png", width = 15, height = 12)    

### get the main of oil traits for two parents 
pheno.parent.oil <- 
pheno.parent %>% 
  dplyr::filter(class == "oil_traits") %>% 
  dplyr::filter(PHENOTYPE != "Capric_acid" & 
                PHENOTYPE != "Lauric_acid" & 
                PHENOTYPE != "Pentadecanoic_acid") 

pheno.parent.oil <- 
pheno.parent.oil %>% 
  group_by(PHENOTYPE, variable) %>% 
  summarise(mean = mean(value)) 


```

### appendix function for this script 
```{r}
### function to extract pair of traits with high correlation 
high.cor.pair <- function(correlation, threshold){ 

  dup.cordinate <- 
    which(abs(correlation) >= threshold & lower.tri(correlation), arr.ind = T, useNames = F) 
  cor.value <- 
    correlation[which(abs(correlation) >= threshold & lower.tri(correlation), 
                      arr.ind = T, useNames = F)] 

  # dup.number <- length(unique(c(dup.cordinate[,1], dup.cordinate[,2])))
  dup.cordinate.df <- as.data.frame(dup.cordinate)
  dup.cordinate.df$V3 <- cor.value
  sample.ID <- colnames(correlation) 

  # extract duplicate pair information based on their coordicate
  output <- data.frame(matrix(nrow = nrow(dup.cordinate.df), ncol = 3))
  for (i in 1:nrow(dup.cordinate.df)){ 
    output[i,1] <- sample.ID[dup.cordinate.df[i,1]]
    output[i,2] <- sample.ID[dup.cordinate.df[i,2]]
    output[i,3] <- dup.cordinate.df[i,3]
  }

  output   
}      
```

### need to get the coefficient for parents & F1s 
```{r}
width <- read.csv("~/Desktop/F2_paper/width.added(indentical).best.fitted.summary.csv")
lobe_number <- read.csv("~/Desktop/F2_paper/lobe.n.added(identical).best.fitted.summary.csv")
height <- read.csv("~/Desktop/F2_paper/height.added(identical).best.fitted.summary.csv") 

height %>% dim() # 501 11 
height[,1:2] %>% head()  
ID <- gsub("(r_ID__I)(\\[)([[:print:]]+)(,Intercept])", "\\3", height$X)[347:515] 

height <- coef_extract(height, ID = ID) 
lobe_number <- coef_extract(lobe_number, ID = ID)
width <- coef_extract_width(width, ID = ID) 

height <- height[c(1:3),]
lobe_number <- lobe_number[c(1:3),]
width <- width[c(1:3),]

colnames(height)[2:4] <- paste("height", colnames(height)[2:4], sep = "_")
colnames(lobe_number)[2:4] <- paste("lobe_number", colnames(lobe_number)[2:4], sep = "_")
colnames(width)[2:4] <- paste("width", colnames(width)[2:4], sep = "_")

height
lobe_number
width

load("~/Desktop/growth_model_trait.Rdata")
colnames(growth_model_trait)

parent_model <- 
height %>% 
  left_join(lobe_number) %>% 
  left_join(width)

parent_model %>% dim()
test.melt <- parent_model %>% melt()
test.melt %>% head()

growth_model_trait %>% dim()

png(filename = "~/Desktop/F2_paper/model.png",width=15, height=7, units="in", res=300)
par(mfrow=c(3,4), mai = c(0.4, 0.4, 0.4, 0.2)) 

for (i in colnames(growth_model_trait)[2:13]) { 
  hist(growth_model_trait[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i)
  abline(v = test.melt[(test.melt$variable == i & test.melt$line_ID == "Da_Ae"),]$value, col = "red")
  abline(v = test.melt[(test.melt$variable == i & test.melt$line_ID == "Da_Ol1"),]$value, col = "blue") 
  abline(v = test.melt[(test.melt$variable == i & test.melt$line_ID == "F1"),]$value, col = "green", lty = 3)  
  }

dev.off()      
```




