---
title: "505_network_WGCNA"
output: html_document
---

purpose of this script is to do WGCNA network anaysis using 505 SNP data. 

I did network analysis for our 414 F2 population using SNP data (https://github.com/leejimmy93/KIAT/blob/master/F2/F2_network.Rmd & https://github.com/leejimmy93/KIAT/blob/master/F2/F2_network_batch_corrected.Rmd), however, the identified gene co-expression modules correlated with different traits seems to arise from the genetic linkage instead of expression, so we want to see whether we have the same problem for 505. 

### load libs
```{r}
library(WGCNA) 
library(tidyverse) 
library(edgeR) 
library(DESeq2)
library(magrittr) 
library(gplots) 
library(reshape2)
library(sva) 
options(stringsAsFactors = FALSE) 
```

### load expression data, normalize, and correct for batch effect 
```{r}
# load data 
read_count_505 <- read.table("~/505/data/late_silique/read.count.505.tsv", header=T, row.names=1)
dim(read_count_505) # 101040    131 
read_count_505[1:10, 1:10]

# filter low expression tags 
keep <- rowSums(cpm(read_count_505) > 1) >= (131 * 0.25) # 1cpm in at least 1/4 of the samples 
read_count_505 <- read_count_505[keep,]
dim(read_count_505) # 49038   131

colnames(read_count_505) <- gsub("Sample_", "", colnames(read_count_505))
colnames(read_count_505) <- gsub("\\.", "\\-", colnames(read_count_505))

# add batch information 
sample_des_c <- read.csv("~/505/data/phenotype/batch_c.csv", stringsAsFactors=F, header=T, na.strings = "")
sample_des_d <- read.csv("~/505/data/phenotype/batch_d.csv", stringsAsFactors=F, header=T, na.strings = "")
sample_des_e <- read.csv("~/505/data/phenotype/batch_e.csv", stringsAsFactors=F, header=T, na.strings = "")

batch_info <- data.frame(batch = c(rep("C", nrow(sample_des_c)), 
                                  rep("D", nrow(sample_des_d)), 
                                  rep("E", nrow(sample_des_e))), 
                         sample_ID = c(sample_des_c$Sample.ID, sample_des_d$Sample.ID, sample_des_e$Sample.ID))

batch_info <- batch_info[match(colnames(read_count_505), batch_info$sample_ID),] 
all(batch_info$sample_ID == colnames(read_count_505)) # True 

# sample description 
sample.505 <- data.frame(batch = factor(batch_info$batch), 
                        genotype = factor(batch_info$sample_ID))

# vst transformation 
dds.505 <- DESeqDataSetFromMatrix(countData = round(read_count_505), 
                                  colData = sample.505, 
                                  design = ~genotype)

vsd.505 <- varianceStabilizingTransformation(dds.505)
vstMat.505 <- assay(vsd.505)
colnames(vstMat.505) <- colnames(read_count_505) 

# MDS to check batch effect 
vstMat.505 %>% dim() #  49038   131
test1 <- as.matrix(dist(t(vstMat.505)))  

test2 <- as.data.frame(cmdscale(test1))
head(test2) #now we have 2 dimensions 
plot(test2)
test2$sample_ID <- rownames(test2)
test2 %>% 
  left_join(batch_info, by = "sample_ID") %>% 
  ggplot() + 
  geom_point(aes(x = V1, y= V2, color = as.factor(batch))) # looks like there is batch effect on PC2

# correct batch effect 
vstMat.505.batch.corrected <- removeBatchEffect(vstMat.505, batch = batch_info$batch) 
test1 <- as.matrix(dist(t(vstMat.505.batch.corrected)))  
test2 <- as.data.frame(cmdscale(test1))
test2$sample_ID <- rownames(test2)
test2 %>% 
  left_join(batch_info, by = "sample_ID") %>% 
  ggplot() + 
  geom_point(aes(x = V1, y= V2, color = as.factor(batch))) # great, batch effect removed

# get genes with large variation across samples using variance, The variance is a numerical measure of how the data values is dispersed around the mean.
# https://labs.genetics.ucla.edu/horvath/CoexpressionNetwork/Rpackages/WGCNA/faq.html

vstMat.505.batch.corrected <- as.data.frame(vstMat.505.batch.corrected)
vstMat.505.batch.corrected$variance <- as.numeric(apply(vstMat.505.batch.corrected,1,var))  
summary(vstMat.505.batch.corrected$variance) 

vstMat.505.batch.corrected.largeVar<-
  vstMat.505.batch.corrected %>% 
  filter(variance>1) %>% # dispersed one reads arround the mean 
  dplyr::select(-variance) 

dim(vstMat.505.batch.corrected.largeVar) # 14264   132   

save(vstMat.505.batch.corrected.largeVar, file = "~/505/network_analysis/output/vstMat.505.batch.corrected.largeVar.Rdata")
```

### 1) follow Julin's method for network analysis 

### WGCNA for network  

https://labs.genetics.ucla.edu/horvath/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-02-networkConstr-man.pdf  

### unsigned network 
```{r}
enableWGCNAThreads()

load("~/505/network_analysis/output/vstMat.505.batch.corrected.largeVar.Rdata")  
vstMat.505.WGCNA <- vstMat.505.batch.corrected.largeVar
dim(vstMat.505.WGCNA)

# transform expression data 
head(vstMat.505.WGCNA[,1:6])
vstMat.505.WGCNA.t <- t(vstMat.505.WGCNA)
head(vstMat.505.WGCNA.t[,1:6]) 

# check sample quality, this is to see whether there are many missing entry for certain gene, also check variance  
gag.CR <- goodSamplesGenes(vstMat.505.WGCNA.t, verbose = 3)
gag.CR$allOK

# cluster samples to look for outliers based on gene expression data
# hclust 1stly calculate pairwise dissimilarity between each pair of observations, the observations that are most similar to each other are fused, then the next most similar observations... For clusters with multiple observations, use linkage to quantitate their dissimlarity, including complete, single, average, and centroid linkage. 
sampleTREE.CR <- hclust(dist(vstMat.505.WGCNA.t), method = "average") # average is the mean intercluster dissmilarity 
plot(sampleTREE.CR,cex=.6) 

# soft threshold:  In an unsigned coexpression network, modules correspond to clusters of genes with high absolute correlations. In a signed network, modules correspond to positively correlated genes.
powers <- c(c(1:10), seq(from = 12, to=30, by=2))
sft <- pickSoftThreshold(vstMat.505.WGCNA.t, powerVector = powers, verbose = 5, networkType = "unsigned")

sizeGrWindow(9, 5)
par(mfrow = c(1,2)) 
cex1 <- 0.9 
# Scale-free topology fit index as a fCRction of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red")
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a fCRction of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red") 
# For this unsigned networks, while I didn't get plautea but 6 is the suggested power to use and it gives me high R squared to fit a scale free network model. 

# choose 14 
softPower <- 14
adjacency <- adjacency(vstMat.505.WGCNA.t, power = softPower)
# Turn adjacency into topological overlap, topological overlap leads to a network distance measure
dim(adjacency) # 14264 14264
system.time(TOM <- TOMsimilarity(adjacency)); # Calculation of the topological overlap matrix from a given adjacency matrix. takes 30 mins 
dissTOM <- 1-TOM

# Call the hierarchical clustering fCRction
geneTree <- hclust(as.dist(dissTOM), method = "average")
# Plot the resulting clustering tree (dendrogram)
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, hang = 0.04)

# define modules 
# We like large modules, so we set the minimum module size relatively high:
minModuleSize <- 30;
# Module identification using dynamic tree cut:
dynamicMods <- cutreeDynamic(dendro = geneTree, distM = dissTOM,
                             deepSplit <- 2, pamRespectsDendro = FALSE,
                             minClusterSize = minModuleSize);
table(dynamicMods)
table(dynamicMods) %>% length() # 46 modules 

# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and colors CRderneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene dendrogram and module colors")

# merge similar modules
# Calculate eigengenes 
MEList <- moduleEigengenes(vstMat.505.WGCNA.t, colors = dynamicColors) # Calculates module eigengenes (1st principal component) of modules in a given single dataset. 
MEs <- MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss <- 1-cor(MEs);
# Cluster module eigengenes
METree <- hclust(as.dist(MEDiss), method = "average");
# Plot the result
sizeGrWindow(7, 6)
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")

# merge with correlation > 0.8

MEDissThres = 0.2
# Plot the cut line into the dendrogram
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")
abline(h=MEDissThres, col = "red")
# Call an automatic merging fCRction
merge = mergeCloseModules(vstMat.505.WGCNA.t, dynamicColors, cutHeight = MEDissThres, verbose = 3)
# The merged module colors
mergedColors = merge$colors
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs

# compare pre and post merge

sizeGrWindow(12, 9)
#pdf(file = "Plots/geneDendro-3.pdf", wi = 9, he = 6)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
c("Dynamic Tree Cut", "Merged dynamic"),
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)
#dev.off()

# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs

save(MEs, file = "~/505/network_analysis/output/MEs_505.Rdata")
```

### signed network 
```{r}
load("~/505/network_analysis/output/vstMat.505.batch.corrected.largeVar.Rdata")  
vstMat.505.WGCNA <- vstMat.505.batch.corrected.largeVar
dim(vstMat.505.WGCNA)

# transform expression data 
head(vstMat.505.WGCNA[,1:6])
vstMat.505.WGCNA.t <- t(vstMat.505.WGCNA)
head(vstMat.505.WGCNA.t[,1:6]) 

# soft threshold:  In an unsigned coexpression network, modules correspond to clusters of genes with high absolute correlations. In a signed network, modules correspond to positively correlated genes.
powers <- c(c(1:10), seq(from = 12, to=30, by=2))
sft <- pickSoftThreshold(vstMat.505.WGCNA.t, powerVector = powers, verbose = 5, networkType = "signed")

sizeGrWindow(9, 5)
par(mfrow = c(1,2)) 
cex1 <- 0.9 
# Scale-free topology fit index as a fCRction of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red")
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a fCRction of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red") 
# For this signed networks, while I didn't get plautea but 12 is the suggested power to use and it gives me high R squared to fit a scale free network model. 

# choose 12
softPower <- 12
adjacency <- adjacency(vstMat.505.WGCNA.t, power = softPower, type = "signed")
# Turn adjacency into topological overlap, topological overlap leads to a network distance measure
dim(adjacency) # 14264 14264
system.time(TOM <- TOMsimilarity(adjacency)); # Calculation of the topological overlap matrix from a given adjacency matrix. takes 30 mins 
dissTOM <- 1-TOM

# Call the hierarchical clustering fCRction
geneTree <- hclust(as.dist(dissTOM), method = "average")
# Plot the resulting clustering tree (dendrogram)
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, hang = 0.04)

# define modules 
# We like large modules, so we set the minimum module size relatively high:
minModuleSize <- 30;
# Module identification using dynamic tree cut:
dynamicMods <- cutreeDynamic(dendro = geneTree, distM = dissTOM,
                             deepSplit <- 2, pamRespectsDendro = FALSE,
                             minClusterSize = minModuleSize);
table(dynamicMods)
table(dynamicMods) %>% length() # 46 modules 

# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and colors CRderneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene dendrogram and module colors")

# merge similar modules
# Calculate eigengenes 
MEList <- moduleEigengenes(vstMat.505.WGCNA.t, colors = dynamicColors) # Calculates module eigengenes (1st principal component) of modules in a given single dataset. 
MEs <- MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss <- 1-cor(MEs);
# Cluster module eigengenes
METree <- hclust(as.dist(MEDiss), method = "average");
# Plot the result
sizeGrWindow(7, 6)
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")

# merge with correlation > 0.8

MEDissThres = 0.2
# Plot the cut line into the dendrogram
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")
abline(h=MEDissThres, col = "red")
# Call an automatic merging fCRction
merge = mergeCloseModules(vstMat.505.WGCNA.t, dynamicColors, cutHeight = MEDissThres, verbose = 3)
# The merged module colors
mergedColors = merge$colors
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs

# compare pre and post merge

sizeGrWindow(12, 9)
#pdf(file = "Plots/geneDendro-3.pdf", wi = 9, he = 6)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
c("Dynamic Tree Cut", "Merged dynamic"),
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)
#dev.off()

# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs

MEs.signed <- MEs
save(MEs.signed, file = "~/505/network_analysis/output/MEs_505_signed.Rdata")
``` 

### GO enrichment of all modules
```{r}
load("~/505/network_analysis/output/MEs_505.Rdata")
load("~/505/network_analysis/output/MEs_505_signed.Rdata")



```

###  phenotype data for 505 
```{r}
pheno.505 <- read_csv("~/505/data/505_Phenotype_Bnapus_38_traits.csv") # 1st year data, 2016?
pheno.505 <- pheno.505[,!apply(pheno.505,2,function(x) all(is.na(x)))]
pheno.505 %>% dim() 
pheno.505.t <- as.data.frame(t(pheno.505[,-1]))
colnames(pheno.505.t) <- paste(pheno.505$PHENOTYPE, "year1", sep = "_")
pheno.505.t$ID <- rownames(pheno.505.t)

load("~/505/genomic_prediction/data/pheno_data_fatty_acid_all.Rdata") # 2nd year data, on FA related traits 
pheno_data_505 %>% dim()
colnames(pheno_data_505) <- paste(colnames(pheno_data_505), "year2", sep = "_")
colnames(pheno_data_505)[1] <- "ID"
pheno.505.t$ID
pheno_data_505$ID
pheno_data_505$ID <- gsub("\\_|\\-", "", pheno_data_505$ID)
pheno_data_505$ID <- gsub("(505)(K)([[:digit:]]+)", "ID_\\1_\\2\\3", pheno_data_505$ID)

# merge two dataset 
pheno_data_505 <- 
pheno_data_505 %>% 
  left_join(pheno.505.t) 

dim(pheno_data_505)   # 131 57 

# plot the data to see missing rate, distribution, etc... 
summary(pheno_data_505)
colnames(pheno_data_505)

par(mfrow=c(5, 4))

for (i in colnames(pheno_data_505)[2:21]) { 
  hist(pheno_data_505[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i) 
  } 
 
for (i in colnames(pheno_data_505)[22:41]) { 
  hist(pheno_data_505[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i) 
  } 
 
for (i in colnames(pheno_data_505)[42:57]) { 
  hist(pheno_data_505[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i, mgp=c(.1,.5,.0)) 
  } 

save(pheno_data_505, file = "~/505/network_analysis/output/pheno_data_505.Rdata")

# check correlation between phenotypes 
```

### plot several clusters gene expression value & their eigengenes 
```{r}

```

### related modules to trait and identify important genes
```{r}
load("~/505/network_analysis/output/MEs_505_signed.Rdata")
load("~/505/network_analysis/output/pheno_data_505.Rdata")

# include triat value  
dim(MEs.signed) # 36 moduels 
rownames(vstMat.505.WGCNA.t) 
rownames(MEs.signed) 
rownames(MEs.signed) <- rownames(vstMat.505.WGCNA.t)
MEs.signed$ID <- rownames(MEs.signed)

# merge trait and module values 
MEs.signed$ID <- gsub("\\_|\\-", "", MEs.signed$ID)
MEs.signed$ID <- gsub("(505)(K)([[:digit:]]+)", "ID_\\1_\\2\\3", MEs.signed$ID)

# There are two K200 sequencing info (this line was sequenced twice), modify the name, change the 1st K200 to K200-1, the 2nd to K200-2, for both module and trait info  
MEs.signed$ID[15] <- "ID_505_K200-1"
MEs.signed$ID[16] <- "ID_505_K200-2"

pheno_data_505$ID[16] <- "ID_505_K200-1"
pheno_data_505$ID[17] <- "ID_505_K200-2"

ME.blups.505 <- 
MEs.signed %>% 
  left_join(pheno_data_505, by = "ID") 

dim(ME.blups.505) # 36 module + 56 traits + 1 "line"
colnames(ME.blups.505)

blups.cor <- cor(ME.blups.505[,38:93], ME.blups.505[,1:36], use = "pairwise.complete.obs") # the 1st part is phenotype, 2nd part is the module info, calculate correlation between eigengene and trait value
blups.cor.P <- corPvalueStudent(blups.cor,nrow(vstMat.505.WGCNA.t))
blups.cor.sig <- blups.cor
blups.cor.sig[blups.cor.P>0.05] <- NA 
blups.cor.sig 

MEs.signed <- MEs.signed %>% dplyr::select(-ID) 
ME.blups.505 <- ME.blups.505 %>% dplyr::select(-ID)  
blups.505 <- pheno_data_505 %>% dplyr::select(-ID)  
save(blups.505, file = "~/505/network_analysis/output/blups.505.Rdata")

# plot it 
# Will display correlations and their p-values
png(filename = "~/505/network_analysis/output/Module-trait_heatmap_505.png",width=20, height=16, units="in", res=300)
par(mar = c(8, 18, 3, 3)); 
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = blups.cor,
               yLabels = names(blups.505),
               xLabels = names(MEs.signed), 
               xSymbols = names(MEs.signed),
               ySymbols = names(blups.505),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = signif(blups.cor.sig,2),
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("505 Module-trait relationships")) # this figure shows the correlation between traits and eigene gene of different modules, only correlation with P-value smaller than 0.05 were shown with correlation value 

dev.off()    

# again follow Julin's method 
# Arbitrary, but let's take the max and min for each trait (so long as they are significant)

blups.cor.5 <- blups.cor.sig  
dim(blups.cor.5) # 56 36 

cor.top <- t(apply(blups.cor.5,1,function(x) { # for each trait 
  maxx = max(x,na.rm=TRUE) 
  minx = min(x,na.rm=TRUE)
  ifelse(x == maxx | x == minx, x, NA) # only keep the largest and smallest correlation for each trait 
}
)) 
dim(cor.top) # 56 36  

cor.top <- cor.top[,apply(cor.top,2,function(x) !all(is.na(x)))] # keep modules which have significant correlations with traits
cor.top %>% dim() # 56 29 
write.csv(cor.top,"~/505/network_analysis/output/Eigen_trait_cor_505_threshold.csv") 

# write the Eigen genes
head(MEs.signed[,colnames(cor.top)]) 
MEs.signed[,colnames(cor.top)] %>% dim # get 29 modules which are significantly correlated with traits 
write.csv(MEs.signed[,colnames(cor.top)], file = "~/505/network_analysis/output/Top_Eigen_genes_505.csv") 
```

### looks like there are so many modules correlated with different traits, what if I use unsigned network 
```{r}
load("~/505/network_analysis/output/MEs_505.Rdata")
load("~/505/network_analysis/output/pheno_data_505.Rdata")

# include triat value  
dim(MEs) # 38 moduels 
rownames(MEs) <- rownames(vstMat.505.WGCNA.t)
MEs$ID <- rownames(MEs.signed)

# merge trait and module values 
MEs$ID <- gsub("\\_|\\-", "", MEs$ID)
MEs$ID <- gsub("(505)(K)([[:digit:]]+)", "ID_\\1_\\2\\3", MEs$ID)

# There are two K200 sequencing info (this line was sequenced twice), modify the name, change the 1st K200 to K200-1, the 2nd to K200-2, for both module and trait info  
MEs$ID[15] <- "ID_505_K200-1"
MEs$ID[16] <- "ID_505_K200-2"

pheno_data_505$ID[16] <- "ID_505_K200-1"
pheno_data_505$ID[17] <- "ID_505_K200-2"

ME.blups.505 <- 
MEs %>% 
  left_join(pheno_data_505, by = "ID") 

dim(ME.blups.505) # 38 module + 56 traits + 1 "line"
colnames(ME.blups.505)

blups.cor <- cor(ME.blups.505[,40:95], ME.blups.505[,1:38], use = "pairwise.complete.obs") # the 1st part is phenotype, 2nd part is the module info, calculate correlation between eigengene and trait value
blups.cor.P <- corPvalueStudent(blups.cor,nrow(vstMat.505.WGCNA.t))
blups.cor.sig <- blups.cor
blups.cor.sig[blups.cor.P>0.05] <- NA 
blups.cor.sig 

MEs <- MEs %>% dplyr::select(-ID) 
ME.blups.505 <- ME.blups.505 %>% dplyr::select(-ID)  
blups.505 <- pheno_data_505 %>% dplyr::select(-ID)  
save(blups.505, file = "~/505/network_analysis/output/blups.505.unsigned.Rdata")

# plot it 
# Will display correlations and their p-values
png(filename = "~/505/network_analysis/output/Module-trait_heatmap_505_unsigned.png",width=20, height=16, units="in", res=300)
par(mar = c(8, 18, 3, 3)); 
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = blups.cor,
               yLabels = names(blups.505),
               xLabels = names(MEs), 
               xSymbols = names(MEs),
               ySymbols = names(blups.505),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = signif(blups.cor.sig,2),
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("505 Module-trait relationships")) # this figure shows the correlation between traits and eigene gene of different modules, only correlation with P-value smaller than 0.05 were shown with correlation value 

dev.off()     
```

### do GWAS on eigene gene 
```{r}

```

### GWAS on all traits  
```{r}

```

### compare the GWAS result between traits and their associated eigene genes 
```{r}

```








