---
title: "505_network_WGCNA"
output: html_document
---

purpose of this script is to do WGCNA network anaysis using 505 SNP data: identify the different modules and check their correlation with pheno data.  

I did network analysis for our 414 F2 population using SNP data (https://github.com/leejimmy93/KIAT/blob/master/F2/F2_network.Rmd & https://github.com/leejimmy93/KIAT/blob/master/F2/F2_network_batch_corrected.Rmd), however, the identified gene co-expression modules correlated with different traits seems to arise from the genetic linkage instead of expression, so we want to see whether we have the same problem for 505. 

### load libs
```{r}
library(WGCNA) 
library(tidyverse) 
library(edgeR) 
library(DESeq2)
library(magrittr) 
library(gplots) 
library(reshape2)
library(sva) 
options(stringsAsFactors = FALSE) 
```

### load expression data, normalize, and correct for batch effect 
```{r}
# load data 
read_count_505 <- read.table("~/505/data/late_silique/read.count.505.tsv", header=T, row.names=1)
dim(read_count_505) # 101040    131 
read_count_505[1:10, 1:10]

# filter low expression tags 
keep <- rowSums(cpm(read_count_505) > 1) >= 3 # 1cpm in at least 1/4 of the samples 
read_count_505 <- read_count_505[keep,]
dim(read_count_505) # 60222   131

colnames(read_count_505) <- gsub("Sample_", "", colnames(read_count_505))
colnames(read_count_505) <- gsub("\\.", "\\-", colnames(read_count_505))

# add batch information 
sample_des_c <- read.csv("~/505/data/phenotype/batch_c.csv", stringsAsFactors=F, header=T, na.strings = "")
sample_des_d <- read.csv("~/505/data/phenotype/batch_d.csv", stringsAsFactors=F, header=T, na.strings = "")
sample_des_e <- read.csv("~/505/data/phenotype/batch_e.csv", stringsAsFactors=F, header=T, na.strings = "")

batch_info <- data.frame(batch = c(rep("C", nrow(sample_des_c)), 
                                  rep("D", nrow(sample_des_d)), 
                                  rep("E", nrow(sample_des_e))), 
                         sample_ID = c(sample_des_c$Sample.ID, sample_des_d$Sample.ID, sample_des_e$Sample.ID))

batch_info <- batch_info[match(colnames(read_count_505), batch_info$sample_ID),] 
all(batch_info$sample_ID == colnames(read_count_505)) # True 

# sample description 
sample.505 <- data.frame(batch = factor(batch_info$batch), 
                        genotype = factor(batch_info$sample_ID))

# vst transformation 
dds.505 <- DESeqDataSetFromMatrix(countData = round(read_count_505), 
                                  colData = sample.505, 
                                  design = ~genotype)

vsd.505 <- varianceStabilizingTransformation(dds.505)
vstMat.505 <- assay(vsd.505)
colnames(vstMat.505) <- colnames(read_count_505) 

# MDS to check batch effect 
vstMat.505 %>% dim() #  60222   131 
test1 <- as.matrix(dist(t(vstMat.505)))  

test2 <- as.data.frame(cmdscale(test1))
head(test2) #now we have 2 dimensions 
test2$sample_ID <- rownames(test2)

P.batch.before <- 
test2 %>% 
  left_join(batch_info, by = "sample_ID") %>% 
  ggplot() + 
  geom_point(aes(x = V1, y= V2, color = as.factor(batch))) # looks like there is batch effect on PC2

# correct batch effect 
vstMat.505.batch.corrected <- removeBatchEffect(vstMat.505, batch = batch_info$batch) 
test1 <- as.matrix(dist(t(vstMat.505.batch.corrected)))  
test2 <- as.data.frame(cmdscale(test1))
test2$sample_ID <- rownames(test2)

P.batch.after <- 
test2 %>% 
  left_join(batch_info, by = "sample_ID") %>% 
  ggplot() + 
  geom_point(aes(x = V1, y= V2, color = as.factor(batch))) # great, batch effect removed

ggsave(P.batch.before, filename = "~/505/network_analysis/output/P.batch.before.png", width = 9, height = 4)
ggsave(P.batch.after, filename = "~/505/network_analysis/output/P.batch.after.png", width = 9, height = 4)

# get genes with large variation across samples using CV, The coefficient of variation (CV) is defined as the ratio of the standard deviation sigma  to the mean mu (standerdized deviation):
# https://labs.genetics.ucla.edu/horvath/CoexpressionNetwork/Rpackages/WGCNA/faq.html

vstMat.505.batch.corrected <- as.data.frame(vstMat.505.batch.corrected) 

gene.CV <- 
sapply(rownames(vstMat.505.batch.corrected), function(i) {
  sd(as.numeric(vstMat.505.batch.corrected[i,]), na.rm = T)/mean(as.numeric(vstMat.505.batch.corrected[i,]), na.rm = T) * 100
}) 

vstMat.505.batch.corrected$CV <- as.numeric(gene.CV) 
summary(vstMat.505.batch.corrected$CV) 
hist(vstMat.505.batch.corrected$CV, breaks = 30)

save(vstMat.505.batch.corrected, file = "~/505/network_analysis/output/vstMat.505.batch.corrected.Rdata")

vstMat.505.batch.corrected.largeVar<-
  vstMat.505.batch.corrected %>% 
  filter(CV>15) %>% # only the last quantile was left for co-expression analysis 
  dplyr::select(-CV) 

rownames(vstMat.505.batch.corrected.largeVar) <- rownames(vstMat.505.batch.corrected[vstMat.505.batch.corrected$CV >15,]) # keep the last quantile & make sure include the two FAE1 genes "BnaA08g11130D" & "BnaC03g65980D"


dim(vstMat.505.batch.corrected.largeVar) # 13454   131  

save(vstMat.505.batch.corrected.largeVar, file = "~/505/network_analysis/output/vstMat.505.batch.corrected.largeVar.Rdata")
```

### 1) follow Julin's method for network analysis 

### WGCNA for network  

https://labs.genetics.ucla.edu/horvath/CoexpressionNetwork/Rpackages/WGCNA/Tutorials/FemaleLiver-02-networkConstr-man.pdf  

### data QC 
```{r}
enableWGCNAThreads()

load("~/505/network_analysis/output/vstMat.505.batch.corrected.largeVar.Rdata")  
vstMat.505.WGCNA <- vstMat.505.batch.corrected.largeVar
dim(vstMat.505.WGCNA)

# transform expression data 
head(vstMat.505.WGCNA[,1:6])
vstMat.505.WGCNA.t <- t(vstMat.505.WGCNA)
head(vstMat.505.WGCNA.t[,1:6]) 

# check sample quality, this is to see whether there are many missing entry for certain gene, also check variance  
gag.CR <- goodSamplesGenes(vstMat.505.WGCNA.t, verbose = 3)
gag.CR$allOK

# cluster samples to look for outliers based on gene expression data
# hclust 1stly calculate pairwise dissimilarity between each pair of observations, the observations that are most similar to each other are fused, then the next most similar observations... For clusters with multiple observations, use linkage to quantitate their dissimlarity, including complete, single, average, and centroid linkage. 
sampleTREE.CR <- hclust(dist(vstMat.505.WGCNA.t), method = "average") # average is the mean intercluster dissmilarity 
plot(sampleTREE.CR,cex=.6) 

### MDS plot to make sure I get the right data to start with 
vstMat.505.WGCNA.t %>% dim()
vstMat.505.WGCNA.t[1:10, 1:10]
rnames <- rownames(vstMat.505.WGCNA.t)

genDist <- as.matrix(dist(as.matrix(vstMat.505.WGCNA.t)))

#perform the multi-dimensional scaling
geno.mds <- as.data.frame(cmdscale(genDist))
head(geno.mds) #now we have 2 dimensions
dim(geno.mds) # 131 2 
plot(geno.mds)  

load("~/505/output/131_sample/sample_des_revised.Rdata")
sample_des_revised 

geno.mds$ID <- rownames(geno.mds)

geno.mds <- 
geno.mds %>% 
  left_join(sample_des_revised, by = c("ID" = "Sample.ID"))  

library(ggrepel)
p.mds.WGCNA <- ggplot(data=geno.mds, aes(V1, V2, color=Sample.Description.Origin)) + theme_bw() + 
  geom_point(size=1) + 
# p.mds.country.2 <- p.mds.country.2 + scale_color_brewer(type="qual",palette="Set1")
  scale_color_manual(values = c("lightblue", "orange", "red", "green", "blue", "grey", "purple", "pink", "darkred", "black")) + 
  geom_text_repel(aes(x=V1, y=V2, label=factor(ID))) + 
  theme(legend.position = c(0.1, 0.8), legend.background = element_rect(fill="lightgrey", size=0.5, linetype="solid"), axis.title.x=element_blank(), axis.title.y =element_blank()) + 
  labs(title="MDS of 505 collection") + theme(plot.title = element_text(hjust = 0.5, size = 16, face="bold")) 
p.mds.WGCNA

ggsave(p.mds.WGCNA, filename="~/505/network_analysis/output/p.mds.WGCNA.png", height=10, width=10)  
save(vstMat.505.WGCNA.t, file = "~/505/network_analysis/output/vstMat.505.WGCNA.t.Rdata")
```

### WGCNA module detection 
```{r}
load("~/505/network_analysis/output/vstMat.505.batch.corrected.largeVar.Rdata")  

# soft threshold:  In an unsigned coexpression network, modules correspond to clusters of genes with high absolute correlations. In a signed network, modules correspond to positively correlated genes.
powers <- c(c(1:10), seq(from = 12, to=30, by=2))
sft <- pickSoftThreshold(vstMat.505.WGCNA.t, powerVector = powers, verbose = 5, networkType = "signed")

sizeGrWindow(9, 5)
par(mfrow = c(1,2)) 
cex1 <- 0.9 
# Scale-free topology fit index as a fCRction of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red")
# this line corresponds to using an R^2 cut-off of h
abline(h=0.90,col="red")
# Mean connectivity as a fCRction of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red") 

# choose 14 
softPower <- 14
adjacency <- adjacency(vstMat.505.WGCNA.t, power = softPower, type = "signed")
# Turn adjacency into topological overlap, topological overlap leads to a network distance measure
dim(adjacency) # 13454 13454 
system.time(TOM <- TOMsimilarity(adjacency)); # Calculation of the topological overlap matrix from a given adjacency matrix. takes 30 mins 
dissTOM <- 1-TOM

# Call the hierarchical clustering fCRction
geneTree <- hclust(as.dist(dissTOM), method = "average") 
# Plot the resulting clustering tree (dendrogram)
sizeGrWindow(12,9)
plot(geneTree, xlab="", sub="", main = "Gene clustering on TOM-based dissimilarity",
     labels = FALSE, hang = 0.04) 

# define modules 
# We like large modules, so we set the minimum module size relatively high:
minModuleSize <- 30;
# Module identification using dynamic tree cut:
dynamicMods <- cutreeDynamic(dendro = geneTree, distM = dissTOM,
                             deepSplit <- 2, pamRespectsDendro = FALSE,
                             minClusterSize = minModuleSize);
table(dynamicMods)
table(dynamicMods) %>% length() # 43 modules 

# Convert numeric lables into colors
dynamicColors = labels2colors(dynamicMods)
table(dynamicColors)
# Plot the dendrogram and colors CRderneath
sizeGrWindow(8,6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05,
                    main = "Gene dendrogram and module colors")

# merge similar modules
# Calculate eigengenes 
MEList <- moduleEigengenes(vstMat.505.WGCNA.t, colors = dynamicColors) # Calculates module eigengenes (1st principal component) of modules in a given single dataset. 
MEs <- MEList$eigengenes
# Calculate dissimilarity of module eigengenes
MEDiss <- 1-cor(MEs);
# Cluster module eigengenes
METree <- hclust(as.dist(MEDiss), method = "average");
# Plot the result
sizeGrWindow(7, 6)
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")

# merge with correlation > 0.8

MEDissThres = 0.2
# Plot the cut line into the dendrogram
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")
abline(h=MEDissThres, col = "red")
# Call an automatic merging fCRction
merge = mergeCloseModules(vstMat.505.WGCNA.t, dynamicColors, cutHeight = MEDissThres, verbose = 3)
# The merged module colors
mergedColors = merge$colors
# Eigengenes of the new merged modules:
mergedMEs = merge$newMEs

# compare pre and post merge

sizeGrWindow(12, 9)
#pdf(file = "Plots/geneDendro-3.pdf", wi = 9, he = 6)
plotDendroAndColors(geneTree, cbind(dynamicColors, mergedColors),
c("Dynamic Tree Cut", "Merged dynamic"),
dendroLabels = FALSE, hang = 0.03,
addGuide = TRUE, guideHang = 0.05)
#dev.off()

# Rename to moduleColors
moduleColors = mergedColors
# Construct numerical labels corresponding to the colors
colorOrder = c("grey", standardColors(50));
moduleLabels = match(moduleColors, colorOrder)-1;
MEs = mergedMEs # 39 modules 

save(MEs, moduleColors, file = "~/505/network_analysis/output/MEs_505_signed.Rdata")
```

###  phenotype data for 505 
```{r}
pheno.505 <- read_csv("~/505/data/505_Phenotype_Bnapus_38_traits.csv") # 1st year data, 2016?
pheno.505 <- pheno.505[,!apply(pheno.505,2,function(x) all(is.na(x)))]
pheno.505 %>% dim() 
pheno.505.t <- as.data.frame(t(pheno.505[,-1]))
colnames(pheno.505.t) <- paste(pheno.505$PHENOTYPE, "year1", sep = "_")
pheno.505.t$ID <- rownames(pheno.505.t)

load("~/505/genomic_prediction/data/pheno_data_fatty_acid_all.Rdata") # 2nd year data, on FA related traits 
pheno_data_505 %>% dim()
colnames(pheno_data_505) <- paste(colnames(pheno_data_505), "year2", sep = "_")
colnames(pheno_data_505)[1] <- "ID"
pheno.505.t$ID
pheno_data_505$ID
pheno_data_505$ID <- gsub("\\_|\\-", "", pheno_data_505$ID)
pheno_data_505$ID <- gsub("(505)(K)([[:digit:]]+)", "ID_\\1_\\2\\3", pheno_data_505$ID)

# merge two dataset 
pheno_data_505 <- 
pheno_data_505 %>% 
  left_join(pheno.505.t) 

dim(pheno_data_505)   # 131 57 

# plot the data to see missing rate, distribution, etc... 
summary(pheno_data_505)
colnames(pheno_data_505)

par(mfrow=c(5, 4))

for (i in colnames(pheno_data_505)[2:21]) { 
  hist(pheno_data_505[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i) 
  } 
 
for (i in colnames(pheno_data_505)[22:41]) { 
  hist(pheno_data_505[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i) 
  } 
 
for (i in colnames(pheno_data_505)[42:57]) { 
  hist(pheno_data_505[,i], breaks = 15, col = "grey", border = "black", xlab = "", ylab = "", main = i, mgp=c(.1,.5,.0)) 
  } 

save(pheno_data_505, file = "~/505/network_analysis/output/pheno_data_505.Rdata")

# check correlation between phenotypes 
library(pheatmap)
pheno_data_505[1:10, 1:10]
trait_cor <- round(cor(pheno_data_505[,c(2,6:57)], use = "pairwise.complete.obs"), digits = 2) 

png(filename = "~/505/network_analysis/output/figure/trait_heatmap.png",width=220, height=200, units="mm", res=300)
par(mar = c(3, 3, 3, 3))
pheatmap(as.matrix(t(trait_cor)),
     # scale="row",
     clustering_method = "average", 
     # cellheight = 10,
     # cellwidth = 5,
     border_color=NA,
     fontsize_row = 6, 
     fontsize_col = 6, 
     legend = T
)   
dev.off()     
```

### plot several clusters gene expression value & their eigengenes 
```{r}
load("~/505/network_analysis/output/MEs_505_signed.Rdata")

moduleColors.signed %>% length()
MEs.signed 

load("~/505/network_analysis/output/vstMat.505.batch.corrected.largeVar.Rdata")  
vstMat.505.batch.corrected.largeVar %>% dim()

table(moduleColors.signed) 
dim(MEs.signed) # 30 

# check "white" module 
vstMat.505.signed <- 
  vstMat.505.batch.corrected.largeVar[moduleColors.signed == "orange",] %>% 
  t() %>% as.data.frame()

# plot heatmap 
pheatmap(as.matrix(t(vstMat.505.signed)),  
     cluster_rows = T,
     cluster_cols = T,
     clustering_method = "average", 
     # cellheight = 10,
     # cellwidth = 5,
     border_color=NA,
     fontsize_col = 8, 
     legend = T, 
     main = "signed"
) 

### unsigned modules 
load("~/505/network_analysis/output/MEs_505_unsigned.Rdata")

moduleColors %>% length()
MEs
table(moduleColors) 
length(MEs) # 30 

# check "white" module 
vstMat.505.unsigned <- 
  vstMat.505.batch.corrected.largeVar[moduleColors == "royalblue",] %>% 
  t() %>% as.data.frame()

# plot heatmap 
pheatmap(as.matrix(t(vstMat.505.unsigned)),  
     cluster_rows = T,
     cluster_cols = T,
     clustering_method = "average", 
     # cellheight = 10,
     # cellwidth = 5,
     border_color=NA,
     fontsize_col = 8, 
     legend = T, 
     main = "unsigned"
)  
```

### check to see the three genes I found from F2, BnaA08g12780D, BnaA08g11060D, and BnaA08g11140D 
```{r}
# BnaA08g11060D is not expressed at all 
read_count_505 <- read.table("~/505/data/late_silique/read.count.505.tsv", header=T, row.names=1)
read_count_505["BnaA08g11140D",]

# BnaA08g11060D and BnaA08g12780D were expressed but not varied a lot among lines 
load("~/505/network_analysis/output/vstMat.505.batch.corrected.Rdata")
load("~/505/network_analysis/output/pheno_data_505.Rdata") 

vstMat.505.batch.corrected["BnaA08g12780D",]$CV
vstMat.505.batch.corrected["BnaA08g11060D",]$CV 

test1 <- 
vstMat.505.batch.corrected [c("BnaA08g12780D", "BnaA08g11060D"),-15] %>% 
  dplyr::select(-CV) %>% 
  t() %>% as.data.frame()

test1$ID <- rownames(test1)

tmp <- 
test1 %>% 
  left_join(pheno_data_505)

cor(tmp$BnaA08g12780D, tmp$Erucic_acid_year1, use = "pairwise.complete.obs") # 0.240422 this is not very high correlation compared to what we found from network analysis 
cor(tmp$BnaA08g11060D, tmp$Erucic_acid_year1, use = "pairwise.complete.obs") # -0.1287613 
```




